### 2017-09-05 degraded-paper.Rmd
Move degraded paper outside of repo

---
title: "Algorithmic Approaches to Match Degraded Land Impressions"
author: "Eric Hare, Heike Hofmann, Alicia Carriquiry"
date: "9/5/2017"
bibliography: references.bib
output: 
  bookdown::pdf_document2:
    keep_tex: true
    toc: false
    includes:
      in_header: preamble.tex
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

\textbf{Abstract}

Bullet matching is a process used to determine whether two bullets may have been fired from the same gun barrel. Historically, this has been a manual process performed by trained forensic examiners. Recent work however has shown that it is possible to add statistical validity and objectivity to the procedure. In this paper, we build upon the algorithms explored in Automatic Matching of Bullet Lands [@2016arXiv160105788H] by formalizing and defining a set of features, computed on pairs of bullet lands, which can be used in machine learning models to assess the probability of a match. We then use these features to perform an analysis of the two Hamby [@hamby:2009] bullet sets (Set 252 and Set 44), to assess the presence of microscope operator effects in scanning. We also take some first steps to address the issue of degraded bullet lands, and provide a range of degradation at which the matching algorithm still performs well. Finally, we discuss generalizing land to land comparisons to full bullet comparisons as would be used for this procedure in a criminal justice situation.

```{r, echo=FALSE, message=FALSE, cache=FALSE}
library(ggplot2)
library(xtable)
library(RMySQL)
library(tidyverse)
library(bulletr)
library(gridExtra)
library(randomForest)
library(caret)
library(broom)
library(matrixcalc)

dbname <- "bullets"
user <- "buser"
password <- readLines("buser_pass.txt")
host <- "127.0.0.1"
port <- 3306

my_db <- src_mysql(dbname, host, port, user, password)
my_metadata <- tbl(my_db, "metadata")
my_metadata_derived <- tbl(my_db, "metadata_derived")
my_profiles <- tbl(my_db, "profiles")
my_signatures <- tbl(my_db, "signatures")
my_ccf <- tbl(my_db, "ccf")
```

# Background

Intense scrutiny has been focused on the process of bullet matching in recent years [e.g., @giannelli:2011]. Bullet matching, the process of determining whether two bullets could have been fired from the same gun barrel, has traditionally been performed without meaningful determination of error rates or statistical assessments of uncertainty [@NAS:2009]. There have been some attempts towards developing mathematical and statistical approaches to bullet matching.One such attempt was the definition of CMS, the Consecutively Matching Striae [@biasotti:1959], with a cutoff of six to separate matches from non-matches. Still, rigorous assessments of the applicability of such cutoffs have not to this point been described [@pcast2016].

Recently, several authors have addressed these well-known shortcomings. Focusing on firing pin impressions and breech faces, @riva:2014 have described an automated algorithm using 3D images that enables comparison between pairs of exemplars. Other examples of work in this and related areas include @petraco:2012, @chu:2011, @vorburger:2011, and others. In our approach to this problem, Automatic Matching of Bullet Lands, we used the Hamby 252 set [@hamby:2009] to train and develop a random forest in order to provide a matching probability for two bullet lands [@2016arXiv160105788H]. While the algorithm had a very strong performance on this set, some limitations were immediately clear. For instance, performance was assessed only on this single set of 35 bullets fired from a consecutively manufactured set of only ten gun barrels. Each of these bullets was part of controlled study, and the full lands were available for matching. While there were some data quality issues, this was still a near ideal test case for the algorithm.

Real world applications of bullet matching often involve the recovery of fragments of bullets from the crime scene. Traditional features used in forensic examination work well for a full land, but there has been less investigation into their performance in the case of a fragmented land. For example, the CMS is naturally limited by the portion of the land that can be recovered.

In this paper, we take steps to address these and other concerns. Specifically, we begin by reviewing features from the literature, computed on pairs of bullet lands, and presenting some of our own features. We propose an approach to  standardize the featues, to account for the fact that only a portion of the land impression may be recovered from the crime scene. With the standardized features, we tackle two issues that were not addressed in  [@2016arXiv160105788H]. The first is the effect of the microscope operator  on the resulting images and consequent algorithm performance. The second issue has to do with the robustness of the land matching algorithm in [@2016arXiv160105788H] relative to the degree of degradation of the questioned land impression. Finally, we describe some of the initial steps toward generalizing a matching algorithm based on land-to-land comparisons, to one based on bullet-to-bullet comparisons, as would be of interest in a real world application of these ideas.

# Feature Standardization

To start, we introduce a standardized version of each of the features used in the matching routine proposed by [@2016arXiv160105788H]. These features are computed on *aligned pairs of bullet land impressions* rather than on individual lands. This enables us, for instance, to compute the number of matching striae. Table \ref{tab:ccf1} and Table \ref{tab:ccf2} provide an example of six land-to-land comparisons and the derived features for these comparisons. The two `profile_id` columns identify a particular land from the two Hamby datasets, and the remaining columns are the derived features for these comparisons.

```{r, echo=FALSE, results='asis'}
result <- my_ccf %>%
    filter(compare_id == 4) %>%
    head %>%
    collect %>%
    select(2:8) %>%
    as.data.frame

print(xtable(result, digits = c(0, 0, 0, 4, 4, 4, 4, 4), label = "tab:ccf1", caption = "A sample of six land-to-land comparisons, and derived features for these comparisons."), comment = FALSE, include.rownames = FALSE)
```

```{r, echo=FALSE, results='asis'}
result2 <- my_ccf %>%
    filter(compare_id == 4) %>%
    head %>%
    collect %>%
    select(9:ncol(.)) %>%
    as.data.frame

print(xtable(result2, digits = c(4, 4, 4, 4, 4, 4, 4, 4), label = "tab:ccf2", caption = "The remaining derived features for the previous six land-to-land comparisons."), comment = FALSE, include.rownames = FALSE)
```

We generalize the definitions of these features to account for the possibility that we may be handling degraded bullet lands, where only fragments can be recovered. The definition of each feature is given below, where $f(t)$ represents the height values of the first profile, and $g(t)$ the height values of the second:

- **ccf** (%) is the maximum value of the Cross-Correlation function evaluated at the optimal alignment. The Cross-Correlation function is defined as $C(\tau) = \int_{-\infty}^{\infty} f(t)g(t + \tau)dt$ where $\tau$ represents the the lag of the second signature [@vorburger:2011]. 
- **rough_cor** (%) is a new feature that quantifies the correlation between the two signatures after performing a second LOESS smoothing stage and then subtracting the result from the original signatures. This attempts to model the roughness of the surface after removing structure such as waviness.
- **lag** (mm) Is the optimal lag for the ccf value.
- **D** (mm) is the Euclidean vertical distance between each height value of the aligned signatures. This is defined as $D^2 = \frac{1}{\text{\#}t}\sum_t \left[f(t) - g(t)\right]^2$. This is a measure of the total variation between two functions [@clarkson1933definitions].
- **sd_D** (mm) provides the standard deviation of the values of *D* from above.
- **signature_length** (mm) is the overall length of the smallest of the two aligned signatures.
- **overlap** (%) provides the percentage of the two signatures that overlap after the alignment stage.
- **matches** (per mm) is the number of matching peaks/valleys (striae) per millimeter of the overlapping portion of the aligned signatures.
- **mismatches** (per mm) is the number of mismatching peaks/valleys (striae) per millimeter of the overlapping portion of the aligned signatures.
- **cms** (per mm) is the number of consecutively matching peaks/valleys (striae) per millimeter of the overlapping portion of the aligned signatures [@biasotti:1959, @thompson:2013].
- **non_cms** (per mm) is the number of consecutive mismatching peaks/valleys (striae) per millimeter of the overlapping portion of the aligned signatures.
- **sum_peaks** (per mm) is the the sum of the average heights of matched striae.

The features that are expressed on a per millimeter level are intended to support the degraded land case, as discussed earlier. Note that the computation differs slightly depending on the feature. For example, to standardize the number of matches, the first count the raw number of matching striae, and then divide this number by the length of the overlapping region of the two lands (`overlap` from above). In most cases, the overlapping region will be very close to the length of the smaller signature. But depending on the alignment, this may not always be true. This ensures that we do not punish a particular cross-comparison for having a smaller region in which matches could occur. On the other hand, the number of mismatches is divided by the total length of the two aligned signatures, since mismatched striae can occur even in the non-overlapping region of the two signatures.

The `rough_cor` or Roughness Correlation is derived by performing a second smoothing step, and subtracting the result from the original signatures. This creates a new signature which eliminates some of the overall structure, allowing global deformations to have less of an influence on the model output. Where the roughness correlation is most useful is in a scenario like Figure \@ref(fig:roughcorgood). This figure shows the alignment of profile 40977 with 47600. The top panel shows the smoothed signatures. The middle panel overlays a LOESS fit to the average of the two signatures. Finally, to derive the roughness correlation, this LOESS is subtracted from the original signature to create a new set of roughness residuals, which are then given in the bottom panel. Note that these two profiles do not match, yet the ccf is 0.7724. The roughness correlation (-0.0324) correctly indicates the lack of matching. The roughness correlation acts as a check against false positives which can arise when there are significant deformations in the overall structure, as in the case with both these profiles.

```{r roughcorgood, echo=FALSE, fig.cap='Alignment of profile 40977 with 47600. The top panel shows the smoothed signatures. The middle panel overlays a LOESS fit to the average of the two signatures. Finally, to derive the roughness correlation, this LOESS is subtracted from the original signature to create a new set of roughness residuals, which are then given in the bottom panel. Note that these two profiles do not match, yet the ccf is 0.7724. The roughness correlation (-0.0324) correctly indicates the lack of matching.'}
bullets_smoothed <- my_signatures %>% filter(run_id == 3) %>% collect(n = Inf)

br1 <- filter(bullets_smoothed, profile_id == 40977) %>%
    dplyr::select(-run_id) %>%
    rename(bullet = profile_id) %>%
    filter(!is.na(l30))
br2 <- filter(bullets_smoothed, profile_id == 47600) %>%
    dplyr::select(-run_id) %>%
    rename(bullet = profile_id) %>%
    filter(!is.na(l30))

res <- bulletGetMaxCMS(br1, br2, column = "l30", span = 5)

prof1_zeroed <- br1 %>% 
    mutate(y = y - min(y),
           profile_id = factor(bullet))

prof2_zeroed <- br2 %>%
    mutate(y = y - min(y)) %>%
    mutate(y = y + res$lag,
           profile_id = factor(bullet))

final_profs <- rbind(prof1_zeroed, prof2_zeroed)

doublesmoothed <- final_profs %>%
    group_by(y) %>%
    mutate(avgl30 = mean(l30, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(smoothavgl30 = smoothloess(x = y, y = avgl30, span = 0.3),
           l50 = l30 - smoothavgl30)

ys <- intersect(prof1_zeroed$y, prof2_zeroed$y)

final_doublesmoothed <- doublesmoothed %>%
    filter(y %in% ys)

p1 <- ggplot(data = final_profs, aes(x = y, y = l30, colour = profile_id)) +
    geom_line() +
    theme_bw()

p2 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l30, colour = profile_id)) +
    geom_line() +
    geom_smooth(inherit.aes = FALSE, aes(x = y, y = avgl30), colour = "purple") +
    theme_bw()

p3 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l50, colour = profile_id)) +
    geom_line() +
    theme_bw()

grid.arrange(p1, p2, p3, nrow = 3)
```

In a typical comparison between two profiles, such as in Figure \@ref(fig:roughcorfine), the roughness correlation does not meaningfully impact the matching probability given the presence of the ccf in the model. In this figure, we see the alignment of profile 8752 with profile 136676. In this case, the waviness or the deformation pattern in the signatures is less pronounced, and hence the resulting roughness signature resembles the original signature more closely. These profiles match, and both ccf (0.6891) and rough_cor (0.7980) provide values indicative of matching.

```{r roughcorfine, echo=FALSE, fig.cap='Alignment of profile 8752 with profile 136676. In this case, the waviness or the deformation pattern in the signatures is more minor, and hence the resulting roughness signature resembles the original signature more closely. These profiles match, and both ccf (0.6891) and rough_cor (0.7980) provide values indicative of matching.'}
br1 <- filter(bullets_smoothed, profile_id == 8752) %>%
    dplyr::select(-run_id) %>%
    rename(bullet = profile_id) %>%
    filter(!is.na(l30))
br2 <- filter(bullets_smoothed, profile_id == 136676) %>%
    dplyr::select(-run_id) %>%
    rename(bullet = profile_id) %>%
    filter(!is.na(l30))

res <- bulletGetMaxCMS(br1, br2, column = "l30", span = 5)

prof1_zeroed <- br1 %>% 
    mutate(y = y - min(y),
           profile_id = factor(bullet))

prof2_zeroed <- br2 %>%
    mutate(y = y - min(y)) %>%
    mutate(y = y + res$lag,
           profile_id = factor(bullet))

final_profs <- rbind(prof1_zeroed, prof2_zeroed)

doublesmoothed <- final_profs %>%
    group_by(y) %>%
    mutate(avgl30 = mean(l30, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(smoothavgl30 = smoothloess(x = y, y = avgl30, span = 0.3),
           l50 = l30 - smoothavgl30)

ys <- intersect(prof1_zeroed$y, prof2_zeroed$y)

final_doublesmoothed <- doublesmoothed %>%
    filter(y %in% ys)

p1 <- ggplot(data = final_profs, aes(x = y, y = l30, colour = profile_id)) +
    geom_line() +
    theme_bw()

p2 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l30, colour = profile_id)) +
    geom_line() +
    geom_smooth(inherit.aes = FALSE, aes(x = y, y = avgl30), colour = "purple") +
    theme_bw()

p3 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l50, colour = profile_id)) +
    geom_line() +
    theme_bw()

grid.arrange(p1, p2, p3, nrow = 3)
```

```{r, echo=FALSE, cache=FALSE}
con <- dbConnect(MySQL(), user = user, password = password,
                 dbname = dbname, host = host)

all_bullets_metadata <- dbReadTable(con, "metadata")
my_matches <- dbReadTable(con, "matches") %>% mutate(match = 1)
profiles <- dbReadTable(con, "profiles")
ccf <- dbReadTable(con, "ccf") %>% filter(compare_id == 4)

CCFs_withlands <- ccf %>%
    left_join(dplyr::select(profiles, profile_id, land_id), by = c("profile1_id" = "profile_id")) %>%
    left_join(dplyr::select(profiles, profile_id, land_id), by = c("profile2_id" = "profile_id")) %>%
    left_join(my_matches, by = c("land_id.x" = "land1_id", "land_id.y" = "land2_id")) %>%
    left_join(dplyr::select(all_bullets_metadata, land_id, study, barrel, bullet, land), by = c("land_id.x" = "land_id")) %>%
    left_join(dplyr::select(all_bullets_metadata, land_id, study, barrel, bullet, land), by = c("land_id.y" = "land_id")) %>%
    filter(study.x != "Cary", study.y != "Cary") %>%
    filter(study.y != "Hamby44" | barrel.y != "E") %>%
    filter(study.x != "Hamby44" | barrel.x != "E") %>%
    arrange(study.x, study.y) %>%
    mutate(match = as.logical(replace(match, is.na(match), 0)))
```

We can observe the distributions of both CCF and the Roughness Correlation side by side, differentiating between known matches and known non-matches. Figure \@ref(fig:roughcordist) shows this. It can be seen that the separation of the two groups along both variables is quite strong, but many known-matches have low values for each. The distributions of known non-matches, on the other hand, is relatively symmetric and normal.

```{r roughcordist, echo=FALSE, fig.cap='Distributions of the Roughness Correlation compared to the CCF for known matches and known non-matches'}
cor_compare <- CCFs_withlands %>%
  select(ccf, rough_cor, match) %>%
  gather(key = feature, value = value, ccf, rough_cor) %>%
  mutate(match = factor(match, labels = c("Known non-matches (KNM)", "Known matches (KM)")))

qplot(value, group=match, geom="density", data=cor_compare, fill=match, 
      alpha=I(0.6), colour=I("grey20")) + 
  facet_wrap(~feature, nrow = 2) +
  theme_bw() + ylab("") + xlab("") +
  scale_fill_brewer("Bullet Land Pairs", palette="Paired") +
  theme(legend.background = element_rect(colour="grey75"),
        axis.title.y=element_blank(),
        plot.margin = unit(c(0,0,0,0), unit="cm"))
```

# Model Training

Using these features, we can train a randomForest [@randomForest] model which attempts to predict whether two lands match given the value of the features.  There are currently three studies for bullets included in the NIST ballistics database (REF). Those are Hamby (Set 252), Hamby (Set 44), and Cary. For purposes of analysis we describe in this paper, we exclude the Cary bullets from consideration, because the study was designed to assess the persistence of striation markings over a series of fires from the same barrel. Thus, every Cary bullet is a known match to every other Cary bullet, although early firings do not produce the same markings that later firings do. Hence, we will consider Hamby (Set 252) and Hamby (Set 44) only.

We can now train the forest using the features we defined earlier. We split the data into an 80% training, 20% testing sets to assess the algorithm's out of sample performance, using the `caret` package [@caretpkg]. Table \ref{tab:avgforest} displays the results in the form of a confusion matrix on the test set, averaged over ten independent random forests trained on ten random data subsets. There are a total of 83,028 land-to-land comparisons, of which 1208 are among known matching land impressions and 81,820 are among known non-matching land impressions. It can be seen that false positives are exceedingly rare, but false negatives occur more frequently (approximately 65 false negative land to land comparisons on the test set, compared with an average of less than four false positives).

```{r, echo=FALSE, results='asis'}
includes <- setdiff(names(CCFs_withlands), c("compare_id", "profile1_id", "profile2_id",
                                         "study.x", "study.y", "barrel.x", "barrel.y",
                                         "bullet.x", "bullet.y", "land.x", "land.y",
                                         "land_id.x", "land_id.y", "signature_length"))

set.seed(20170222)
trainIndex <- createDataPartition(1:nrow(CCFs_withlands), p = 0.8, list = TRUE, times = 10)
match_result <- trainIndex %>% purrr::map_df(function(index) {
    CCFs_train <- CCFs_withlands[index,]
    CCFs_test = setdiff(CCFs_withlands, CCFs_train)

    rtrees <- randomForest(factor(match) ~ ., data = CCFs_train[,includes], ntree = 300)

    CCFs_test$forest <- predict(rtrees, newdata = CCFs_test, type = "prob")[,2]
    imp <- data.frame(importance(rtrees))
    confmat <- as.data.frame(xtabs(~(forest > 0.5) + match + study.x + study.y, data = CCFs_test))
    names(confmat) <- c("Prediction", "Actual", "study.x", "study.y", "Count")

    mydf <- confmat %>%
        mutate(Study = paste(study.x, study.y, sep = "_"),
               Result = rep(c("True Negative", "False Positive", "False Negative", "True Positive"), times = length(unique(Study)))) %>%
        mutate(Study = replace(Study, which(Study == "Hamby44_Hamby252"), "Hamby252_Hamby44")) %>%
        dplyr::select(Study, Result, Count) %>%
        group_by(Study, Result) %>%
        summarise(Count = sum(Count)) %>%
        as.data.frame

    return(mydf)
}) 
match_result <- cbind(id = rep(1:10, each = nrow(match_result) / 10), match_result)
match_result_wide <- match_result %>%
    spread(key = Result, value = Count)

result <- match_result %>%
    group_by(Result, id) %>%
    summarise(Count = sum(Count)) %>%
    group_by(Result) %>%
    summarise(Count = mean(Count))
sens_result <- result$Count[result$Result == "True Positive"] / (result$Count[result$Result == "True Positive"] + result$Count[result$Result == "False Negative"])
spec_result <- result$Count[result$Result == "True Negative"] / (result$Count[result$Result == "True Negative"] + result$Count[result$Result == "False Positive"])

print(xtable(result, digits = c(0, 0, 1), label = "tab:avgforest", caption = "The average confusion matrix for the 10 random forests. It can be seen that false positives are exceedingly rare, but false negatives occur more frequently."), comment = FALSE, include.rownames = FALSE, table.placement = "H")
```

These results suggest that our algorithm is  too conservative in predicting a match when in fact the bullets were fired from the same gun barrel. We can break down the confusion matrix by the study from which each of the two land impressions originated. Table \ref{tab:avgforeststudy} shows the average confusion matrix for the 10 random forests, broken down by study. It can be seen that Hamby252 to Hamby252 comparisons exhibit the fewest errors, while Hamby252 to Hamby44 comparisons exhibit the most errors on average. This intuitively makes some sense given the potential presence of scanner operator effects, which we address further in this section.

```{r, echo=FALSE, results='asis'}
result <- match_result %>%
    group_by(Study, Result) %>%
    summarise(Count = mean(Count)) %>%
    spread(key = Result, value = Count)

result_wide_row <- result
result_wide_row[,-1] <- t(apply(result_wide_row[,-1], 1, function(x) paste0(round(100 * x / sum(x), digits = 2), "%")))

print(xtable(result_wide_row, label = "tab:avgforeststudy", caption = "The average confusion matrix for the 10 random forests, broken down by study. It can be seen that Hamby252 to Hamby252 comparisons exhibit the fewest errors, while Hamby252 to Hamby44 comparisons exhibit the most on average."), comment = FALSE, include.rownames = FALSE, table.placement = "H")
```

# Feature Robustness

As a first stage to assessing feature robustness, we produce parallel coordinate plots of the various features based on true positives, true negatives, false positives, and false negatives. Figures \ref{fig:pcp} and \ref{fig:pcp2} display these plots. The means of the true positive and the false negative groups are shown, respectively, in the two figures. False positives tend to occur with anomalously high values of `sum_peaks`, `matches`, and `ccf`, while false negatives tend to resemble very closely the feature distribution of the true negatives, with a slightly higher average `ccf`.

```{r pcp, echo=FALSE, message=FALSE, fig.cap="Parallel coordinate plot of the features based on the random forest confusion matrix for true and false positives.", fig.height=6, fig.width=10}
set.seed(20170222)
CCFs_train <- sample_frac(CCFs_withlands, size = .8)
CCFs_test = setdiff(CCFs_withlands, CCFs_train)

includes <- setdiff(names(CCFs_withlands), c("compare_id", "profile1_id", "profile2_id",
                                         "study.x", "study.y", "barrel.x", "barrel.y",
                                         "bullet.x", "bullet.y", "land.x", "land.y",
                                         "land_id.x", "land_id.y", "signature_length", "lag", "overlap"))

new_rtrees <- randomForest(factor(match) ~ ., data = CCFs_train[,includes], ntree = 300)
CCFs_test$forest <- predict(new_rtrees, newdata = CCFs_test, type = "prob")[,2]

CCFs_test <- CCFs_test %>%
  mutate(Type = ifelse(forest >= .5 & match, "True Positive", 
                       ifelse(forest >= .5 & !match, "False Positive",
                              ifelse(forest < .5 & match, "False Negative",
                                     "True Negative"))))

CCFs_test_normalized <- CCFs_test %>% select(Type, ccf, rough_cor, D, matches, mismatches, cms, non_cms, sum_peaks)
CCFs_test_normalized[,-1] <- lapply(CCFs_test_normalized[,-1], function(x) { (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE) })
pcp_data <- CCFs_test_normalized %>%
  group_by(Type) %>%
  summarise_each(funs(mean))
pcp_data_tall <- pcp_data %>%
  gather(key = Feature, value = Value, 2:ncol(.))
CCFs_tall <- CCFs_test_normalized %>%
  mutate(id = 1:nrow(.)) %>%
  select(id, everything()) %>%
  gather(key = Feature, value = Value, 3:ncol(.))

ggplot() +
  geom_line(data = filter(CCFs_tall, Type %in% c("True Positive", "False Positive")), aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type, size = Type)) +
  geom_line(data = filter(pcp_data_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 3) +
  ylab("Normalized Value") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("#e60000", "#000099")) +
  scale_alpha_manual(values = c(1, .05)) +
  scale_size_manual(values = c(1, 1))
```

```{r pcp2, echo=FALSE, message=FALSE, fig.cap="Parallel coordinate plot of the features based on the random forest confusion matrix for true and false negatives.", fig.height=6, fig.width=10}
mysamp <- sample(CCFs_tall$id[CCFs_tall$Type == "True Negative"], 1000)
trueneg_sample <- CCFs_tall %>%
  filter(id %in% mysamp) %>%
  rbind(CCFs_tall %>% filter(Type == "False Negative"))

ggplot() +
  #geom_line(data = filter(CCFs_tall, Type %in% c("True Negative", "False Negative")), aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type, size = Type)) +
  geom_line(data = trueneg_sample, aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type)) +
  theme_bw() +
  geom_line(data = filter(pcp_data_tall, Type %in% c("False Negative")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 3) +
  ylab("Normalized Value") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("#ff9999", "#b3b3ff")) +
  scale_alpha_manual(values = c(.5, .1))
```

## Operator Effects

We attempt to quantify the effect of the study on the matching probability by fitting a new random forest which is designed to predict the study based on the derived features. Ideally, if the assumption of independence between lands holds across different operators, this forest should have poor performance - The set of derived features should be relatively consistent among known matches and known non-matches regardless of the study since the Hamby data in both sets originated from the same gun barrels.

```{r, echo=FALSE}
includes_study <- setdiff(names(CCFs_withlands), c("compare_id", "profile1_id", "profile2_id",
                                         "barrel.x", "barrel.y", "match",
                                         "bullet.x", "bullet.y", "land.x", "land.y",
                                         "land_id.x", "land_id.y", "study.x", "study.y", "signature_length"))

result_study <- trainIndex %>% purrr::map_df(function(index) {
    CCFs_train <- CCFs_withlands[index,]
    CCFs_test = setdiff(CCFs_withlands, CCFs_train)

    CCFs_train$Study <- paste(CCFs_train$study.x, CCFs_train$study.y, sep = "_")
    CCFs_test$Study <- paste(CCFs_test$study.x, CCFs_test$study.y, sep = "_")
    CCFs_train$Study <- factor(replace(CCFs_train$Study, which(CCFs_train$Study == "Hamby44_Hamby252"), "Hamby252_Hamby44"))
    CCFs_test$Study <- factor(replace(CCFs_test$Study, which(CCFs_test$Study == "Hamby44_Hamby252"), "Hamby252_Hamby44"))

    includes_study <- setdiff(names(CCFs_train), c("compare_id", "profile1_id", "profile2_id", "barrel.x", "barrel.y", "match", "bullet.x", "bullet.y", "land.x", "land.y", "land_id.x", "land_id.y", "study.x", "study.y"))

    rtrees_study <- randomForest(Study ~ ., data = CCFs_train[,includes_study], ntree = 300)

    CCFs_test$study_forest <- predict(rtrees_study, newdata = CCFs_test)
    confmat <- as.data.frame(xtabs(~study_forest + Study, data = CCFs_test))
    names(confmat) <- c("Prediction", "Actual", "Count")

    return(confmat)
})
```

Table \ref{tab:studypred} shows the confusion matrix, with column proportions, for the random forest with study as the response. It can be seen that while overall the random forest performs poorly, as hoped, comparisons between Hamby252 bullets are more distinguishable from other comparisons.

```{r, echo=FALSE, results='asis'}
study_result <- cbind(id = rep(1:10, each = nrow(result_study) / 10), result_study)
study_result_wide <- study_result %>%
    spread(key = Actual, value = Count) %>%
    group_by(Prediction) %>%
    summarise_each(funs(mean)) %>%
    select(-id) %>%
    as.data.frame
study_result_wide_col <- study_result_wide
study_result_wide_col[,-1] <- apply(study_result_wide_col[,-1], 2, function(x) paste0(round(100 * x / sum(x), digits = 2), "%"))
names(study_result_wide_col)[1] <- "Prediction \\ Actual"

print(xtable(study_result_wide_col, digits = c(0, 0, 3, 3, 3), label = "tab:studypred", caption = "Confusion Matrix (Column Proportions) for the random forest with study as the response. It can be seen that while overall the random forest performs poorly, as hoped, comparisons between Hamby252 bullets is more distinguishable from other comparisons."), comment = FALSE, include.rownames = FALSE, table.placement = "H")
```

Figure \@ref(fig:ccfstudy) shows the distributions of the features defined above, faceted by whether the lands are known to be fired from the same gun barrel, across different study to study comparisons. The distributions among the known non-matches seem relatively consistent across study based on visual inspection. On the other hand, among known matches, Hamby252 to Hamby252 comparisons exhibit more pronounced features, including a higher average ccf, higher number of matches, and higher value of sum_peaks.

```{r ccfstudy, echo=FALSE, message=FALSE, fig.cap="Distribution of the features, facetted by match, for different study to study comparisons of lands.", fig.height=13, fig.width=10}
CCFs_features <- CCFs_withlands %>%
    mutate(Study = factor(paste(study.x, study.y, sep = "_"))) %>%
    mutate(Study = replace(Study, which(Study == "Hamby44_Hamby252"), "Hamby252_Hamby44")) %>%
    select(-matches("_id|study.x|study.y|barrel|bullet|land|signature_length|non_cms|sd_D|lag")) %>%
    select(Study, match, everything()) %>%
    mutate(match = factor(match, labels = c("Known Non-Match", "Known Match"))) %>%
    gather(key = feature, value = value, 3:ncol(.))

ggplot(CCFs_features, aes(x = Study, y = value)) +
    geom_boxplot() +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90)) +
    facet_grid(feature~match, scales = "free") +
    ggtitle("Distributions of Features by Study and Match") +
    xlab("Study") +
    ylab("Value")
```

Though visual inspection clearly shows differences, we can more formally assess the differences between distributions with a Kolmogrov-Smirnov test. Table \ref{tab:kstests} gives the results of pairwise tests, for each feature, between different set comparisons, and between known matches compared with known non-matches. Although the tests are significant, looking at the raw values of the D statistic suggest that the largest effect sizes do in fact occur in comparisons with two Hamby252 lands, as the visual inspection of the boxplots also suggested.

These results strongly suggest the need for controlling for more effects when performing the analysis. Specifically, microscope operator effects resulting in variations in scan quality and scan parameters seem to play a role in the utlimate performance of the matching algorithm. Land to land comparisons from Hamby252 consistently result in more pronounced expression of features among known matches, and therefore result in higher accuracy in the random forest. Rigorous procedures to ensure scan quality and consistency across operators need to be in place to minimize the effect of the study and ensure that the assumption of land to land independence is satisfied.

Another way to demostrate the study/operator effect is by observing the distribution of our algorithm's ideal cross section by study. Figure \@ref(fig:crosscompare) gives the distributions of the ideal cross sections by study. It can be seen that the Hamby44 ideal cross sections are more likely to be close to the base of the bullet when compared to the position of the ideal cross sections in Hamby252.

```{r crosscompare, echo=FALSE, fig.cap='Distributions of the ideal cross sections by study. It can be seen that the Hamby44 ideal cross sections are much more likely to be close to the base of the bullet compared to Hamby252.'}
result <- dbReadTable(con, "metadata_derived") %>%
    left_join(dplyr::select(all_bullets_metadata, land_id, study)) %>%
    filter(run_id == max(run_id)) %>%
    dplyr::select(land_id, ideal_crosscut, study) %>%
    filter(!is.na(ideal_crosscut), study != "Cary")

ggplot(result, aes(x = study, y = ideal_crosscut)) +
    geom_boxplot() +
    theme_bw() +
    ylim(c(0, 400)) +
    ggtitle("Distribution of Ideal Cross Section by Study")
```

Indeed, another Kolmogorov-Smirnov test confirms a significant difference in the distributions of these values ($D = 0.6239, p < 0.0001$). This result strongly suggests that the operator effect in the bullet scanning procedure must be taken into account in order to assume pairwise independence.

```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
mytest <- ks.test(x = result$ideal_crosscut[result$study == "Hamby252"], y = result$ideal_crosscut[result$study == "Hamby44"])

print(xtable(tidy(mytest), digits = c(0, 4, 4, 0, 0), label = "tab:mykstest", caption = "Results for the Kolmogrov-Smirnov distributional test between values of the ideal cross section for Hamby252 compared with Hamby44."), comment = FALSE, include.rownames = FALSE)
```

## Degraded Lands

We now turn our attention to matching degraded bullet lands, in which only fragments of the land can be recovered. Because the NIST database currently contains only full bullet lands, we artificially degrade bullets under some simplifying assumptions. Essentially, we delete portions of lands to simulate the situation where we only recover a fragment from the crime scene.  We  simulate various levels of degradation from the left, right, and middle of the land impression. We vary the proportion of the land impression that is recovered, between 100% (no degradation) and 25% (significant degradation). For example, a left-fixed 75% scenario implies that the left hand portion of the land was recovered, and the 25% rightmost portion was lost. We will do this by subsetting the signatures. Note that this is a a simplified scenario because the signatures themselves are somewhat dependent on the data that are missing because of the properties of the LOESS smoother.

```{r, cache=FALSE, echo=FALSE}
data(rtrees)

ccf <- dbReadTable(con, "ccf_new") %>% filter(compare_id %in% c(4:16, 25:30))
compares <- dbReadTable(con, "compares")
CCFs_withlands_degrade <- ccf %>%
    mutate(profile1_id_bak = profile1_id,
           profile1_id = pmin(profile1_id, profile2_id),
           profile2_id = pmax(profile2_id, profile1_id_bak)) %>%
    select(-profile1_id_bak) %>%
    left_join(dplyr::select(profiles, profile_id, land_id), by = c("profile1_id" = "profile_id")) %>%
    left_join(dplyr::select(profiles, profile_id, land_id), by = c("profile2_id" = "profile_id")) %>%
    left_join(my_matches, by = c("land_id.x" = "land1_id", "land_id.y" = "land2_id")) %>%
    left_join(dplyr::select(all_bullets_metadata, land_id, study, barrel, bullet, land), by = c("land_id.x" = "land_id")) %>%
    left_join(dplyr::select(all_bullets_metadata, land_id, study, barrel, bullet, land), by = c("land_id.y" = "land_id")) %>%
    left_join(dplyr::select(compares, compare_id, degrade_left, degrade_right), by = c("compare_id" = "compare_id")) %>%
    filter(study.x != "Cary", study.y != "Cary") %>%
    arrange(study.x, study.y) %>%
    mutate(match = as.logical(replace(match, is.na(match), 0)),
           degrade_right = 1 - degrade_right)

CCFs_withlands_degrade$forest <- predict(rtrees, newdata = CCFs_withlands_degrade, type = "prob")[,2]
CCFs_withlands_plot <- CCFs_withlands_degrade
CCFs_withlands_plot$match <- factor(CCFs_withlands_plot$match, labels = c("Known Non-Match", "Known Match"))
```

Figure \@ref(fig:senspe) gives the sensitivity (true positive rate) and specificity (true negative rate) of the random forest predictions for given levels of degradation. It can be seen that the sensitivity drops a bit until 50% of the land is available and then rises again. This occurs because the algorithm begins producing more positive predictions in general, likely as the result of the ccf being arbitrarily higher for known non-matches due to the small signature. On the other hand, the specificity drops dramatically for left, middle, and right fixed degraded lands when less than 50%. of the land impression is available for examination.  For a more in-depth exploration of the matching probabilities, Figure \@ref(fig:deghist) provides histograms of the matching probability  by  degradation level and by known match versus known non-match categories. The matching probabilities suffer when compared with the probablities obtained from comparisons between full land impressions in all cases. The jump seems to be most noticeable beginning at about 25% degradation (75% land recovered), and the algorithm struggles beyond 50%.

```{r senspe, echo=FALSE, fig.cap='Sensitivity and specificity of the random forest for given levels of degradation. It can be seen that both metrics decline as a function of the land proportion, except for the sensitivity, which rises for very low levels of the land proportion due to an increase in the amount of positive predictions.'}
result <- CCFs_withlands_degrade %>%
    mutate(fixed = ifelse(degrade_left == 0 & degrade_right == 0, "full", ifelse(degrade_left == 0, "left", ifelse(degrade_right == 0, "right", "middle"))),
           level = degrade_left + degrade_right) %>%
    dplyr::select(profile1_id, profile2_id, forest, match, fixed, level) %>%
    mutate(pred = (forest > .5)) %>%
    group_by(fixed, level) %>%
    summarise(sensitivity = mean(sensitivity(factor(!pred), factor(!match))),
              specificity = mean(specificity(factor(!pred), factor(!match)))) %>%
    gather(key = metric, value = value, 3:4) %>%
    arrange(fixed, level)
result$value[1:2] <- c(sens_result, spec_result)

## Produce plot
ggplot(result, aes(x = level, y = value, colour = fixed)) +
    facet_wrap(~metric, nrow = 2) +
    geom_point() +
    geom_line() +
    theme_bw() +
    scale_x_continuous(breaks = c(0, .125, .25, .375, .5, .625, .75), labels = c("100%", "87.5%", "75%", "62.5%", "50%", "37.5%", "25%")) +
    xlab("Land Proportion")
```

Figure \@ref(fig:featexp) gives  feature expression for known matches, as a function of the proportion of land impression recovered. It is immediately obvious that the variability in feature expression is large when only a small fraction of the land is recovered, such as 25%. For instance, `sum_peaks` and  `cms` both drop, while `D` rises. Interestingly, some of the features are better expressed for the middle-fixed case. Overall, feature expression remains relatively consistent as long as we recover 50% or more of the land impression.

```{r featexp, echo=FALSE, fig.cap='Feature expression for known matches, as a function of land proportion. It can be seen that when we fix the middle portion of the bullet land, the features tend to be better expressed.'}
feature_exp <- CCFs_withlands_degrade %>%
    filter(match) %>%
    dplyr::select(degrade_left, degrade_right, ccf, rough_cor, D, matches, mismatches, cms, non_cms, sum_peaks) %>%
    gather(key = feature, value = value, 3:ncol(.)) %>%
    group_by(degrade_left, degrade_right, feature) %>%
    summarise(value = mean(value, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(fixed = ifelse(degrade_left == 0 & degrade_right == 0, "full", ifelse(degrade_left == 0, "left", ifelse(degrade_right == 0, "right", "middle"))),
           level = degrade_left + degrade_right)

ggplot(data = feature_exp, aes(x = level, y = value, colour = fixed)) +
    geom_point() +
    geom_line() +
    theme_bw() +
    theme(legend.position = "bottom") +
    scale_x_continuous(breaks = c(0, .25, .5, .75), labels = c("100%", "75%", "50%", "25%")) +
    xlab("Land Proportion") +
    facet_wrap(~feature, scales = "free_y")
```

To come full circle, we now attempt to match a particular land which exhibits bad tank rash. Figure \@ref(fig:br924) provides an image of the surface of this land impression. Due to the tank rash, this particular land impression was originally excluded from consideration (see [@2016arXiv160105788H]). However, it appears that approximately half of the bullet land remains relatively unaffected. We extract a signature from the unaffected half and attempt to match this signature to its full known match.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{images/br9-2-4-greyflip.png}
\caption{Land 4 of Bullet 2, from Barrel 9 of Hamby Set 252. It can be seen that this particular land exhibits some major tank rash on the right half.}
\label{fig:br924}
\end{figure}

Table \ref{tab:br924pred} shows the values of the features, after extracting only the first 50% of the Hamby Barrel 9 Bullet 2, 4th land (and hence, simulating a left-fixed 50% degraded scenario), compared with a feature comparison between both full lands (and hence, including the tank rash striae). The features are derived in a comparison with its known match, the complete Bullet 1 third land fired from Barrel 9. The features, including the ccf and the matches, are expressed enough to (barely) indicate a match in the case of the degraded bullet. Using the pre-trained random forest, the predicted matching probability is 52%. This is encouraging in that attempting to match the full bullet land, by comparison, yields a matching probability of 0.0067%. This is due to the relatively higher values of the ccf, cms, and matches for the degraded comparison, and suggests that the feature standardization is working as intended.

```{r, echo=FALSE, results='asis'}
br924 <- read_x3p("images/Hamby (2009) Barrel/bullets/Br9 Bullet 2-4.x3p")

myprof <- get_crosscut(bullet = br924, x = 125)

procbul <- processBullets(br924, name = "br924", x = 125, grooves = get_grooves(myprof)$groove) %>% bulletSmooth()

br1 <- procbul %>%
    filter(!is.na(l30)) %>%
    filter(y <= quantile(y, .5)) %>%
    select(bullet, y, value, fitted, resid, se, l30) %>%
    ungroup() %>%
    mutate(bullet = 116408)
br2 <- filter(bullets_smoothed, profile_id == 116407) %>%
    dplyr::select(-run_id) %>%
    rename(bullet = profile_id) %>%
    filter(!is.na(l30))

res <- bulletGetMaxCMS(br1, br2, column = "l30", span = 25)

lofX <- res$bullets
b12 <- unique(lofX$bullet)

subLOFx1 <- subset(lofX, bullet==b12[1])
subLOFx2 <- subset(lofX, bullet==b12[2])

ys <- intersect(subLOFx1$y, subLOFx2$y)

idx1 <- which(subLOFx1$y %in% ys)
idx2 <- which(subLOFx2$y %in% ys)
distr.dist <- sqrt(mean(((subLOFx1$val[idx1] - subLOFx2$val[idx2]) * 1.5625 / 1000)^2, na.rm=TRUE))
distr.sd <- sd(subLOFx1$val * 1.5625 / 1000, na.rm=TRUE) + sd(subLOFx2$val * 1.5625 / 1000, na.rm=TRUE)
km <- which(res$lines$match)
knm <- which(!res$lines$match)
if (length(km) == 0) km <- c(length(knm)+1,0)
if (length(knm) == 0) knm <- c(length(km)+1,0)
#browser()    
# feature extraction
signature.length <- min(nrow(subLOFx1), nrow(subLOFx2))

doublesmoothed <- lofX %>%
    group_by(y) %>%
    mutate(avgl30 = mean(l30, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(smoothavgl30 = smoothloess(x = y, y = avgl30, span = .3),
           l50 = l30 - smoothavgl30)

final_doublesmoothed <- doublesmoothed %>%
    filter(y %in% ys)

rough_cor <- cor(na.omit(final_doublesmoothed$l50[final_doublesmoothed$bullet == b12[1]]), 
                  na.omit(final_doublesmoothed$l50[final_doublesmoothed$bullet == b12[2]]),
                  use = "pairwise.complete.obs")

testdf <- data.frame(ccf=res$ccf, rough_cor = rough_cor, lag=res$lag / 1000, 
           D=distr.dist, 
           sd_D = distr.sd,
           b1=b12[1], b2=b12[2],
           signature_length = signature.length * 1.5625 / 1000,
           overlap = length(ys) / signature.length,
           matches = sum(res$lines$match) * (1000 / 1.5625) / length(ys),
           mismatches = sum(!res$lines$match) * 1000 / abs(diff(range(c(subLOFx1$y, subLOFx2$y)))),
           cms = res$maxCMS * (1000 / 1.5625) / length(ys),
           cms2 = bulletr::maxCMS(subset(res$lines, type==1 | is.na(type))$match) * (1000 / 1.5625) / length(ys),
           non_cms = bulletr::maxCMS(!res$lines$match) * 1000 / abs(diff(range(c(subLOFx1$y, subLOFx2$y)))),
           left_cms = max(knm[1] - km[1], 0) * (1000 / 1.5625) / length(ys),
           right_cms = max(km[length(km)] - knm[length(knm)],0) * (1000 / 1.5625) / length(ys),
           left_noncms = max(km[1] - knm[1], 0) * 1000 / abs(diff(range(c(subLOFx1$y, subLOFx2$y)))),
           right_noncms = max(knm[length(knm)]-km[length(km)],0) * 1000 / abs(diff(range(c(subLOFx1$y, subLOFx2$y)))),
           sum_peaks = sum(abs(res$lines$heights[res$lines$match])) * (1000 / 1.5625) / length(ys)
)

mypred <- predict(rtrees, newdata = testdf, type = "prob")[,2]

br1 <- procbul %>%
    filter(!is.na(l30)) %>%
    #filter(y <= quantile(y, .5)) %>%
    select(bullet, y, value, fitted, resid, se, l30) %>%
    ungroup() %>%
    mutate(bullet = 116408)
br2 <- filter(bullets_smoothed, profile_id == 116407) %>%
    dplyr::select(-run_id) %>%
    rename(bullet = profile_id) %>%
    filter(!is.na(l30))

res <- bulletGetMaxCMS(br1, br2, column = "l30", span = 25)

lofX <- res$bullets
b12 <- unique(lofX$bullet)

subLOFx1 <- subset(lofX, bullet==b12[1])
subLOFx2 <- subset(lofX, bullet==b12[2])

ys <- intersect(subLOFx1$y, subLOFx2$y)

idx1 <- which(subLOFx1$y %in% ys)
idx2 <- which(subLOFx2$y %in% ys)
distr.dist <- sqrt(mean(((subLOFx1$val[idx1] - subLOFx2$val[idx2]) * 1.5625 / 1000)^2, na.rm=TRUE))
distr.sd <- sd(subLOFx1$val * 1.5625 / 1000, na.rm=TRUE) + sd(subLOFx2$val * 1.5625 / 1000, na.rm=TRUE)
km <- which(res$lines$match)
knm <- which(!res$lines$match)
if (length(km) == 0) km <- c(length(knm)+1,0)
if (length(knm) == 0) knm <- c(length(km)+1,0)
#browser()    
# feature extraction
signature.length <- min(nrow(subLOFx1), nrow(subLOFx2))

doublesmoothed <- lofX %>%
    group_by(y) %>%
    mutate(avgl30 = mean(l30, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(smoothavgl30 = smoothloess(x = y, y = avgl30, span = .3),
           l50 = l30 - smoothavgl30)

final_doublesmoothed <- doublesmoothed %>%
    filter(y %in% ys)

rough_cor <- cor(na.omit(final_doublesmoothed$l50[final_doublesmoothed$bullet == b12[1]]), 
                  na.omit(final_doublesmoothed$l50[final_doublesmoothed$bullet == b12[2]]),
                  use = "pairwise.complete.obs")

testdf2 <- data.frame(ccf=res$ccf, rough_cor = rough_cor, lag=res$lag / 1000, 
           D=distr.dist, 
           sd_D = distr.sd,
           b1=b12[1], b2=b12[2],
           signature_length = signature.length * 1.5625 / 1000,
           overlap = length(ys) / signature.length,
           matches = sum(res$lines$match) * (1000 / 1.5625) / length(ys),
           mismatches = sum(!res$lines$match) * 1000 / abs(diff(range(c(subLOFx1$y, subLOFx2$y)))),
           cms = res$maxCMS * (1000 / 1.5625) / length(ys),
           cms2 = bulletr::maxCMS(subset(res$lines, type==1 | is.na(type))$match) * (1000 / 1.5625) / length(ys),
           non_cms = bulletr::maxCMS(!res$lines$match) * 1000 / abs(diff(range(c(subLOFx1$y, subLOFx2$y)))),
           left_cms = max(knm[1] - km[1], 0) * (1000 / 1.5625) / length(ys),
           right_cms = max(km[length(km)] - knm[length(knm)],0) * (1000 / 1.5625) / length(ys),
           left_noncms = max(km[1] - knm[1], 0) * 1000 / abs(diff(range(c(subLOFx1$y, subLOFx2$y)))),
           right_noncms = max(knm[length(knm)]-km[length(km)],0) * 1000 / abs(diff(range(c(subLOFx1$y, subLOFx2$y)))),
           sum_peaks = sum(abs(res$lines$heights[res$lines$match])) * (1000 / 1.5625) / length(ys)
)

mypred2 <- predict(rtrees, newdata = testdf2, type = "prob")[,2]

outputdf <- testdf %>%
    select(ccf, rough_cor, D, overlap, matches, mismatches, cms, non_cms, sum_peaks) %>%
    gather(key = Feature, value = `Degraded Land`)
outputdf2 <- testdf2 %>%
    select(ccf, rough_cor, D, overlap, matches, mismatches, cms, non_cms, sum_peaks) %>%
    gather(key = Feature, value = `Full Land`)

finaloutput <- outputdf %>%
    left_join(outputdf2) %>%
    rbind(data.frame(Feature = "matchprob", `Degraded Land` = mypred, `Full Land` = mypred2, check.names = FALSE))

print(xtable(finaloutput, digits = c(0, 0, 4, 4), label = "tab:br924pred", caption = "Features extracted for a comparison of the full Hamby Barrel 9 Bullet 1 Land 3, with a left-fixed 50 percent degraded portion of Hamby Barrel 9 Bullet 2 Land 4. These two lands are known matches, and indeed the random forest does predict a match."), comment = FALSE, include.rownames = FALSE)
```

\addtocontents{toc}{\protect\newpage}

# From Lands to Bullets

Another area that deserves more study is the question of generalizing these algorithms to matching entire bullets rather than indvidual lands, as would be done in a criminal justice application. One such approach is to recognize that (at least for the Hamby bullets) there should be six matching pairs of lands for any two bullets that were fired from the same gun barrel. Therefore, for each pair of bullets, we can extract the six highest matching probabilities and average them. If we do so, we obtain a clear separation between the scores that are obtained when matching bullets known to be matches and the scores obtained from known non-matches. This is shown in Figure \@ref(fig:firstscore). No known-matches have a score below 50%, while all known non-matches have a score below 10%.

We can improve on this approach by exploiting the rotation of the bullet to compute a score. Under the assumption of land to land independence, we can define the probability that two bullets match (M) as one minus the probability that the two bullets do not match (NM). Exploiting the idea that when two bullets do not match, none of the individual lands match either, we can write the matching probability as the probability that at least one land pair in the matrix matches. Specifically,

\begin{align}
P(M) &= 1 - P(NM) \\
     &= 1 - (P(NM1) \times P(NM2) \times ... \times P(NM6)) \\
     &= 1 - ((1 - P(M1)) \times (1 - P(M2)) \times ... \times (1 - P(M6)))
\end{align}

Where $M$ is the event that two bullets match, $NM$ is the event that two bullets do not match, $M1$, $M2$, ..., $M6$ are the probabilities of land one, land two, ..., land six matching, and $NM1$, $NM2$, ..., $NM6$ are the probabilities that land one, land two, ..., land six do not match. However, to compute this probability, we need to know the alignment of the two sets of lands. Fortunately, the consistent rotation of the bullet permits this. For instance, if we knew that land 1 of bullet 1 matches land 4 of bullet 2, then we immediately know that land 2 of bullet 1 matches to land 5 of bullet 2, land 3 of bullet 1 matches to land 6 of bullet 2, etc. Hence, we can take look across six diagonals of the $6 \otimes 6$ matrix containing match probabilities. Table \ref{tab:diag} gives an example of the matrix of matching probabilities between two sets of six lands from bullets that are known matches. The matching diagonal is clear based on the high probabilities (cell $(1, 3)$, cell $(2, 4)$, cell $(3, 5)$, etc.) although it can be seen that one of the six comparisons has a relatively lower matching probability. This procedure is based on the Sequence Average Maximum (SAM) by @sensorfar in their bullet matching software application `SensoMatch`.

```{r, echo=FALSE, results='asis'}
CCFs_withlands$forest <- predict(rtrees, newdata = CCFs_withlands, type = "prob")[,2]

result <- CCFs_withlands %>% 
    filter(study.x == "Hamby252", barrel.x == 3, bullet.x == 1, study.y == "Hamby252", barrel.y == 3, bullet.y == 2) %>%
    dplyr::select(profile1_id, profile2_id, forest) %>%
    spread(key = profile2_id, value = forest)

print(xtable(result, digits = c(0, 0, 4, 4, 4, 4, 4, 4), label = "tab:diag", caption = "Matrix of matching probabilities between two sets of six lands from bullets that are known matches."), comment = FALSE, include.rownames = FALSE)
```

We derive a score by computing the bullet matching probability on each set of six matrix diagonals using the previously defined formula, under the assumption of land to land indepndence. Finally, we take the maximum score obtained out of the six results as the final matching score for a bullet pair. After doing so, we can plot the scores for known matches and known non-matches separately. Figure \@ref(fig:scores) provides the distribution of matching scores for known matches compared to known non-matches. It can be seen that the known matches all have scores of around 100%, while no non-match achieves a score of above 30%, and hence this procedure provides perfect discrimination between all pairs of bullets between and within the two Hamby datasets.

```{r scores, echo=FALSE, fig.cap='Distribution of matching scores for known matches compared to known non-matches. It can be seen that the known matches all have scores of around 100%, while no non-match achieves a score of above 30%.'}
myfunc <- function(x) {
     x <- x %>% 
            dplyr::select(profile1_id, profile2_id, forest) %>%
            spread(key = profile2_id, value = forest)
        testmat <- rbind(as.matrix(x[,-1]), as.matrix(x[,-1]))
        result <- numeric(6)
        for (i in 1:6) {
            result[i] <- 1 - prod(1 - diag(testmat[-(1:6),]), na.rm = TRUE)
            testmat <- shift.down(testmat)
        }
        return(max(result))
}

test <- CCFs_withlands %>% 
    group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
    do(score = myfunc(.)) %>%
    mutate(score = unlist(score)) %>%
    filter(study.x != study.y | barrel.x != barrel.y | bullet.x != bullet.y)
test2 <- CCFs_withlands %>%
    group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
    dplyr::summarise(match = any(match)) %>%
    left_join(test)

ggplot(data = test2, aes(x = match, y = score)) +
    geom_boxplot() +
    theme_bw()
```

On the other hand, flipping this procedure around by assuming that a match occurs if and only if all six lands match does not discriminate quite as well, as can be seen in Figure \@ref(fig:scores1). Every known bullet non-match achieves a score of about zero, but so do about 15 known bullet matches. This method performs poorly because our matching algorithm exhibits a larger false negative rate than the rate of false positives. Multiplying the probabilities together compounds the issue of false negatives and leads to some misidentification of matching bullets.

```{r scores1, echo=FALSE, fig.cap='Distribution of matching scores for known matches compared to known non-matches when assuming a match occurs if and only if all six lands match. Now, it can be seen that the known non-matches all have scores of around 0, while most though not all known matches achieve much higher scores.'}
myfunc1 <- function(x) {
     x <- x %>% 
            dplyr::select(profile1_id, profile2_id, forest) %>%
            spread(key = profile2_id, value = forest)
        testmat <- rbind(as.matrix(x[,-1]), as.matrix(x[,-1]))
        result <- numeric(6)
        for (i in 1:6) {
            result[i] <- prod(diag(testmat[-(1:6),]), na.rm = TRUE)
            testmat <- shift.down(testmat)
        }
        return(max(result))
}

test <- CCFs_withlands %>% 
    group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
    do(score = myfunc1(.)) %>%
    mutate(score = unlist(score)) %>%
    filter(study.x != study.y | barrel.x != barrel.y | bullet.x != bullet.y)
test2 <- CCFs_withlands %>%
    group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
    dplyr::summarise(match = any(match)) %>%
    left_join(test)

ggplot(data = test2, aes(x = match, y = score)) +
    geom_boxplot() +
    theme_bw()
```

If we construct a hybrid of these two approaches, we can average the probabilities along the diagonal rather than multiplying those probabilities. Doing so yields the distributions shown in Figure \@ref(fig:scores10). Now, we once again differentiate the two groups  well with no known non-match achieving a score above 10%, and no known match with a score below 40%.

```{r scores10, echo=FALSE, fig.cap='Distribution of matching scores for known matches compared to known non-matches obtained by averaging the probabilities along the maximal diagonal. Once again, it can be seen that the known non-matches all have scores of around 0, but the known matches this time are well separated, with nearly all achieving scores above .5.'}
myfunc10 <- function(x) {
     x <- x %>% 
            dplyr::select(profile1_id, profile2_id, forest) %>%
            spread(key = profile2_id, value = forest)
        testmat <- rbind(as.matrix(x[,-1]), as.matrix(x[,-1]))
        result <- numeric(6)
        for (i in 1:6) {
            result[i] <- mean(diag(testmat[-(1:6),]), na.rm = TRUE)
            testmat <- shift.down(testmat)
        }
        return(max(result))
}

test <- CCFs_withlands %>% 
    group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
    do(score = myfunc10(.)) %>%
    mutate(score = unlist(score)) %>%
    filter(study.x != study.y | barrel.x != barrel.y | bullet.x != bullet.y)
test2 <- CCFs_withlands %>%
    group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
    dplyr::summarise(match = any(match)) %>%
    left_join(test)

ggplot(data = test2, aes(x = match, y = score)) +
    geom_boxplot() +
    theme_bw()
```

One more approach to match bullets would exploit the SAM procedure on individual features. For each diagonal in the $6 \otimes 6$ matrix, we can compute an average value for each feature in our model. This  yields six sets of feature values for all six diagonals. We can then feed all six sets of features into the random forest in order to obtain a matching probability for each, taking the highest resulting probability to locate the diagonal and thus identify land to land alignment. Figure \@ref(fig:scores2) shows boxplots of the matching scores using this hybrid procedure. It can be seen that while this procedure does discriminate well, it yields some false negatives (matching bullets that our forest identifies as a non-match).

```{r scores2, echo=FALSE, fig.cap='Distribution of matching scores using a SAM procedure on the feature values for known matches compared to known non-matches.'}
myfunc2 <- function(x, feature) {
     x <- x %>% 
            select_("profile1_id", "profile2_id", feature) %>%
            spread_(key = "profile2_id", value = feature)
        testmat <- rbind(as.matrix(x[,-1]), as.matrix(x[,-1]))
        result <- numeric(6)
        for (i in 1:6) {
            result[i] <- mean(diag(testmat[-(1:6),]), na.rm = TRUE)
            testmat <- shift.down(testmat)
        }

        return(result)
}

test <- CCFs_withlands %>% 
    group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
    do(ccf = myfunc2(., feature = "ccf"),
       rough_cor = myfunc2(., feature = "rough_cor"),
       lag = myfunc2(., feature = "lag"),
       D = myfunc2(., feature = "D"),
       sd_D = myfunc2(., feature = "sd_D"),
       overlap = myfunc2(., feature = "overlap"),
       matches = myfunc2(., feature = "matches"),
       mismatches = myfunc2(., feature = "mismatches"),
       cms = myfunc2(., feature = "cms"),
       non_cms = myfunc2(., feature = "non_cms"),
       sum_peaks = myfunc2(., feature = "sum_peaks"))

hmm <- cbind(id = rep(1:nrow(test), each = 6), as.data.frame(apply(test[,7:17], 2, unlist)))

test2 <- CCFs_withlands %>%
    group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
    summarise(match = any(match)) %>%
    ungroup() %>%
    mutate(id = 1:nrow(.))

hmm$forest <- predict(rtrees, newdata = hmm, type = "prob")[,2]
final_hmm <- hmm %>%
    group_by(id) %>%
    summarise(forest = max(forest, na.rm = TRUE))

final_result <- test2 %>%
    left_join(final_hmm) %>%
    select(-id)

#final_result %>% filter(match) %>% arrange(forest) %>% head

#xtabs(~(forest > .5)+match, data = final_result)

ggplot(data = final_result, aes(x = match, y = forest)) +
    geom_boxplot() +
    theme_bw()
```

# Conclusion

In this paper, we have introduced a set of robust features that can be used to train bullet matching models. We have used these features to train a random forest and assess its out-of-sample accuracy. In doing so, we noted strong evidence of operator effects that resulted in differences in the quality of the microscope scans.

While these effects were clearly identified, the best approach to account for them in practice is less clear. In the ideal case, bullets fired from a particular gun barrel should yield surface scans that are of identical quality and properties, regardless of the operator performing the scan. To achieve this, rigorous standards may need to be put in place with regards to the alignment of the bullet under the objective, and the procedure used to scan the bullet surface. To appropriately design a set of best practices requires more research. For instance, because of the significant difference between the placement of the ideal cross section across the two studies, a best practice may specify the margin from the edge of the objective at which the bullet can be placed.

We begain exploring the robustness of the matching algorithm proposed in [@2016arXiv160105788H] to land degradation. As suspected, the algorithm performance declines as a function of the rate of degradation. However, there is a relatively clear threshold around about 50%; if 50% of the land or more is recovered, the algorithm still performs reasonably well.  When the proportion of the land that is recovered is below 50%, the accuracy with which we can compare land impressions is low.

As we have stated before, the lack of 3D images of bullets limits the extent to which these algorithms can be tested and validated.   The degraded land simulation itself may be too simplistic and not faithfully represent realistic scenarios. However, as more data are collected, we can continue to update, train and test the matching algorithm in order to improve its performance in real datasets.

\clearpage

# Appendix

```{r, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, results='asis'}
all_combinations <- combn(unique(c(as.character(CCFs_features$Study), as.character(CCFs_features$Study))), 2, simplify = FALSE)

myresult <- all_combinations %>% map_df(function(x) {
    final_result <- try({
        set1 <- filter(CCFs_features, Study == x[1])
        set2 <- filter(CCFs_features, Study == x[2])

        result <- set1 %>%
            rbind(set2) %>%
            group_by(feature) %>%
            do(matchtest = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Match"])$p.value,
               matchd = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Match"])$statistic,
               nonmatchtest = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Non-Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Non-Match"])$p.value,
               nonmatchd = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Non-Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Non-Match"])$statistic,
               set1_matchcount = length(filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Match"]),
               set1_nonmatchcount = length(filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Non-Match"]),
               set2_matchcount = length(filter(., Study == x[2])$value[filter(., Study == x[1])$match == "Known Match"]),
               set2_nonmatchcount = length(filter(., Study == x[2])$value[filter(., Study == x[1])$match == "Known Non-Match"])

            )
    })

    if (inherits(final_result, "try-error")) return(NULL)

    return(final_result %>% mutate(set1 = x[1], set2 = x[2]) %>% select(set1, set2, everything()) %>% as.data.frame)
})

final_result <- myresult %>%
    mutate_each(funs(unlist)) %>%
    mutate(set1 = gsub("Hamby", "H", set1),
           set2 = gsub("Hamby", "H", set2),
           matchtest = round(matchtest, digits = 4),
           nonmatchtest = round(nonmatchtest, digits = 4)) %>%
    as.data.frame
final_result$matchtest[final_result$matchtest < .0001] <- "< 0.0001"
final_result$nonmatchtest[final_result$nonmatchtest < .0001] <- "< 0.0001"

print(xtable(final_result[,1:7], digits = c(0, 0, 0, 0, 0, 4, 0, 4), label = "tab:kstests", caption = "Results for the Kolmogrov-Smirnov distributional test."), comment = FALSE, include.rownames = FALSE)
```

```{r deghist, echo=FALSE, fig.cap='Histograms of matching probability, facetted by the degradation level and known match versus known non-match.', fig.height=9.5, fig.width=8}
p1 <- ggplot(data = filter(CCFs_withlands_plot, degrade_left == 0), aes(x = forest)) +
    geom_histogram() +
    facet_grid(match~degrade_right, scales = "free_y") +
    ggtitle("Left Fixed") +
    theme_bw()
p2 <- ggplot(data = filter(CCFs_withlands_plot, degrade_right == 0), aes(x = forest)) +
    geom_histogram() +
    facet_grid(match~degrade_left, scales = "free_y") +
    ggtitle("Right Fixed") +
    theme_bw()
p3 <- ggplot(data = filter(CCFs_withlands_plot, (degrade_left == 0 & degrade_right == 0) | (degrade_right != 0 & degrade_left != 0)), aes(x = forest)) +
    geom_histogram() +
    facet_grid(match~degrade_left+degrade_right, scales = "free_y") +
    ggtitle("Middle Fixed") +
    theme_bw()

grid.arrange(p1, p2, p3, nrow = 3)
```

```{r firstscore, echo=FALSE, fig.cap='Score distributions for the naive approach to bullet matching, for known matches and known non-matches.'}
result <- CCFs_withlands %>%
    group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
    summarise(forest = mean(head(sort(forest, decreasing = TRUE))), match = any(match))

ggplot(data = result, aes(x = match, y = forest)) +
    geom_boxplot() +
    theme_bw()
```

### 2017-09-05 degraded-paper.tex
Move degraded paper outside of repo

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Algorithmic Approaches to Match Degraded Land Impressions},
            pdfauthor={Eric Hare, Heike Hofmann, Alicia Carriquiry},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Algorithmic Approaches to Match Degraded Land Impressions}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Eric Hare, Heike Hofmann, Alicia Carriquiry}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{9/5/2017}

\usepackage{float}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\maketitle

\textbf{Abstract}

Bullet matching is a process used to determine whether two bullets may
have been fired from the same gun barrel. Historically, this has been a
manual process performed by trained forensic examiners. Recent work
however has shown that it is possible to add statistical validity and
objectivity to the procedure. In this paper, we build upon the
algorithms explored in Automatic Matching of Bullet Lands (Hare,
Hofmann, and Carriquiry 2016) by formalizing and defining a set of
features, computed on pairs of bullet lands, which can be used in
machine learning models to assess the probability of a match. We then
use these features to perform an analysis of the two Hamby (Hamby,
Brundage, and Thorpe 2009) bullet sets (Set 252 and Set 44), to assess
the presence of microscope operator effects in scanning. We also take
some first steps to address the issue of degraded bullet lands, and
provide a range of degradation at which the matching algorithm still
performs well. Finally, we discuss generalizing land to land comparisons
to full bullet comparisons as would be used for this procedure in a
criminal justice situation.

\section{Background}\label{background}

Intense scrutiny has been focused on the process of bullet matching in
recent years (e.g., Giannelli 2011). Bullet matching, the process of
determining whether two bullets could have been fired from the same gun
barrel, has traditionally been performed without meaningful
determination of error rates or statistical assessments of uncertainty
(National Research Council 2009). There have been some attempts towards
developing mathematical and statistical approaches to bullet
matching.One such attempt was the definition of CMS, the Consecutively
Matching Striae (Biasotti 1959), with a cutoff of six to separate
matches from non-matches. Still, rigorous assessments of the
applicability of such cutoffs have not to this point been described
(Advisors on Science and Technology 2016).

Recently, several authors have addressed these well-known shortcomings.
Focusing on firing pin impressions and breech faces, Riva and Champod
(2014) have described an automated algorithm using 3D images that
enables comparison between pairs of exemplars. Other examples of work in
this and related areas include Petraco and Chan (2012), W. Chu et al.
(2011), T. Vorburger et al. (2011), and others. In our approach to this
problem, Automatic Matching of Bullet Lands, we used the Hamby 252 set
(Hamby, Brundage, and Thorpe 2009) to train and develop a random forest
in order to provide a matching probability for two bullet lands (Hare,
Hofmann, and Carriquiry 2016). While the algorithm had a very strong
performance on this set, some limitations were immediately clear. For
instance, performance was assessed only on this single set of 35 bullets
fired from a consecutively manufactured set of only ten gun barrels.
Each of these bullets was part of controlled study, and the full lands
were available for matching. While there were some data quality issues,
this was still a near ideal test case for the algorithm.

Real world applications of bullet matching often involve the recovery of
fragments of bullets from the crime scene. Traditional features used in
forensic examination work well for a full land, but there has been less
investigation into their performance in the case of a fragmented land.
For example, the CMS is naturally limited by the portion of the land
that can be recovered.

In this paper, we take steps to address these and other concerns.
Specifically, we begin by reviewing features from the literature,
computed on pairs of bullet lands, and presenting some of our own
features. We propose an approach to standardize the featues, to account
for the fact that only a portion of the land impression may be recovered
from the crime scene. With the standardized features, we tackle two
issues that were not addressed in (Hare, Hofmann, and Carriquiry 2016).
The first is the effect of the microscope operator on the resulting
images and consequent algorithm performance. The second issue has to do
with the robustness of the land matching algorithm in (Hare, Hofmann,
and Carriquiry 2016) relative to the degree of degradation of the
questioned land impression. Finally, we describe some of the initial
steps toward generalizing a matching algorithm based on land-to-land
comparisons, to one based on bullet-to-bullet comparisons, as would be
of interest in a real world application of these ideas.

\section{Feature Standardization}\label{feature-standardization}

To start, we introduce a standardized version of each of the features
used in the matching routine proposed by (Hare, Hofmann, and Carriquiry
2016). These features are computed on \emph{aligned pairs of bullet land
impressions} rather than on individual lands. This enables us, for
instance, to compute the number of matching striae. Table \ref{tab:ccf1}
and Table \ref{tab:ccf2} provide an example of six land-to-land
comparisons and the derived features for these comparisons. The two
\texttt{profile\_id} columns identify a particular land from the two
Hamby datasets, and the remaining columns are the derived features for
these comparisons.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
profile1\_id & profile2\_id & ccf & rough\_cor & lag & D & sd\_D \\ 
  \hline
49 & 540 & 0.2689 & 0.1091 & -0.1734 & 0.0026 & 0.0043 \\ 
  49 & 1032 & 0.3878 & 0.3582 & 0.1938 & 0.0017 & 0.0030 \\ 
  49 & 1540 & 0.2410 & 0.1320 & 0.2781 & 0.0023 & 0.0039 \\ 
  49 & 2044 & 0.2892 & 0.1136 & 0.0828 & 0.0023 & 0.0038 \\ 
  49 & 2583 & 0.1769 & 0.0214 & -0.0563 & 0.0026 & 0.0041 \\ 
  49 & 3074 & 0.2715 & 0.1254 & 0.0938 & 0.0020 & 0.0033 \\ 
   \hline
\end{tabular}
\caption{A sample of six land-to-land comparisons, and derived features for these comparisons.} 
\label{tab:ccf1}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
signature\_length & overlap & matches & mismatches & cms & non\_cms & sum\_peaks \\ 
  \hline
1.9328 & 0.9103 & 1.1368 & 10.2251 & 0.5684 & 3.7182 & 2.0038 \\ 
  1.9359 & 0.9193 & 5.0571 & 6.5593 & 3.3714 & 2.3426 & 6.5264 \\ 
  1.9359 & 0.8757 & 1.7696 & 10.3662 & 1.1797 & 5.8592 & 2.0192 \\ 
  1.9172 & 0.9829 & 1.5920 & 7.9441 & 0.5307 & 3.4756 & 1.8139 \\ 
  1.8234 & 0.9692 & 1.1317 & 8.8479 & 0.5659 & 4.9155 & 0.8555 \\ 
  1.9172 & 0.9690 & 2.6913 & 10.7645 & 1.0765 & 3.4251 & 3.5272 \\ 
   \hline
\end{tabular}
\caption{The remaining derived features for the previous six land-to-land comparisons.} 
\label{tab:ccf2}
\end{table}

We generalize the definitions of these features to account for the
possibility that we may be handling degraded bullet lands, where only
fragments can be recovered. The definition of each feature is given
below, where \(f(t)\) represents the height values of the first profile,
and \(g(t)\) the height values of the second:

\begin{itemize}
\tightlist
\item
  \textbf{ccf} (\%) is the maximum value of the Cross-Correlation
  function evaluated at the optimal alignment. The Cross-Correlation
  function is defined as
  \(C(\tau) = \int_{-\infty}^{\infty} f(t)g(t + \tau)dt\) where \(\tau\)
  represents the the lag of the second signature (T. Vorburger et al.
  2011).
\item
  \textbf{rough\_cor} (\%) is a new feature that quantifies the
  correlation between the two signatures after performing a second LOESS
  smoothing stage and then subtracting the result from the original
  signatures. This attempts to model the roughness of the surface after
  removing structure such as waviness.
\item
  \textbf{lag} (mm) Is the optimal lag for the ccf value.
\item
  \textbf{D} (mm) is the Euclidean vertical distance between each height
  value of the aligned signatures. This is defined as
  \(D^2 = \frac{1}{\text{\#}t}\sum_t \left[f(t) - g(t)\right]^2\). This
  is a measure of the total variation between two functions (Clarkson
  and Adams 1933).
\item
  \textbf{sd\_D} (mm) provides the standard deviation of the values of
  \emph{D} from above.
\item
  \textbf{signature\_length} (mm) is the overall length of the smallest
  of the two aligned signatures.
\item
  \textbf{overlap} (\%) provides the percentage of the two signatures
  that overlap after the alignment stage.
\item
  \textbf{matches} (per mm) is the number of matching peaks/valleys
  (striae) per millimeter of the overlapping portion of the aligned
  signatures.
\item
  \textbf{mismatches} (per mm) is the number of mismatching
  peaks/valleys (striae) per millimeter of the overlapping portion of
  the aligned signatures.
\item
  \textbf{cms} (per mm) is the number of consecutively matching
  peaks/valleys (striae) per millimeter of the overlapping portion of
  the aligned signatures (Biasotti 1959, Wei Chu et al. (2013)).
\item
  \textbf{non\_cms} (per mm) is the number of consecutive mismatching
  peaks/valleys (striae) per millimeter of the overlapping portion of
  the aligned signatures.
\item
  \textbf{sum\_peaks} (per mm) is the the sum of the average heights of
  matched striae.
\end{itemize}

The features that are expressed on a per millimeter level are intended
to support the degraded land case, as discussed earlier. Note that the
computation differs slightly depending on the feature. For example, to
standardize the number of matches, the first count the raw number of
matching striae, and then divide this number by the length of the
overlapping region of the two lands (\texttt{overlap} from above). In
most cases, the overlapping region will be very close to the length of
the smaller signature. But depending on the alignment, this may not
always be true. This ensures that we do not punish a particular
cross-comparison for having a smaller region in which matches could
occur. On the other hand, the number of mismatches is divided by the
total length of the two aligned signatures, since mismatched striae can
occur even in the non-overlapping region of the two signatures.

The \texttt{rough\_cor} or Roughness Correlation is derived by
performing a second smoothing step, and subtracting the result from the
original signatures. This creates a new signature which eliminates some
of the overall structure, allowing global deformations to have less of
an influence on the model output. Where the roughness correlation is
most useful is in a scenario like Figure \ref{fig:roughcorgood}. This
figure shows the alignment of profile 40977 with 47600. The top panel
shows the smoothed signatures. The middle panel overlays a LOESS fit to
the average of the two signatures. Finally, to derive the roughness
correlation, this LOESS is subtracted from the original signature to
create a new set of roughness residuals, which are then given in the
bottom panel. Note that these two profiles do not match, yet the ccf is
0.7724. The roughness correlation (-0.0324) correctly indicates the lack
of matching. The roughness correlation acts as a check against false
positives which can arise when there are significant deformations in the
overall structure, as in the case with both these profiles.

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/roughcorgood-1.pdf}
\caption{\label{fig:roughcorgood}Alignment of profile 40977 with 47600. The
top panel shows the smoothed signatures. The middle panel overlays a
LOESS fit to the average of the two signatures. Finally, to derive the
roughness correlation, this LOESS is subtracted from the original
signature to create a new set of roughness residuals, which are then
given in the bottom panel. Note that these two profiles do not match,
yet the ccf is 0.7724. The roughness correlation (-0.0324) correctly
indicates the lack of matching.}
\end{figure}

In a typical comparison between two profiles, such as in Figure
\ref{fig:roughcorfine}, the roughness correlation does not meaningfully
impact the matching probability given the presence of the ccf in the
model. In this figure, we see the alignment of profile 8752 with profile
136676. In this case, the waviness or the deformation pattern in the
signatures is less pronounced, and hence the resulting roughness
signature resembles the original signature more closely. These profiles
match, and both ccf (0.6891) and rough\_cor (0.7980) provide values
indicative of matching.

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/roughcorfine-1.pdf}
\caption{\label{fig:roughcorfine}Alignment of profile 8752 with profile
136676. In this case, the waviness or the deformation pattern in the
signatures is more minor, and hence the resulting roughness signature
resembles the original signature more closely. These profiles match, and
both ccf (0.6891) and rough\_cor (0.7980) provide values indicative of
matching.}
\end{figure}

We can observe the distributions of both CCF and the Roughness
Correlation side by side, differentiating between known matches and
known non-matches. Figure \ref{fig:roughcordist} shows this. It can be
seen that the separation of the two groups along both variables is quite
strong, but many known-matches have low values for each. The
distributions of known non-matches, on the other hand, is relatively
symmetric and normal.

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/roughcordist-1.pdf}
\caption{\label{fig:roughcordist}Distributions of the Roughness Correlation
compared to the CCF for known matches and known non-matches}
\end{figure}

\section{Model Training}\label{model-training}

Using these features, we can train a randomForest (Liaw and Wiener 2002)
model which attempts to predict whether two lands match given the value
of the features. There are currently three studies for bullets included
in the NIST ballistics database (REF). Those are Hamby (Set 252), Hamby
(Set 44), and Cary. For purposes of analysis we describe in this paper,
we exclude the Cary bullets from consideration, because the study was
designed to assess the persistence of striation markings over a series
of fires from the same barrel. Thus, every Cary bullet is a known match
to every other Cary bullet, although early firings do not produce the
same markings that later firings do. Hence, we will consider Hamby (Set
252) and Hamby (Set 44) only.

We can now train the forest using the features we defined earlier. We
split the data into an 80\% training, 20\% testing sets to assess the
algorithm's out of sample performance, using the \texttt{caret} package
(Jed Wing et al. 2016). Table \ref{tab:avgforest} displays the results
in the form of a confusion matrix on the test set, averaged over ten
independent random forests trained on ten random data subsets. There are
a total of 83,028 land-to-land comparisons, of which 1208 are among
known matching land impressions and 81,820 are among known non-matching
land impressions. It can be seen that false positives are exceedingly
rare, but false negatives occur more frequently (approximately 65 false
negative land to land comparisons on the test set, compared with an
average of less than four false positives).

\begin{table}[H]
\centering
\begin{tabular}{lr}
  \hline
Result & Count \\ 
  \hline
False Negative & 65.3 \\ 
  False Positive & 3.8 \\ 
  True Negative & 16363.1 \\ 
  True Positive & 171.8 \\ 
   \hline
\end{tabular}
\caption{The average confusion matrix for the 10 random forests. It can be seen that false positives are exceedingly rare, but false negatives occur more frequently.} 
\label{tab:avgforest}
\end{table}

These results suggest that our algorithm is too conservative in
predicting a match when in fact the bullets were fired from the same gun
barrel. We can break down the confusion matrix by the study from which
each of the two land impressions originated. Table
\ref{tab:avgforeststudy} shows the average confusion matrix for the 10
random forests, broken down by study. It can be seen that Hamby252 to
Hamby252 comparisons exhibit the fewest errors, while Hamby252 to
Hamby44 comparisons exhibit the most errors on average. This intuitively
makes some sense given the potential presence of scanner operator
effects, which we address further in this section.

\begin{table}[H]
\centering
\begin{tabular}{lllll}
  \hline
Study & False Negative & False Positive & True Negative & True Positive \\ 
  \hline
Hamby252\_Hamby252 & 0.29\% & 0.01\% & 98.77\% & 0.93\% \\ 
  Hamby252\_Hamby44 & 0.47\% & 0.02\% & 98.35\% & 1.16\% \\ 
  Hamby44\_Hamby44 & 0.34\% & 0.04\% & 98.72\% & 0.9\% \\ 
   \hline
\end{tabular}
\caption{The average confusion matrix for the 10 random forests, broken down by study. It can be seen that Hamby252 to Hamby252 comparisons exhibit the fewest errors, while Hamby252 to Hamby44 comparisons exhibit the most on average.} 
\label{tab:avgforeststudy}
\end{table}

\section{Feature Robustness}\label{feature-robustness}

As a first stage to assessing feature robustness, we produce parallel
coordinate plots of the various features based on true positives, true
negatives, false positives, and false negatives. Figures \ref{fig:pcp}
and \ref{fig:pcp2} display these plots. The means of the true positive
and the false negative groups are shown, respectively, in the two
figures. False positives tend to occur with anomalously high values of
\texttt{sum\_peaks}, \texttt{matches}, and \texttt{ccf}, while false
negatives tend to resemble very closely the feature distribution of the
true negatives, with a slightly higher average \texttt{ccf}.

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/pcp-1.pdf}
\caption{\label{fig:pcp}Parallel coordinate plot of the features based on
the random forest confusion matrix for true and false positives.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/pcp2-1.pdf}
\caption{\label{fig:pcp2}Parallel coordinate plot of the features based on
the random forest confusion matrix for true and false negatives.}
\end{figure}

\subsection{Operator Effects}\label{operator-effects}

We attempt to quantify the effect of the study on the matching
probability by fitting a new random forest which is designed to predict
the study based on the derived features. Ideally, if the assumption of
independence between lands holds across different operators, this forest
should have poor performance - The set of derived features should be
relatively consistent among known matches and known non-matches
regardless of the study since the Hamby data in both sets originated
from the same gun barrels.

Table \ref{tab:studypred} shows the confusion matrix, with column
proportions, for the random forest with study as the response. It can be
seen that while overall the random forest performs poorly, as hoped,
comparisons between Hamby252 bullets are more distinguishable from other
comparisons.

\begin{table}[H]
\centering
\begin{tabular}{llll}
  \hline
Prediction $\backslash$ Actual & Hamby252\_Hamby252 & Hamby252\_Hamby44 & Hamby44\_Hamby44 \\ 
  \hline
Hamby252\_Hamby252 & 28.5\% & 10.23\% & 3.24\% \\ 
  Hamby252\_Hamby44 & 70.57\% & 83.27\% & 79.78\% \\ 
  Hamby44\_Hamby44 & 0.93\% & 6.51\% & 16.98\% \\ 
   \hline
\end{tabular}
\caption{Confusion Matrix (Column Proportions) for the random forest with study as the response. It can be seen that while overall the random forest performs poorly, as hoped, comparisons between Hamby252 bullets is more distinguishable from other comparisons.} 
\label{tab:studypred}
\end{table}

Figure \ref{fig:ccfstudy} shows the distributions of the features
defined above, faceted by whether the lands are known to be fired from
the same gun barrel, across different study to study comparisons. The
distributions among the known non-matches seem relatively consistent
across study based on visual inspection. On the other hand, among known
matches, Hamby252 to Hamby252 comparisons exhibit more pronounced
features, including a higher average ccf, higher number of matches, and
higher value of sum\_peaks.

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/ccfstudy-1.pdf}
\caption{\label{fig:ccfstudy}Distribution of the features, facetted by
match, for different study to study comparisons of lands.}
\end{figure}

Though visual inspection clearly shows differences, we can more formally
assess the differences between distributions with a Kolmogrov-Smirnov
test. Table \ref{tab:kstests} gives the results of pairwise tests, for
each feature, between different set comparisons, and between known
matches compared with known non-matches. Although the tests are
significant, looking at the raw values of the D statistic suggest that
the largest effect sizes do in fact occur in comparisons with two
Hamby252 lands, as the visual inspection of the boxplots also suggested.

These results strongly suggest the need for controlling for more effects
when performing the analysis. Specifically, microscope operator effects
resulting in variations in scan quality and scan parameters seem to play
a role in the utlimate performance of the matching algorithm. Land to
land comparisons from Hamby252 consistently result in more pronounced
expression of features among known matches, and therefore result in
higher accuracy in the random forest. Rigorous procedures to ensure scan
quality and consistency across operators need to be in place to minimize
the effect of the study and ensure that the assumption of land to land
independence is satisfied.

Another way to demostrate the study/operator effect is by observing the
distribution of our algorithm's ideal cross section by study. Figure
\ref{fig:crosscompare} gives the distributions of the ideal cross
sections by study. It can be seen that the Hamby44 ideal cross sections
are more likely to be close to the base of the bullet when compared to
the position of the ideal cross sections in Hamby252.

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/crosscompare-1.pdf}
\caption{\label{fig:crosscompare}Distributions of the ideal cross sections
by study. It can be seen that the Hamby44 ideal cross sections are much
more likely to be close to the base of the bullet compared to Hamby252.}
\end{figure}

Indeed, another Kolmogorov-Smirnov test confirms a significant
difference in the distributions of these values
(\(D = 0.6239, p < 0.0001\)). This result strongly suggests that the
operator effect in the bullet scanning procedure must be taken into
account in order to assume pairwise independence.

\subsection{Degraded Lands}\label{degraded-lands}

We now turn our attention to matching degraded bullet lands, in which
only fragments of the land can be recovered. Because the NIST database
currently contains only full bullet lands, we artificially degrade
bullets under some simplifying assumptions. Essentially, we delete
portions of lands to simulate the situation where we only recover a
fragment from the crime scene. We simulate various levels of degradation
from the left, right, and middle of the land impression. We vary the
proportion of the land impression that is recovered, between 100\% (no
degradation) and 25\% (significant degradation). For example, a
left-fixed 75\% scenario implies that the left hand portion of the land
was recovered, and the 25\% rightmost portion was lost. We will do this
by subsetting the signatures. Note that this is a a simplified scenario
because the signatures themselves are somewhat dependent on the data
that are missing because of the properties of the LOESS smoother.

Figure \ref{fig:senspe} gives the sensitivity (true positive rate) and
specificity (true negative rate) of the random forest predictions for
given levels of degradation. It can be seen that the sensitivity drops a
bit until 50\% of the land is available and then rises again. This
occurs because the algorithm begins producing more positive predictions
in general, likely as the result of the ccf being arbitrarily higher for
known non-matches due to the small signature. On the other hand, the
specificity drops dramatically for left, middle, and right fixed
degraded lands when less than 50\%. of the land impression is available
for examination. For a more in-depth exploration of the matching
probabilities, Figure \ref{fig:deghist} provides histograms of the
matching probability by degradation level and by known match versus
known non-match categories. The matching probabilities suffer when
compared with the probablities obtained from comparisons between full
land impressions in all cases. The jump seems to be most noticeable
beginning at about 25\% degradation (75\% land recovered), and the
algorithm struggles beyond 50\%.

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/senspe-1.pdf}
\caption{\label{fig:senspe}Sensitivity and specificity of the random forest
for given levels of degradation. It can be seen that both metrics
decline as a function of the land proportion, except for the
sensitivity, which rises for very low levels of the land proportion due
to an increase in the amount of positive predictions.}
\end{figure}

Figure \ref{fig:featexp} gives feature expression for known matches, as
a function of the proportion of land impression recovered. It is
immediately obvious that the variability in feature expression is large
when only a small fraction of the land is recovered, such as 25\%. For
instance, \texttt{sum\_peaks} and \texttt{cms} both drop, while
\texttt{D} rises. Interestingly, some of the features are better
expressed for the middle-fixed case. Overall, feature expression remains
relatively consistent as long as we recover 50\% or more of the land
impression.

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/featexp-1.pdf}
\caption{\label{fig:featexp}Feature expression for known matches, as a
function of land proportion. It can be seen that when we fix the middle
portion of the bullet land, the features tend to be better expressed.}
\end{figure}

To come full circle, we now attempt to match a particular land which
exhibits bad tank rash. Figure \ref{fig:br924} provides an image of the
surface of this land impression. Due to the tank rash, this particular
land impression was originally excluded from consideration (see (Hare,
Hofmann, and Carriquiry 2016)). However, it appears that approximately
half of the bullet land remains relatively unaffected. We extract a
signature from the unaffected half and attempt to match this signature
to its full known match.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{images/br9-2-4-greyflip.png}
\caption{Land 4 of Bullet 2, from Barrel 9 of Hamby Set 252. It can be seen that this particular land exhibits some major tank rash on the right half.}
\label{fig:br924}
\end{figure}

Table \ref{tab:br924pred} shows the values of the features, after
extracting only the first 50\% of the Hamby Barrel 9 Bullet 2, 4th land
(and hence, simulating a left-fixed 50\% degraded scenario), compared
with a feature comparison between both full lands (and hence, including
the tank rash striae). The features are derived in a comparison with its
known match, the complete Bullet 1 third land fired from Barrel 9. The
features, including the ccf and the matches, are expressed enough to
(barely) indicate a match in the case of the degraded bullet. Using the
pre-trained random forest, the predicted matching probability is 52\%.
This is encouraging in that attempting to match the full bullet land, by
comparison, yields a matching probability of 0.0067\%. This is due to
the relatively higher values of the ccf, cms, and matches for the
degraded comparison, and suggests that the feature standardization is
working as intended.

\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
Feature & Degraded Land & Full Land \\ 
  \hline
ccf & 0.6004 & 0.4442 \\ 
  rough\_cor & 0.3671 & 0.1633 \\ 
  D & 0.0018 & 0.0023 \\ 
  overlap & 0.9968 & 0.9968 \\ 
  matches & 10.2236 & 5.6275 \\ 
  mismatches & 7.5949 & 5.0713 \\ 
  cms & 9.2013 & 4.6043 \\ 
  non\_cms & 6.5823 & 2.5357 \\ 
  sum\_peaks & 12.0020 & 6.3148 \\ 
  matchprob & 0.5200 & 0.0067 \\ 
   \hline
\end{tabular}
\caption{Features extracted for a comparison of the full Hamby Barrel 9 Bullet 1 Land 3, with a left-fixed 50 percent degraded portion of Hamby Barrel 9 Bullet 2 Land 4. These two lands are known matches, and indeed the random forest does predict a match.} 
\label{tab:br924pred}
\end{table}

\addtocontents{toc}{\protect\newpage}

\section{From Lands to Bullets}\label{from-lands-to-bullets}

Another area that deserves more study is the question of generalizing
these algorithms to matching entire bullets rather than indvidual lands,
as would be done in a criminal justice application. One such approach is
to recognize that (at least for the Hamby bullets) there should be six
matching pairs of lands for any two bullets that were fired from the
same gun barrel. Therefore, for each pair of bullets, we can extract the
six highest matching probabilities and average them. If we do so, we
obtain a clear separation between the scores that are obtained when
matching bullets known to be matches and the scores obtained from known
non-matches. This is shown in Figure \ref{fig:firstscore}. No
known-matches have a score below 50\%, while all known non-matches have
a score below 10\%.

We can improve on this approach by exploiting the rotation of the bullet
to compute a score. Under the assumption of land to land independence,
we can define the probability that two bullets match (M) as one minus
the probability that the two bullets do not match (NM). Exploiting the
idea that when two bullets do not match, none of the individual lands
match either, we can write the matching probability as the probability
that at least one land pair in the matrix matches. Specifically,

\begin{align}
P(M) &= 1 - P(NM) \\
     &= 1 - (P(NM1) \times P(NM2) \times ... \times P(NM6)) \\
     &= 1 - ((1 - P(M1)) \times (1 - P(M2)) \times ... \times (1 - P(M6)))
\end{align}

Where \(M\) is the event that two bullets match, \(NM\) is the event
that two bullets do not match, \(M1\), \(M2\), \ldots{}, \(M6\) are the
probabilities of land one, land two, \ldots{}, land six matching, and
\(NM1\), \(NM2\), \ldots{}, \(NM6\) are the probabilities that land one,
land two, \ldots{}, land six do not match. However, to compute this
probability, we need to know the alignment of the two sets of lands.
Fortunately, the consistent rotation of the bullet permits this. For
instance, if we knew that land 1 of bullet 1 matches land 4 of bullet 2,
then we immediately know that land 2 of bullet 1 matches to land 5 of
bullet 2, land 3 of bullet 1 matches to land 6 of bullet 2, etc. Hence,
we can take look across six diagonals of the \(6 \otimes 6\) matrix
containing match probabilities. Table \ref{tab:diag} gives an example of
the matrix of matching probabilities between two sets of six lands from
bullets that are known matches. The matching diagonal is clear based on
the high probabilities (cell \((1, 3)\), cell \((2, 4)\), cell
\((3, 5)\), etc.) although it can be seen that one of the six
comparisons has a relatively lower matching probability. This procedure
is based on the Sequence Average Maximum (SAM) by Sensorfar (2017) in
their bullet matching software application \texttt{SensoMatch}.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
profile1\_id & 45604 & 46104 & 46601 & 47069 & 47600 & 48069 \\ 
  \hline
42594 & 0.0000 & 0.0000 & 1.0000 & 0.0000 & 0.0000 & 0.0000 \\ 
  43063 & 0.0000 & 0.0000 & 0.0000 & 1.0000 & 0.0067 & 0.0000 \\ 
  43581 & 0.0000 & 0.0000 & 0.0000 & 0.0000 & 0.8433 & 0.0000 \\ 
  44211 & 0.0000 & 0.0000 & 0.0133 & 0.0000 & 0.0000 & 0.6700 \\ 
  44568 & 1.0000 & 0.0000 & 0.0033 & 0.0000 & 0.0000 & 0.0000 \\ 
  45070 & 0.0000 & 1.0000 & 0.0000 & 0.0000 & 0.0000 & 0.0000 \\ 
   \hline
\end{tabular}
\caption{Matrix of matching probabilities between two sets of six lands from bullets that are known matches.} 
\label{tab:diag}
\end{table}

We derive a score by computing the bullet matching probability on each
set of six matrix diagonals using the previously defined formula, under
the assumption of land to land indepndence. Finally, we take the maximum
score obtained out of the six results as the final matching score for a
bullet pair. After doing so, we can plot the scores for known matches
and known non-matches separately. Figure \ref{fig:scores} provides the
distribution of matching scores for known matches compared to known
non-matches. It can be seen that the known matches all have scores of
around 100\%, while no non-match achieves a score of above 30\%, and
hence this procedure provides perfect discrimination between all pairs
of bullets between and within the two Hamby datasets.

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/scores-1.pdf}
\caption{\label{fig:scores}Distribution of matching scores for known matches
compared to known non-matches. It can be seen that the known matches all
have scores of around 100\%, while no non-match achieves a score of
above 30\%.}
\end{figure}

On the other hand, flipping this procedure around by assuming that a
match occurs if and only if all six lands match does not discriminate
quite as well, as can be seen in Figure \ref{fig:scores1}. Every known
bullet non-match achieves a score of about zero, but so do about 15
known bullet matches. This method performs poorly because our matching
algorithm exhibits a larger false negative rate than the rate of false
positives. Multiplying the probabilities together compounds the issue of
false negatives and leads to some misidentification of matching bullets.

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/scores1-1.pdf}
\caption{\label{fig:scores1}Distribution of matching scores for known
matches compared to known non-matches when assuming a match occurs if
and only if all six lands match. Now, it can be seen that the known
non-matches all have scores of around 0, while most though not all known
matches achieve much higher scores.}
\end{figure}

If we construct a hybrid of these two approaches, we can average the
probabilities along the diagonal rather than multiplying those
probabilities. Doing so yields the distributions shown in Figure
\ref{fig:scores10}. Now, we once again differentiate the two groups well
with no known non-match achieving a score above 10\%, and no known match
with a score below 40\%.

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/scores10-1.pdf}
\caption{\label{fig:scores10}Distribution of matching scores for known
matches compared to known non-matches obtained by averaging the
probabilities along the maximal diagonal. Once again, it can be seen
that the known non-matches all have scores of around 0, but the known
matches this time are well separated, with nearly all achieving scores
above .5.}
\end{figure}

One more approach to match bullets would exploit the SAM procedure on
individual features. For each diagonal in the \(6 \otimes 6\) matrix, we
can compute an average value for each feature in our model. This yields
six sets of feature values for all six diagonals. We can then feed all
six sets of features into the random forest in order to obtain a
matching probability for each, taking the highest resulting probability
to locate the diagonal and thus identify land to land alignment. Figure
\ref{fig:scores2} shows boxplots of the matching scores using this
hybrid procedure. It can be seen that while this procedure does
discriminate well, it yields some false negatives (matching bullets that
our forest identifies as a non-match).

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/scores2-1.pdf}
\caption{\label{fig:scores2}Distribution of matching scores using a SAM
procedure on the feature values for known matches compared to known
non-matches.}
\end{figure}

\section{Conclusion}\label{conclusion}

In this paper, we have introduced a set of robust features that can be
used to train bullet matching models. We have used these features to
train a random forest and assess its out-of-sample accuracy. In doing
so, we noted strong evidence of operator effects that resulted in
differences in the quality of the microscope scans.

While these effects were clearly identified, the best approach to
account for them in practice is less clear. In the ideal case, bullets
fired from a particular gun barrel should yield surface scans that are
of identical quality and properties, regardless of the operator
performing the scan. To achieve this, rigorous standards may need to be
put in place with regards to the alignment of the bullet under the
objective, and the procedure used to scan the bullet surface. To
appropriately design a set of best practices requires more research. For
instance, because of the significant difference between the placement of
the ideal cross section across the two studies, a best practice may
specify the margin from the edge of the objective at which the bullet
can be placed.

We begain exploring the robustness of the matching algorithm proposed in
(Hare, Hofmann, and Carriquiry 2016) to land degradation. As suspected,
the algorithm performance declines as a function of the rate of
degradation. However, there is a relatively clear threshold around about
50\%; if 50\% of the land or more is recovered, the algorithm still
performs reasonably well. When the proportion of the land that is
recovered is below 50\%, the accuracy with which we can compare land
impressions is low.

As we have stated before, the lack of 3D images of bullets limits the
extent to which these algorithms can be tested and validated. The
degraded land simulation itself may be too simplistic and not faithfully
represent realistic scenarios. However, as more data are collected, we
can continue to update, train and test the matching algorithm in order
to improve its performance in real datasets.

\clearpage

\section{Appendix}\label{appendix}

\begin{table}[ht]
\centering
\begin{tabular}{llllrlr}
  \hline
set1 & set2 & feature & matchtest & matchd & nonmatchtest & nonmatchd \\ 
  \hline
H252\_H252 & H252\_H44 & ccf & $<$ 0.0001 & 0.2723 & 1e-04 & 0.0189 \\ 
  H252\_H252 & H252\_H44 & cms & $<$ 0.0001 & 0.1751 & $<$ 0.0001 & 0.0245 \\ 
  H252\_H252 & H252\_H44 & D & $<$ 0.0001 & 0.2567 & $<$ 0.0001 & 0.1049 \\ 
  H252\_H252 & H252\_H44 & matches & $<$ 0.0001 & 0.1933 & $<$ 0.0001 & 0.0327 \\ 
  H252\_H252 & H252\_H44 & mismatches & $<$ 0.0001 & 0.2015 & 0.3537 & 0.0079 \\ 
  H252\_H252 & H252\_H44 & overlap & 0.0492 & 0.0984 & $<$ 0.0001 & 0.0276 \\ 
  H252\_H252 & H252\_H44 & rough\_cor & $<$ 0.0001 & 0.2647 & $<$ 0.0001 & 0.0970 \\ 
  H252\_H252 & H252\_H44 & sum\_peaks & 8e-04 & 0.1426 & 0.0015 & 0.0162 \\ 
  H252\_H252 & H44\_H44 & ccf & $<$ 0.0001 & 0.2160 & $<$ 0.0001 & 0.0257 \\ 
  H252\_H252 & H44\_H44 & cms & $<$ 0.0001 & 0.2515 & $<$ 0.0001 & 0.0467 \\ 
  H252\_H252 & H44\_H44 & D & $<$ 0.0001 & 0.2342 & $<$ 0.0001 & 0.1946 \\ 
  H252\_H252 & H44\_H44 & matches & $<$ 0.0001 & 0.2770 & $<$ 0.0001 & 0.0713 \\ 
  H252\_H252 & H44\_H44 & mismatches & $<$ 0.0001 & 0.2505 & 0.0414 & 0.0138 \\ 
  H252\_H252 & H44\_H44 & overlap & 0.2432 & 0.0906 & $<$ 0.0001 & 0.0408 \\ 
  H252\_H252 & H44\_H44 & rough\_cor & $<$ 0.0001 & 0.2242 & $<$ 0.0001 & 0.1718 \\ 
  H252\_H252 & H44\_H44 & sum\_peaks & 1e-04 & 0.1926 & $<$ 0.0001 & 0.0289 \\ 
  H252\_H44 & H44\_H44 & ccf & 0.1149 & 0.0883 & $<$ 0.0001 & 0.0259 \\ 
  H252\_H44 & H44\_H44 & cms & 0.111 & 0.0888 & $<$ 0.0001 & 0.0262 \\ 
  H252\_H44 & H44\_H44 & D & 0.2923 & 0.0724 & $<$ 0.0001 & 0.0906 \\ 
  H252\_H44 & H44\_H44 & matches & 0.0603 & 0.0977 & $<$ 0.0001 & 0.0423 \\ 
  H252\_H44 & H44\_H44 & mismatches & 0.3301 & 0.0700 & 0.1633 & 0.0096 \\ 
  H252\_H44 & H44\_H44 & overlap & 0.8231 & 0.0465 & 1e-04 & 0.0190 \\ 
  H252\_H44 & H44\_H44 & rough\_cor & 0.2671 & 0.0741 & $<$ 0.0001 & 0.0769 \\ 
  H252\_H44 & H44\_H44 & sum\_peaks & 0.047 & 0.1011 & 0.006 & 0.0147 \\ 
   \hline
\end{tabular}
\caption{Results for the Kolmogrov-Smirnov distributional test.} 
\label{tab:kstests}
\end{table}

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/deghist-1.pdf}
\caption{\label{fig:deghist}Histograms of matching probability, facetted by
the degradation level and known match versus known non-match.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{degraded-paper_files/figure-latex/firstscore-1.pdf}
\caption{\label{fig:firstscore}Score distributions for the naive approach to
bullet matching, for known matches and known non-matches.}
\end{figure}

\hypertarget{refs}{}
\hypertarget{ref-pcast2016}{}
Advisors on Science, President's Council of, and Technology. 2016.
``Report on Forensic Science in Criminal Courts: Ensuring Scientific
Validity of Feature-Comparison Methods.''
\url{https://www.whitehouse.gov/sites/default/files/microsites/ostp/PCAST/pcast_forensic_science_report_final.pdf}.

\hypertarget{ref-biasotti:1959}{}
Biasotti, Alfred A. 1959. ``A Statistical Study of the Individual
Characteristics of Fired Bullets.'' \emph{Journal of Forensic Sciences}
4 (1): 34--50.

\hypertarget{ref-chu:2011}{}
Chu, W., J. Song, T. Vorburger, R. Thompson, and R. Silver. 2011.
``Selecting Valid Correlation Areas for Automated Bullet Identification
System Based on Striation Detection.'' \emph{Journal of Research of the
National Institute of Standards and Technology} 116 (3): 649.

\hypertarget{ref-thompson:2013}{}
Chu, Wei, Robert M Thompson, John Song, and Theodore V Vorburger. 2013.
``Automatic identification of bullet signatures based on consecutive
matching striae (CMS) criteria.'' \emph{Forensic Science International}
231 (1--3): 137--41.

\hypertarget{ref-clarkson1933definitions}{}
Clarkson, James A, and C Raymond Adams. 1933. ``On Definitions of
Bounded Variation for Functions of Two Variables.'' \emph{Transactions
of the American Mathematical Society} 35 (4). JSTOR: 824--54.

\hypertarget{ref-giannelli:2011}{}
Giannelli, Paul C. 2011. ``Ballistics Evidence Under Fire.''
\emph{Criminal Justice} 25 (4): 50--51.

\hypertarget{ref-hamby:2009}{}
Hamby, James E., David J. Brundage, and James W. Thorpe. 2009. ``The
Identification of Bullets Fired from 10 Consecutively Rifled 9mm Ruger
Pistol Barrels: A Research Project Involving 507 Participants from 20
Countries.'' \emph{AFTE Journal} 41 (2): 99--110.

\hypertarget{ref-2016arXiv160105788H}{}
Hare, E., H. Hofmann, and A. Carriquiry. 2016. ``Automatic Matching of
Bullet Lands.'' \emph{ArXiv E-Prints}, January.

\hypertarget{ref-caretpkg}{}
Jed Wing, Max Kuhn. Contributions from, Steve Weston, Andre Williams,
Chris Keefer, Allan Engelhardt, Tony Cooper, Zachary Mayer, et al. 2016.
\emph{Caret: Classification and Regression Training}.
\url{https://CRAN.R-project.org/package=caret}.

\hypertarget{ref-randomForest}{}
Liaw, Andy, and Matthew Wiener. 2002. ``Classification and Regression by
RandomForest.'' \emph{R News} 2 (3): 18--22.
\url{http://CRAN.R-project.org/doc/Rnews/}.

\hypertarget{ref-NAS:2009}{}
National Research Council. 2009. \emph{Strengthening Forensic Science in
the United States: A Path Forward}. Washington, DC: The National
Academies Press.
doi:\href{https://doi.org/10.17226/12589}{10.17226/12589}.

\hypertarget{ref-petraco:2012}{}
Petraco, Nicholas, and Helen Chan. 2012. \emph{Application of Machine
Learning to Toolmarks: Statistically Based Methods for Impression
Pattern Comparisons}. Mannheim, Germany: Bibliographisches Institut AG.

\hypertarget{ref-riva:2014}{}
Riva, Fabiano, and Christophe Champod. 2014. ``Automatic Comparison and
Evaluation of Impressions Left by a Firearm on Fired Cartridge Cases.''
\emph{Journal of Forensic Sciences} 59 (3): 637--47.
doi:\href{https://doi.org/10.1111/1556-4029.12382}{10.1111/1556-4029.12382}.

\hypertarget{ref-sensorfar}{}
Sensorfar. 2017. \emph{SensoMATCH Bullet Comparison Software}.

\hypertarget{ref-vorburger:2011}{}
Vorburger, T.V., J.-F. Song, W. Chu, L. Ma, S.H. Bui, A. Zheng, and T.B.
Renegar. 2011. ``Applications of Cross-Correlation Functions.''
\emph{Wear} 271 (3--4): 529--33.
doi:\href{https://doi.org/http://dx.doi.org/10.1016/j.wear.2010.03.030}{http://dx.doi.org/10.1016/j.wear.2010.03.030}.

\end{document}

### 2017-09-06 degraded-paper.Rmd
Add in feature expression for non-matches

@@
 Overall, feature expression remains relatively consistent as long as we recover 50% or more of the land impression.
+Feature \@ref(fig:featexp2) shows the feature expression for known non-matches by comparison.
 
@@
 ggplot(data = feature_exp, aes(x = level, y = value, colour = fixed)) +
+    geom_point() +
+    geom_line() +
+    theme_bw() +
+    theme(legend.position = "bottom") +
+    scale_x_continuous(breaks = c(0, .25, .5, .75), labels = c("100%", "75%", "50%", "25%")) +
+    xlab("Land Proportion") +
+    facet_wrap(~feature, scales = "free_y")
+```
+
+```{r featexp2, echo=FALSE, fig.cap='Feature expression for known non-matches, as a function of land proportion. Patterns of feature expression as a function are far less visible in this case.'}
+feature_exp2 <- CCFs_withlands_degrade %>%
+    filter(!match) %>%
+    dplyr::select(degrade_left, degrade_right, ccf, rough_cor, D, matches, mismatches, cms, non_cms, sum_peaks) %>%
+    gather(key = feature, value = value, 3:ncol(.)) %>%
+    group_by(degrade_left, degrade_right, feature) %>%
+    summarise(value = mean(value, na.rm = TRUE)) %>%
+    ungroup() %>%
+    mutate(fixed = ifelse(degrade_left == 0 & degrade_right == 0, "full", ifelse(degrade_left == 0, "left", ifelse(degrade_right == 0, "right", "middle"))),
+           level = degrade_left + degrade_right)
+
+ggplot(data = feature_exp2, aes(x = level, y = value, colour = fixed)) +
     geom_point() +

### 2017-09-06 degraded-paper.tex
Add in feature expression for non-matches

@@
 relatively consistent as long as we recover 50\% or more of the land
-impression.
+impression. Feature \ref{fig:featexp2} shows the feature expression for
+known non-matches by comparison.
 
@@
 portion of the bullet land, the features tend to be better expressed.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{degraded-paper_files/figure-latex/featexp2-1.pdf}
+\caption{\label{fig:featexp2}Feature expression for known non-matches, as a
+function of land proportion. Patterns of feature expression as a
+function are far less visible in this case.}
 \end{figure}

### 2017-09-11 degraded-paper.Rmd
Fix the references section

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
 ```
@@
 
-# Appendix
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, results='asis'}
+```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, fig.pos='H', results='asis'}
 all_combinations <- combn(unique(c(as.character(CCFs_features$Study), as.character(CCFs_features$Study))), 2, simplify = FALSE)
@@
 
-```{r deghist, echo=FALSE, fig.cap='Histograms of matching probability, facetted by the degradation level and known match versus known non-match.', fig.height=9.5, fig.width=8}
+```{r eval=FALSE, deghist, echo=FALSE, fig.cap='Histograms of matching probability, facetted by the degradation level and known match versus known non-match.', fig.height=9.5, fig.width=8, fig.pos='H'}
 p1 <- ggplot(data = filter(CCFs_withlands_plot, degrade_left == 0), aes(x = forest)) +
@@
 
-```{r firstscore, echo=FALSE, fig.cap='Score distributions for the naive approach to bullet matching, for known matches and known non-matches.'}
+```{r eval=FALSE, firstscore, echo=FALSE, fig.cap='Score distributions for the naive approach to bullet matching, for known matches and known non-matches.', fig.pos='H'}
 result <- CCFs_withlands %>%
@@
 ```
+
+# References

### 2017-09-11 degraded-paper.tex
Fix the references section

@@
 
-\section{Appendix}\label{appendix}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{llllrlr}
-  \hline
-set1 & set2 & feature & matchtest & matchd & nonmatchtest & nonmatchd \\ 
-  \hline
-H252\_H252 & H252\_H44 & ccf & $<$ 0.0001 & 0.2723 & 1e-04 & 0.0189 \\ 
-  H252\_H252 & H252\_H44 & cms & $<$ 0.0001 & 0.1751 & $<$ 0.0001 & 0.0245 \\ 
-  H252\_H252 & H252\_H44 & D & $<$ 0.0001 & 0.2567 & $<$ 0.0001 & 0.1049 \\ 
-  H252\_H252 & H252\_H44 & matches & $<$ 0.0001 & 0.1933 & $<$ 0.0001 & 0.0327 \\ 
-  H252\_H252 & H252\_H44 & mismatches & $<$ 0.0001 & 0.2015 & 0.3537 & 0.0079 \\ 
-  H252\_H252 & H252\_H44 & overlap & 0.0492 & 0.0984 & $<$ 0.0001 & 0.0276 \\ 
-  H252\_H252 & H252\_H44 & rough\_cor & $<$ 0.0001 & 0.2647 & $<$ 0.0001 & 0.0970 \\ 
-  H252\_H252 & H252\_H44 & sum\_peaks & 8e-04 & 0.1426 & 0.0015 & 0.0162 \\ 
-  H252\_H252 & H44\_H44 & ccf & $<$ 0.0001 & 0.2160 & $<$ 0.0001 & 0.0257 \\ 
-  H252\_H252 & H44\_H44 & cms & $<$ 0.0001 & 0.2515 & $<$ 0.0001 & 0.0467 \\ 
-  H252\_H252 & H44\_H44 & D & $<$ 0.0001 & 0.2342 & $<$ 0.0001 & 0.1946 \\ 
-  H252\_H252 & H44\_H44 & matches & $<$ 0.0001 & 0.2770 & $<$ 0.0001 & 0.0713 \\ 
-  H252\_H252 & H44\_H44 & mismatches & $<$ 0.0001 & 0.2505 & 0.0414 & 0.0138 \\ 
-  H252\_H252 & H44\_H44 & overlap & 0.2432 & 0.0906 & $<$ 0.0001 & 0.0408 \\ 
-  H252\_H252 & H44\_H44 & rough\_cor & $<$ 0.0001 & 0.2242 & $<$ 0.0001 & 0.1718 \\ 
-  H252\_H252 & H44\_H44 & sum\_peaks & 1e-04 & 0.1926 & $<$ 0.0001 & 0.0289 \\ 
-  H252\_H44 & H44\_H44 & ccf & 0.1149 & 0.0883 & $<$ 0.0001 & 0.0259 \\ 
-  H252\_H44 & H44\_H44 & cms & 0.111 & 0.0888 & $<$ 0.0001 & 0.0262 \\ 
-  H252\_H44 & H44\_H44 & D & 0.2923 & 0.0724 & $<$ 0.0001 & 0.0906 \\ 
-  H252\_H44 & H44\_H44 & matches & 0.0603 & 0.0977 & $<$ 0.0001 & 0.0423 \\ 
-  H252\_H44 & H44\_H44 & mismatches & 0.3301 & 0.0700 & 0.1633 & 0.0096 \\ 
-  H252\_H44 & H44\_H44 & overlap & 0.8231 & 0.0465 & 1e-04 & 0.0190 \\ 
-  H252\_H44 & H44\_H44 & rough\_cor & 0.2671 & 0.0741 & $<$ 0.0001 & 0.0769 \\ 
-  H252\_H44 & H44\_H44 & sum\_peaks & 0.047 & 0.1011 & 0.006 & 0.0147 \\ 
-   \hline
-\end{tabular}
-\caption{Results for the Kolmogrov-Smirnov distributional test.} 
-\label{tab:kstests}
-\end{table}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{degraded-paper_files/figure-latex/deghist-1.pdf}
-\caption{\label{fig:deghist}Histograms of matching probability, facetted by
-the degradation level and known match versus known non-match.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{degraded-paper_files/figure-latex/firstscore-1.pdf}
-\caption{\label{fig:firstscore}Score distributions for the naive approach to
-bullet matching, for known matches and known non-matches.}
-\end{figure}
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
 

### 2017-10-03 response.Rmd
Paper updates

---
title: "Response to reviewers"
author: "Eric Hare, Heike Hofmann, Alicia Carriquiry"
date: "October 5th, 2017"
output:
  pdf_document: default
bibliography: references.bib
---

# Reviewer 1

1. *pp. 4  first paragraph of section “3 Model Training”, the author state: "Thus, every Cary bullet is a known match to every other Cary bullet, although early firings do not produce the same markings that later firings do." The authors did not study this here or elsewhere. The authors also did not cite a study that indicated this conclusion. The second half of that statement may very well be true, but  stating it anecdotally should not appear in a research paper.  I’d suggest the minor change: "Thus every Cary bullet is a known match to every other Cary bullet"*

  **Thank you for the comment. We have made this change.**

2. *pp8-9, 11, 25: The authors highlight operator effect on the data acquisition process in the study. The instrumentation used to acquire data requires skill and experience. Naturally that brings up the question as to who was operating the microscope? The authors must state (in at least rough terms) the level of expertise of the operator(s). Were they students new to the technology? Were the operators familiar with the quirks of the surfaces they were looking at? Were they professional surface metrologists or company technicians? I would expect significant operator effects if data is collected by operators inexperienced  with the instrumentation, surface metrology and/or firearms analysis.*

  **In this case, the operators were experienced surface metrologists at NIST. The presence of operator effects was not a conclusion meant to insinuate that these experts performed the scans poorly, but rather that there is variability in the procedures and standards used. The paper has been updated to include this information.**

3. *Page 11,  last paragraph the authors state: "This result strongly suggests that the operator effect in the bullet scanning procedure must be taken into account in order to assume pairwise independence." I don’t understand this statement at all. Do the authors mean independence between features? The pair-wise comparisons between the feature vectors in this study obviously ARE NOT independent (and the features are likely not independent either…). I’m guessing I'm not understanding something; or the author's logic. I suggest the authors expand this paragraph a bit and clarify the statement.*

  **Apologies. The intent here in terms of independence was pairwise independence of scans between the two studies. Because Hamby 252 and Hamby 44 were scans collected on bullets fired from the same gun barrels, assuming that microscope operation and the scanning procedure is precisely defined and perfect, there should be independence between the scans conducted by different operators, and therefore the features derived from these scans should be the same, independent of who conducted the scan. We have clarified this in the manuscript.**

4. *pp14, last paragraph which discusses tank rash on a surface. The authors state: "We extract a signature from the unaffected half and attempt to match this signature to its full known match." Just out of curiosity, how did the authors segment out tank rash? Automatically or subjectively by hand? Could they please state explicitly."*

  **This was determined subjectively by visual inspection, and we have clarified that point in the paper.**

5. *pp18 section 5 “From Lands to Bullets”. A very similar model was already presented by Chu et al in 2010: Chu et al. J Forensic Sci, March 2010, Vol. 55, No. 2.  pp. 341-347. It should be cited and deserves at least a sentence of discussion/comparison...*

  **Thank you. We have added this citation and added some discussion of the differences compared to our approach.**

# Reviewer 2

1. *In this case, I doubt the usability of their method in other real cases or other set of barrels. So saying the features are robust will be only for this specific barrel unless proven in other cases (datasets)*

  **We do agree that much more work needs to be done to assess the performance of the algorithm in real cases, and in the case of other barrels. At the time of writing, further data was not publicly available to us, and that is why we elected to assess what effects were available (operator or study effects in the bullet scanning process), or to simulate conditions that may be more indicative of a real-world scenario (degraded land impressions). The features used were derived and standardized such that we believe they will be applicable to out-of-sample real world data as we continue to obtain it. Indeed, we've seen some promising results with bullet sets sent to us by police departments since the time of writing.**

2. *The partitioning procedure of the data to 80%-20% is unclear. It seems the authors partitioning after getting the all pairwise land-to-land comparisons. This will skew the error rate and won’t represent the true error rate if new barrel/bullet is applied to this algorithm. Instead, try to keep out one bullet from each barrel/ or keep out a whole barrel to assess the error rate correctly.*

  **Thank you for the feedback. We have redone the partitioning scheme and clarified the method in the paper. As you suggested, we have elected to hold out 20% of the barrels, and then used all land-to-land comparisons between lands in these barrels. The results match very closely to the previous scheme, with a very high specificity and a relatively lower sensitivity.**

3. *The overlayed density plots in Figure 3 needs to have better description. Are you looking at all pairwise known non match scores in the figure? What is the point you are trying to make — symmetric and normally distributed vs skewed ? Does the two features differ? in what way? Is one better than the other? In addition, what bandwidth was used. Did you use the default? In that case, it will depend on the number of matches and non matches. I suspect if you use the same bandwidth, the overlap will be different. Please use Empirical - CDF plots instead to avoid the issue of bandwidth selection for the two groups and will give better view of the plots and corresponding error rates.*

  **We have updated these plots to use empirical CDF plots, as per your advice about the bandwidth issue. Thanks for pointing this out. We have also clarified the descriptions of these plots to indicate the relevant information, including the symmetry of each distribution and the differences across the features.**

4. *Try to summarize Figures 12-15 in one figure- side by side boxplots should work fine.*

  **These figures have been updated. Thank you.**

5. *The latter part of the paper is hard to follow. It appears that some of the figures and tables are missing from the paper (See pages 11 and 18). If this is to be published it needs a significant rewrite and reorganization from the current form.*

  **We have updated this section to include the missing figures and tables, and reorganized them as appropriate. We hope you agree the new structure of this section is much easier to follow.**

6. *P 2 paragraph 2 “Real world applications ... “ the statements made here need citations*

  **Citations have been added for these statements.**

7. *P 3 features description: indicate which ones are new features vs already proposed ones*

  **This has now been included.**

8. *P 8 describe data partitioning method more clearly. Make sure the test set is correlated with the validation set.*

  **Thank you, we have clarified the data partitioning used and revised it based on your advice.**

9. *P 8 “It can be seen that Hamby252 to Hamby252 comparisons exhibit the fewest errors, while Hamby252 to Hamby44 comparisons exhibit the most on average” — the ideal would be to have equal error rates in all cases - data partitioning method may solve this issue*

  **The new data partitioning scheme has improved the results, however hasn't completely eliminated this effect. The discussion has been updated to reflect the new results.**

10. *P 9 “... resemble very closely the feature distribution of the true negatives, with a slightly higher average ccf.” add sentence describing what that section is going to be about — then discuss the plot and rest of the subsection*

  **We have added this clarifying sentence.**

11. *P 9 Figure 4 — make better choice of colors and line types for black and white version of the paper. Same with Figure 5*

  **We have updated the figures accordingly, combining them into one figure in the process.**

12. *P 11 missing table*

  **Apologies. This has been corrected.**

12. *P 18 missing figure*

  **This has also been corrected.**

13. *P19 avoid equation number on each line of the equation - this is a straight forward calculation and do not need to be explained line by line - “Where” after the equation should be “where”*

  **Updated. Thank you.**

### 2017-10-03 degraded-paper.Rmd
Paper updates

@@
 author: "Eric Hare, Heike Hofmann, Alicia Carriquiry"
-date: "9/5/2017"
+date: "10/05/2017"
 bibliography: references.bib
@@
 library(matrixcalc)
+library(parallel)
 
@@
 Bullet matching, the process of determining whether two bullets could have been fired from the same gun barrel, has traditionally been performed without meaningful determination of error rates or statistical assessments of uncertainty [@NAS:2009].
-There have been some attempts towards developing mathematical and statistical approaches to bullet matching.One such attempt was the definition of CMS, the Consecutively Matching Striae [@biasotti:1959], with a cutoff of six to separate matches from non-matches.
+There have been some attempts towards developing mathematical and statistical approaches to bullet matching.
+One such attempt was the definition of CMS, the Consecutively Matching Striae [@biasotti:1959], with a cutoff of six to separate matches from non-matches.
 Still, rigorous assessments of the applicability of such cutoffs have not to this point been described [@pcast2016].
@@
 While the algorithm had a very strong performance on this set, some limitations were immediately clear.
-For instance, performance was assessed only on this single set of 35 bullets fired from a consecutively manufactured set of only ten gun barrels.
+For instance, performance was assessed only on this single set of 35 bullets fired from a consecutively manufactured set of only ten known and 15 unknown gun barrels.
 Each of these bullets was part of controlled study, and the full lands were available for matching.
@@
 
-Real world applications of bullet matching often involve the recovery of fragments of bullets from the crime scene.
+Real world applications of bullet matching often involve the recovery of fragments of bullets from the crime scene [@national2004forensic].
 Traditional features used in forensic examination work well for a full land, but there has been less investigation into their performance in the case of a fragmented land.
-For example, the CMS is naturally limited by the portion of the land that can be recovered.
+For example, the CMS is naturally limited by the portion of the land that can be recovered and varies across manufacturers [@chu:2011].
 
@@
 These features are computed on *aligned pairs of bullet land impressions* rather than on individual lands.
-This enables us, for instance, to compute the number of matching striae.
-Table \ref{tab:ccf1} and Table \ref{tab:ccf2} provide an example of six land-to-land comparisons and the derived features for these comparisons.
-The two `profile_id` columns identify a particular land from the two Hamby datasets, and the remaining columns are the derived features for these comparisons.
-
-```{r, echo=FALSE, results='asis'}
-result <- my_ccf %>%
-    filter(compare_id == 4) %>%
-    head %>%
-    collect %>%
-    select(2:8) %>%
-    as.data.frame
-
-print(xtable(result, digits = c(0, 0, 0, 4, 4, 4, 4, 4), label = "tab:ccf1", caption = "A sample of six land-to-land comparisons, and derived features for these comparisons."), comment = FALSE, include.rownames = FALSE)
-```
-
-```{r, echo=FALSE, results='asis'}
-result2 <- my_ccf %>%
-    filter(compare_id == 4) %>%
-    head %>%
-    collect %>%
-    select(9:ncol(.)) %>%
-    as.data.frame
-
-print(xtable(result2, digits = c(4, 4, 4, 4, 4, 4, 4, 4), label = "tab:ccf2", caption = "The remaining derived features for the previous six land-to-land comparisons."), comment = FALSE, include.rownames = FALSE)
-```
-
+This enables us, for instance, to compute the number of matching striae between two lands.
 We generalize the definitions of these features to account for the possibility that we may be handling degraded bullet lands, where only fragments can be recovered.
-The definition of each feature is given below, where $f(t)$ represents the height values of the first profile, and $g(t)$ the height values of the second:
+The definition of each feature is given below, where $f(t)$ represents the height values of the first profile at position $t$ along the signature, and $g(t)$ the height values of the second.
+An indication of whether the feature is new since @2016arXiv160105788H is also given:
 
@@
 The Cross-Correlation function is defined as $C(\tau) = \int_{-\infty}^{\infty} f(t)g(t + \tau)dt$ where $\tau$ represents the the lag of the second signature [@vorburger:2011].
-- **rough_cor** (%) is a new feature that quantifies the correlation between the two signatures after performing a second LOESS smoothing stage and then subtracting the result from the original signatures.
+- **rough_cor** (new) (%) feature that quantifies the correlation between the two signatures after performing a second LOESS smoothing stage and then subtracting the result from the original signatures.
 This attempts to model the roughness of the surface after removing structure such as waviness.
@@
 - **signature_length** (mm) is the overall length of the smallest of the two aligned signatures.
-- **overlap** (%) provides the percentage of the two signatures that overlap after the alignment stage.
+- **overlap** (new) (%) provides the percentage of the two signatures that overlap after the alignment stage.
 - **matches** (per mm) is the number of matching peaks/valleys (striae) per millimeter of the overlapping portion of the aligned signatures.
@@
 We can observe the distributions of both CCF and the Roughness Correlation side by side, differentiating between known matches and known non-matches.
-Figure \@ref(fig:roughcordist) shows this.
-It can be seen that the separation of the two groups along both variables is quite strong, but many known-matches have low values for each.
-The distributions of known non-matches, on the other hand, is relatively symmetric and normal.
-
-```{r roughcordist, echo=FALSE, fig.cap='Distributions of the Roughness Correlation compared to the CCF for known matches and known non-matches'}
+Figure \@ref(fig:roughcordist) displays this as an empirical CDF plot.
+It can be seen that the separation of known matches and known non-matches along both the CCF and the Roughness Correlation is quite strong and follows similar distributions (the known non-matches are relatively symmetric, while the known matches are very skewed left).
+However, some known non-matches with CCF values that would typically be indicative of a match have relatively lower values for the Roughness Correlation, which indicates that this feature could provide some added value when it comes to discriminating between matches and non-matches.
+
+```{r roughcordist, echo=FALSE, fig.cap='Empirical CDFs of the Roughness Correlation compared to the CCF for known matches and known non-matches.
+It can be seen that the distribution of each feature for the known non-matches is quite symmetric, while the distribution for each feature for the known matches is skewed left.'}
 cor_compare <- CCFs_withlands %>%
@@
 
-qplot(value, group=match, geom="density", data=cor_compare, fill=match, 
-      alpha=I(0.6), colour=I("grey20")) + 
+ggplot(data = cor_compare, aes(x = value, color = match)) +
+  stat_ecdf() +
   facet_wrap(~feature, nrow = 2) +
   theme_bw() + ylab("") + xlab("") +
-  scale_fill_brewer("Bullet Land Pairs", palette="Paired") +
+  scale_color_brewer("Bullet Land Pairs", palette="Paired") +
   theme(legend.background = element_rect(colour="grey75"),
@@
 Using these features, we can train a randomForest [@randomForest] model which attempts to predict whether two lands match given the value of the features.
-There are currently three studies for bullets included in the NIST ballistics database (REF).
+There are currently three studies for bullets included in the NIST Ballistics Toolmark Research Database.
 Those are Hamby (Set 252), Hamby (Set 44), and Cary.
-For purposes of analysis we describe in this paper, we exclude the Cary bullets from consideration, because the study was designed to assess the persistence of striation markings over a series of fires from the same barrel.
-Thus, every Cary bullet is a known match to every other Cary bullet, although early firings do not produce the same markings that later firings do.
+For purposes of the analysis we describe in this paper, we exclude the Cary bullets from consideration, because the study was designed to assess the persistence of striation markings over a series of fires from the same barrel.
+Thus, every Cary bullet is a known match to every other Cary bullet.
 Hence, we will consider Hamby (Set 252) and Hamby (Set 44) only.
+This leaves us a total of 83,028 land-to-land comparisons, of which 1208 are among known matching land impressions and 81,820 are among known non-matching land impressions.
 
 We can now train the forest using the features we defined earlier.
-We split the data into an 80% training, 20% testing sets to assess the algorithm's out of sample performance, using the `caret` package [@caretpkg].
-Table \ref{tab:avgforest} displays the results in the form of a confusion matrix on the test set, averaged over ten independent random forests trained on ten random data subsets.
-There are a total of 83,028 land-to-land comparisons, of which 1208 are among known matching land impressions and 81,820 are among known non-matching land impressions.
-It can be seen that false positives are exceedingly rare, but false negatives occur more frequently (approximately 65 false negative land to land comparisons on the test set, compared with an average of less than four false positives).
+Using the `caret` package [@caretpkg], we perform the following partitioning scheme.
+Out of the 50 barrels total, ten knowns and fifteen unknowns from each of the two Hamby sets, we hold out ten barrels randomly as a testing set, and use the remaining 40 to train the model.
+We repeat this procedure ten times and average the confusion matrix in order to assess the model accuracy with different holdout samples.
+Table \ref{tab:avgforest} displays the results in the form of a confusion matrix on the test set, averaged over these ten independent random forests trained on ten random barrel subsets.
+It can be seen that false positives are exceedingly rare, but false negatives occur more frequently (approximately 21` false negative land to land comparisons on the test set, compared with an average of less than two false positives).
 
@@
                                          "land_id.x", "land_id.y", "signature_length"))
+  CCFs_withlands <- CCFs_withlands %>%
+  mutate(barrel_id.x = interaction(study.x, barrel.x),
+         barrel_id.y = interaction(study.y, barrel.y))
 
 set.seed(20170222)
-trainIndex <- createDataPartition(1:nrow(CCFs_withlands), p = 0.8, list = TRUE, times = 10)
-match_result <- trainIndex %>% purrr::map_df(function(index) {
-    CCFs_train <- CCFs_withlands[index,]
-    CCFs_test = setdiff(CCFs_withlands, CCFs_train)
+mysamps <- replicate(10, expr = {sample(unique(CCFs_withlands$barrel_id.x), ceiling( length(unique(CCFs_withlands$barrel_id.x)) / 5))}, simplify = FALSE)
+
+cl <- makeCluster(10, renice = 0)
+clusterExport(cl, varlist = c("CCFs_withlands", "includes"))
+match_result <- parLapply(cl, mysamps, function(mysamp) {
+  library(dplyr)
+  library(randomForest)
+
+  CCFs_train <- CCFs_withlands %>%
+    filter(!(barrel_id.x %in% mysamp),
+         !(barrel_id.y %in% mysamp))
+
+  CCFs_test <- CCFs_withlands %>%
+    filter(barrel_id.x %in% mysamp,
+         barrel_id.y %in% mysamp)
 
@@
         as.data.frame
-
-    return(mydf)
 })
+
+match_result <- do.call(rbind, match_result)
+
 match_result <- cbind(id = rep(1:10, each = nrow(match_result) / 10), match_result)
@@
 Table \ref{tab:avgforeststudy} shows the average confusion matrix for the 10 random forests, broken down by study.
-It can be seen that Hamby252 to Hamby252 comparisons exhibit the fewest errors, while Hamby252 to Hamby44 comparisons exhibit the most errors on average.
+It can be seen that Hamby252 to Hamby252 comparisons exhibit the fewest errors, while Hamby44 to Hamby44 comparisons exhibit the most errors on average.
 This intuitively makes some sense given the potential presence of scanner operator effects, which we address further in this section.
@@
 print(xtable(result_wide_row, label = "tab:avgforeststudy", caption = "The average confusion matrix for the 10 random forests, broken down by study.
-It can be seen that Hamby252 to Hamby252 comparisons exhibit the fewest errors, while Hamby252 to Hamby44 comparisons exhibit the most on average."), comment = FALSE, include.rownames = FALSE, table.placement = "H")
+It can be seen that Hamby252 to Hamby252 comparisons exhibit the fewest errors, while Hamby44 to Hamby44 comparisons exhibit the most on average."), comment = FALSE, include.rownames = FALSE, table.placement = "H")
 ```
@@
 
-As a first stage to assessing feature robustness, we produce parallel coordinate plots of the various features based on true positives, true negatives, false positives, and false negatives.
-Figures \ref{fig:pcp} and \ref{fig:pcp2} display these plots.
-The means of the true positive and the false negative groups are shown, respectively, in the two figures.
-False positives tend to occur with anomalously high values of `sum_peaks`, `matches`, and `ccf`, while false negatives tend to resemble very closely the feature distribution of the true negatives, with a slightly higher average `ccf`.
-
-```{r pcp, echo=FALSE, message=FALSE, fig.cap="Parallel coordinate plot of the features based on the random forest confusion matrix for true and false positives.", fig.height=6, fig.width=10}
+Our goal is to assess the robustness of the previously defined features as it pertains to our bullet matching routines.
+This goal is both a backward looking assessment of our previous results for full land-to-land comparisons, and a forward looking one to help support the case that these can be used in the degraded land case.
+As a first stage to assessing this robustness, we produce parallel coordinate plots of the various features based on true positive, true negative, false positive, and false negative land-to-land matches.
+Figure \ref{fig:pcp} displays these plots.
+The means of the true positive and the true negative groups are shown as thick blue lines, respectively, in the two panels.
+The dashed lines represent individual land to land comparisons, with errors highlighted larger in red.
+It can be seen that the few false positives tendt o have anomalously high `ccf` or `matches`, while the false negatives tend have a lot of variability, though tending to also have a high `ccf` value.
+
+```{r pcp, echo=FALSE, message=FALSE, fig.cap="Parallel coordinate plot of the features based on the random forest confusion matrix for true and false positives (above), and true and false negatives (below).
+False positives tend to have some feature anomalously high, while false negatives exhibit quite a spread, sometimes having very large values of CCF or CMS, for example.", fig.height=6, fig.width=10}
 set.seed(20170222)
@@
                                          "bullet.x", "bullet.y", "land.x", "land.y",
-                                         "land_id.x", "land_id.y", "signature_length", "lag", "overlap"))
+                                         "land_id.x", "land_id.y", "signature_length", "lag", "overlap", "barrel_id.x", "barrel_id.y"))
 
@@
 
-ggplot() +
-  geom_line(data = filter(CCFs_tall, Type %in% c("True Positive", "False Positive")), aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type, size = Type)) +
-  geom_line(data = filter(pcp_data_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 3) +
+p1 <- ggplot() +
+  geom_line(data = filter(CCFs_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
+  geom_line(data = filter(CCFs_tall, Type %in% c("False Positive")), aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.8, linetype = "dashed") +
+  geom_line(data = filter(pcp_data_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
   ylab("Normalized Value") +
@@
   scale_color_manual(values = c("#e60000", "#000099")) +
-  scale_alpha_manual(values = c(1, .05)) +
   scale_size_manual(values = c(1, 1))
-```
-
-```{r pcp2, echo=FALSE, message=FALSE, fig.cap="Parallel coordinate plot of the features based on the random forest confusion matrix for true and false negatives.", fig.height=6, fig.width=10}
-mysamp <- sample(CCFs_tall$id[CCFs_tall$Type == "True Negative"], 1000)
+
+set.seed(20170222)
+mysamp1 <- sample(CCFs_tall$id[CCFs_tall$Type == "True Negative"], 226)
+mysamp2 <- sample(CCFs_tall$id[CCFs_tall$Type == "False Negative"], 20)
 trueneg_sample <- CCFs_tall %>%
-  filter(id %in% mysamp) %>%
-  rbind(CCFs_tall %>% filter(Type == "False Negative"))
-
-ggplot() +
+  filter(id %in% mysamp1)
+falseneg_sample <- CCFs_tall %>% 
+  filter(id %in% mysamp2)
+
+p2 <- ggplot() +
   #geom_line(data = filter(CCFs_tall, Type %in% c("True Negative", "False Negative")), aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type, size = Type)) +
-  geom_line(data = trueneg_sample, aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type)) +
+  geom_line(data = trueneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
+  geom_line(data = falseneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.4, linetype = "dashed") +
   theme_bw() +
-  geom_line(data = filter(pcp_data_tall, Type %in% c("False Negative")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 3) +
+  geom_line(data = filter(pcp_data_tall, Type %in% c("True Negative")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
   ylab("Normalized Value") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_color_manual(values = c("#ff9999", "#b3b3ff")) +
-  scale_alpha_manual(values = c(.5, .1))
+  scale_color_manual(values = c("#e60000", "#000099"))
+
+grid.arrange(p1, p2, nrow = 2)
 ```
@@
 
-We attempt to quantify the effect of the study on the matching probability by fitting a new random forest which is designed to predict the study based on the derived features.
+We attempt to quantify the effect of the study on the matching probability by fitting a new random forest which is designed to predict the study from which the scans came from based on the derived features.
+It should be noted that the two sets of scans were performed by two trained professionals at NIST, and therefore this analysis is intended to shed light on how slightly different operating procedures for the scans may lead to varying results in algorithms derived from these scans.
 Ideally, if the assumption of independence between lands holds across different operators, this forest should have poor performance - The set of derived features should be relatively consistent among known matches and known non-matches regardless of the study since the Hamby data in both sets originated from the same gun barrels.
@@
                                          "bullet.x", "bullet.y", "land.x", "land.y",
-                                         "land_id.x", "land_id.y", "study.x", "study.y", "signature_length"))
-
-result_study <- trainIndex %>% purrr::map_df(function(index) {
-    CCFs_train <- CCFs_withlands[index,]
-    CCFs_test = setdiff(CCFs_withlands, CCFs_train)
+                                         "land_id.x", "land_id.y", "study.x", "study.y", "barrel_id.x", "barrel_id.y", "signature_length"))
+
+clusterExport(cl, varlist = c("CCFs_withlands", "includes_study"))
+
+result_study <- parLapply(cl, mysamps, function(mysamp) {
+  library(dplyr)
+  library(randomForest)
+
+  CCFs_train <- CCFs_withlands %>%
+    filter(!(barrel_id.x %in% mysamp),
+         !(barrel_id.y %in% mysamp))
+
+  CCFs_test <- CCFs_withlands %>%
+    filter(barrel_id.x %in% mysamp,
+         barrel_id.y %in% mysamp)
 
@@
 
-    includes_study <- setdiff(names(CCFs_train), c("compare_id", "profile1_id", "profile2_id", "barrel.x", "barrel.y", "match", "bullet.x", "bullet.y", "land.x", "land.y", "land_id.x", "land_id.y", "study.x", "study.y"))
+    includes_study <- setdiff(names(CCFs_train), c("compare_id", "profile1_id", "profile2_id", "barrel.x", "barrel.y", "match", "bullet.x", "bullet.y", "land.x", "land.y", "land_id.x", "land_id.y", "study.x", "study.y", "barrel_id.x", "barrel_id.y", "forest"))
 
@@
 })
+
+stopCluster(cl)
 ```
@@
 Table \ref{tab:studypred} shows the confusion matrix, with column proportions, for the random forest with study as the response.
-It can be seen that while overall the random forest performs poorly, as hoped, comparisons between Hamby252 bullets are more distinguishable from other comparisons.
+It can be seen that indeed the random forest performs poorly, as hoped, indicating that a simple model to predict the study using the features available is not enough to detect the operator effects.
 
 ```{r, echo=FALSE, results='asis'}
-study_result <- cbind(id = rep(1:10, each = nrow(result_study) / 10), result_study)
+study_result <- do.call(rbind, result_study)
+study_result <- cbind(id = rep(1:10, each = nrow(study_result) / 10), study_result)
 study_result_wide <- study_result %>%
@@
 print(xtable(study_result_wide_col, digits = c(0, 0, 3, 3, 3), label = "tab:studypred", caption = "Confusion Matrix (Column Proportions) for the random forest with study as the response.
-It can be seen that while overall the random forest performs poorly, as hoped, comparisons between Hamby252 bullets is more distinguishable from other comparisons."), comment = FALSE, include.rownames = FALSE, table.placement = "H")
-```
-
-Figure \@ref(fig:ccfstudy) shows the distributions of the features defined above, faceted by whether the lands are known to be fired from the same gun barrel, across different study to study comparisons.
+It can be seen that the random forest performs poorly, as hoped, indicating that a simple model to predict the study using the features available is not enough to detect the operator effects."), comment = FALSE, include.rownames = FALSE, table.placement = "H")
+```
+
+Figure \@ref(fig:ccfstudy) shows the distributions of the land-to-land features, faceted by whether the lands are known to be fired from the same gun barrel, across different study to study comparisons.
 The distributions among the known non-matches seem relatively consistent across study based on visual inspection.
@@
 Table \ref{tab:kstests} gives the results of pairwise tests, for each feature, between different set comparisons, and between known matches compared with known non-matches.
-Although the tests are significant, looking at the raw values of the D statistic suggest that the largest effect sizes do in fact occur in comparisons with two Hamby252 lands, as the visual inspection of the boxplots also suggested.
+Although most of the tests are significant, looking at the raw values of the D statistic suggest that the largest effect sizes do in fact occur in comparisons with two Hamby252 lands, as the visual inspection of the boxplots also suggested.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, fig.pos='H', results='asis'}
+all_combinations <- combn(unique(c(as.character(CCFs_features$Study), as.character(CCFs_features$Study))), 2, simplify = FALSE)
+
+myresult <- all_combinations %>% map_df(function(x) {
+    final_result <- try({
+        set1 <- filter(CCFs_features, Study == x[1])
+        set2 <- filter(CCFs_features, Study == x[2])
+
+        result <- set1 %>%
+            rbind(set2) %>%
+            group_by(feature) %>%
+            do(matchtest = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Match"])$p.value,
+               matchd = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Match"])$statistic,
+               nonmatchtest = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Non-Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Non-Match"])$p.value,
+               nonmatchd = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Non-Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Non-Match"])$statistic,
+               set1_matchcount = length(filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Match"]),
+               set1_nonmatchcount = length(filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Non-Match"]),
+               set2_matchcount = length(filter(., Study == x[2])$value[filter(., Study == x[1])$match == "Known Match"]),
+               set2_nonmatchcount = length(filter(., Study == x[2])$value[filter(., Study == x[1])$match == "Known Non-Match"])
+
+            )
+    })
+
+    if (inherits(final_result, "try-error")) return(NULL)
+
+    return(final_result %>% mutate(set1 = x[1], set2 = x[2]) %>% select(set1, set2, everything()) %>% as.data.frame)
+})
+
+final_result <- myresult %>%
+    mutate_each(funs(unlist)) %>%
+    mutate(set1 = gsub("Hamby", "H", set1),
+           set2 = gsub("Hamby", "H", set2),
+           matchtest = round(matchtest, digits = 4),
+           nonmatchtest = round(nonmatchtest, digits = 4)) %>%
+    as.data.frame
+final_result$matchtest[final_result$matchtest < .0001] <- "< 0.0001"
+final_result$nonmatchtest[final_result$nonmatchtest < .0001] <- "< 0.0001"
+
+print(xtable(final_result[,1:7], digits = c(0, 0, 0, 0, 0, 4, 0, 4), label = "tab:kstests", caption = "Results for the Kolmogrov-Smirnov distributional test."), comment = FALSE, include.rownames = FALSE, table.placement = 'H')
+```
 
@@
 Indeed, another Kolmogorov-Smirnov test confirms a significant difference in the distributions of these values ($D = 0.6239, p < 0.0001$).
-This result strongly suggests that the operator effect in the bullet scanning procedure must be taken into account in order to assume pairwise independence.
+This result strongly suggests that the operator effect in the bullet scanning procedure must be taken into account in order to assume pairwise independence of bullet land scans between Hamby sets 252 and 44.
 
@@
 
+```{r deghist, echo=FALSE, fig.cap='Histograms of matching probability, facetted by the degradation level and known match versus known non-match.', fig.height=9.5, fig.width=8, fig.pos='H'}
+p1 <- ggplot(data = filter(CCFs_withlands_plot, degrade_left == 0), aes(x = forest)) +
+    geom_histogram() +
+    facet_grid(match~degrade_right, scales = "free_y") +
+    ggtitle("Left Fixed") +
+    theme_bw()
+p2 <- ggplot(data = filter(CCFs_withlands_plot, degrade_right == 0), aes(x = forest)) +
+    geom_histogram() +
+    facet_grid(match~degrade_left, scales = "free_y") +
+    ggtitle("Right Fixed") +
+    theme_bw()
+p3 <- ggplot(data = filter(CCFs_withlands_plot, (degrade_left == 0 & degrade_right == 0) | (degrade_right != 0 & degrade_left != 0)), aes(x = forest)) +
+    geom_histogram() +
+    facet_grid(match~degrade_left+degrade_right, scales = "free_y") +
+    ggtitle("Middle Fixed") +
+    theme_bw()
+
+grid.arrange(p1, p2, p3, nrow = 3)
+```
+
 ```{r senspe, echo=FALSE, fig.cap='Sensitivity and specificity of the random forest for given levels of degradation.
@@
 Feature \@ref(fig:featexp2) shows the feature expression for known non-matches by comparison.
+The non-matches don't exhibit the same pattern of better feature expression for the middle-fixed case, except perhaps for very low degradation levels.
+However, feature expression rises as a function of the land proportion, which indicates why the random forest begins predicting more positives, raising the sensitivity, but drastically lowering the specificity.
 
@@
 
-```{r featexp2, echo=FALSE, fig.cap='Feature expression for known non-matches, as a function of land proportion. Patterns of feature expression as a function are far less visible in this case.'}
+```{r featexp2, echo=FALSE, fig.cap='Feature expression for known non-matches, as a function of land proportion.
+As expected by the fact that the false positive rate increases as the land proportion decreases, so too does the feature expression.
+However, unlike for the known matches, fixing the middle portion of the land does not seem to lead to more expressed features, except perhaps for very low degradation levels.'}
 feature_exp2 <- CCFs_withlands_degrade %>%
@@
 Figure \@ref(fig:br924) provides an image of the surface of this land impression.
-Due to the tank rash, this particular land impression was originally excluded from consideration (see [@2016arXiv160105788H]).
+Due to the tank rash, this particular land impression was originally excluded from consideration (see [@2016arXiv160105788H]) based on our subjective assessment of the quality of the land.
 However, it appears that approximately half of the bullet land remains relatively unaffected.
@@
 
+```{r firstscore, echo=FALSE, fig.cap='Score distributions for the naive approach to bullet matching, for known matches and known non-matches.', fig.pos='H', fig.height=4, fig.width=7}
+CCFs_withlands$forest <- predict(rtrees, newdata = CCFs_withlands, type = "prob")[,2]
+
+result <- CCFs_withlands %>%
+    group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
+    summarise(score = mean(head(sort(forest, decreasing = TRUE))), match = any(match, na.rm = TRUE)) %>%
+    as.data.frame
+result$match <- factor(result$match, labels = c("Known non-matches (KNM)", "Known matches (KM)"))
+
+ggplot(data = result, aes(x = match, y = score, fill = match)) +
+    geom_boxplot() +
+    coord_flip() +
+    scale_fill_brewer("Bullet Land Pairs", palette="Paired") +
+    theme_bw() +
+    theme(legend.background = element_rect(colour="grey75"),
+          axis.text.y = element_blank(),
+          axis.title.y = element_blank())
+```
+
 We can improve on this approach by exploiting the rotation of the bullet to compute a score.
@@
 
-\begin{align}
+\begin{align*}
 P(M) &= 1 - P(NM) \\
@@
      &= 1 - ((1 - P(M1)) \times (1 - P(M2)) \times ... \times (1 - P(M6)))
-\end{align}
-
-Where $M$ is the event that two bullets match, $NM$ is the event that two bullets do not match, $M1$, $M2$, ..., $M6$ are the probabilities of land one, land two, ..., land six matching, and $NM1$, $NM2$, ..., $NM6$ are the probabilities that land one, land two, ..., land six do not match.
+\end{align*}
+
+where $M$ is the event that two bullets match, $NM$ is the event that two bullets do not match, $M1$, $M2$, ..., $M6$ are the probabilities of land one, land two, ..., land six matching, and $NM1$, $NM2$, ..., $NM6$ are the probabilities that land one, land two, ..., land six do not match.
 However, to compute this probability, we need to know the alignment of the two sets of lands.
@@
 This procedure is based on the Sequence Average Maximum (SAM) by @sensorfar in their bullet matching software application `SensoMatch`.
+A similar approach using the cross correlation maximum was first proposed by @chu:2010.
+Compared to that approach, ours uses random forest based probabilities compared to correlation values allowing for elements of probability theory to help determine the resulting bullet match probability.
 
 ```{r, echo=FALSE, results='asis'}
-CCFs_withlands$forest <- predict(rtrees, newdata = CCFs_withlands, type = "prob")[,2]
-
 result <- CCFs_withlands %>% 
@@
 
-We derive a score by computing the bullet matching probability on each set of six matrix diagonals using the previously defined formula, under the assumption of land to land indepndence.
+We now describe four methods for deriving a score from this matrix.
+The results derived from these methods are shown in Figure \@ref(fig:allscores).
+In **Method 1**, we derive a score by computing the bullet matching probability on each set of six matrix diagonals using the previously defined formula, under the assumption of land to land indepndence.
 Finally, we take the maximum score obtained out of the six results as the final matching score for a bullet pair.
 After doing so, we can plot the scores for known matches and known non-matches separately.
-Figure \@ref(fig:scores) provides the distribution of matching scores for known matches compared to known non-matches.
 It can be seen that the known matches all have scores of around 100%, while no non-match achieves a score of above 30%, and hence this procedure provides perfect discrimination between all pairs of bullets between and within the two Hamby datasets.
 
-```{r scores, echo=FALSE, fig.cap='Distribution of matching scores for known matches compared to known non-matches.
-It can be seen that the known matches all have scores of around 100%, while no non-match achieves a score of above 30%.'}
+```{r scores, echo=FALSE}
 myfunc <- function(x) {
@@
     filter(study.x != study.y | barrel.x != barrel.y | bullet.x != bullet.y)
-test2 <- CCFs_withlands %>%
+method1 <- CCFs_withlands %>%
     group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
@@
     left_join(test)
-
-ggplot(data = test2, aes(x = match, y = score)) +
-    geom_boxplot() +
-    theme_bw()
-```
-
-On the other hand, flipping this procedure around by assuming that a match occurs if and only if all six lands match does not discriminate quite as well, as can be seen in Figure \@ref(fig:scores1).
+```
+
+On the other hand, **Method 2** is obtained by flipping this procedure around by assuming that a match occurs if and only if all six lands match.
+As it turns out, this does not discriminate quite as well.
 Every known bullet non-match achieves a score of about zero, but so do about 15 known bullet matches.
@@
 
-```{r scores1, echo=FALSE, fig.cap='Distribution of matching scores for known matches compared to known non-matches when assuming a match occurs if and only if all six lands match.
-Now, it can be seen that the known non-matches all have scores of around 0, while most though not all known matches achieve much higher scores.'}
+```{r scores1, echo=FALSE}
 myfunc1 <- function(x) {
@@
     filter(study.x != study.y | barrel.x != barrel.y | bullet.x != bullet.y)
-test2 <- CCFs_withlands %>%
+method2 <- CCFs_withlands %>%
     group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
@@
     left_join(test)
-
-ggplot(data = test2, aes(x = match, y = score)) +
-    geom_boxplot() +
-    theme_bw()
-```
-
-If we construct a hybrid of these two approaches, we can average the probabilities along the diagonal rather than multiplying those probabilities.
-Doing so yields the distributions shown in Figure \@ref(fig:scores10).
+```
+
+**Method 3** is a hybrid of these two approaches, where we average the probabilities along the diagonal rather than multiplying those probabilities.
 Now, we once again differentiate the two groups well with no known non-match achieving a score above 10%, and no known match with a score below 40%.
 
-```{r scores10, echo=FALSE, fig.cap='Distribution of matching scores for known matches compared to known non-matches obtained by averaging the probabilities along the maximal diagonal.
-Once again, it can be seen that the known non-matches all have scores of around 0, but the known matches this time are well separated, with nearly all achieving scores above .5.'}
+```{r scores10, echo=FALSE}
 myfunc10 <- function(x) {
@@
     filter(study.x != study.y | barrel.x != barrel.y | bullet.x != bullet.y)
-test2 <- CCFs_withlands %>%
+method3 <- CCFs_withlands %>%
     group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
@@
     left_join(test)
-
-ggplot(data = test2, aes(x = match, y = score)) +
-    geom_boxplot() +
-    theme_bw()
-```
-
-One more approach to match bullets would exploit the SAM procedure on individual features.
+```
+
+One more approach to generate bullet matching scores, which we call **Method 4**, would exploit the SAM procedure on individual features.
 For each diagonal in the $6 \otimes 6$ matrix, we can compute an average value for each feature in our model.
@@
 We can then feed all six sets of features into the random forest in order to obtain a matching probability for each, taking the highest resulting probability to locate the diagonal and thus identify land to land alignment.
-Figure \@ref(fig:scores2) shows boxplots of the matching scores using this hybrid procedure.
 It can be seen that while this procedure does discriminate well, it yields some false negatives (matching bullets that our forest identifies as a non-match).
 
-```{r scores2, echo=FALSE, fig.cap='Distribution of matching scores using a SAM procedure on the feature values for known matches compared to known non-matches.'}
+```{r allscores, echo=FALSE, fig.cap='Distribution of matching scores using four methods.
+Method 1 assumes a match if at least one pair of lands match.
+Method 2 assumes a match if all pairs of lands match.
+Method 3 averages the probabilities instead of multiplying them.
+Finally, Method 4 uses a SAM procedure on the feature values for known matches compared to known non-matches.
+While these methods have various levels of discriminatory power, and rely on slightly different assumptions, they do show clear and significant separation between matching bullets and non-matching bullets in general.', fig.height=8.5, fig.width=8, fig.pos='H'}
 myfunc2 <- function(x, feature) {
@@
 
-final_result <- test2 %>%
+method4 <- test2 %>%
     left_join(final_hmm) %>%
@@
 
-ggplot(data = final_result, aes(x = match, y = forest)) +
+final_result <- method1 %>%
+  rbind(method2) %>%
+  rbind(method3) %>%
+  as.data.frame() %>%
+  rbind(as.data.frame(rename(method4, score = forest))) %>%
+  mutate(method = rep(1:4, each = nrow(method1)),
+         match = factor(match, labels = c("Known non-matches (KNM)", "Known matches (KM)")))
+
+mypaste <- function(strings) { paste("Method", strings) }
+
+ggplot(data = final_result, aes(x = match, y = score, fill = match)) +
     geom_boxplot() +
-    theme_bw()
+    coord_flip() +
+    facet_wrap(~method, nrow = 4, labeller = labeller(method = mypaste)) +
+    scale_fill_brewer("Bullet Land Pairs", palette="Paired") +
+    theme_bw() +
+    theme(legend.background = element_rect(colour="grey75"),
+          axis.text.y = element_blank(),
+          axis.title.y = element_blank())
 ```
@@
 In doing so, we noted strong evidence of operator effects that resulted in differences in the quality of the microscope scans.
+These effects were noted despite the experience of the individuals conducting the scans, which implies that such effects could quite likely be more pronounced when scans are done by those more inexperienced, or with fewer standard operating procedures in place.
 
@@
 
-We begain exploring the robustness of the matching algorithm proposed in [@2016arXiv160105788H] to land degradation.
+We began exploring the robustness of the matching algorithm proposed in [@2016arXiv160105788H] to land degradation.
 As suspected, the algorithm performance declines as a function of the rate of degradation.
@@
 
-As we have stated before, the lack of 3D images of bullets limits the extent to which these algorithms can be tested and validated.
+Finally, one pleasing conclusion from these results is the fact that generalizing them to full bullet comparisons rather than the land to land level appears to work quite well.
+Depending on the assumptions made, the out of sample accuracy of bullet to bullet comparisons can range from nearly perfect to perfect.
+This result is encouraging in that real world use of these algorithms would be done on the bullet level, assuming enough of the bullet was recovered to make these procedures possible.
+
+As we have stated before, the lack of 3D images of bullets available in the public domain limits the extent to which these algorithms can be tested and validated.
 The degraded land simulation itself may be too simplistic and not faithfully represent realistic scenarios.
-However, as more data are collected, we can continue to update, train and test the matching algorithm in order to improve its performance in real datasets.
+However, as more data are collected, we can continue to update, train, and test the matching algorithm in order to improve its performance in real datasets.
 
@@
 
-```{r, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, fig.pos='H', results='asis'}
-all_combinations <- combn(unique(c(as.character(CCFs_features$Study), as.character(CCFs_features$Study))), 2, simplify = FALSE)
-
-myresult <- all_combinations %>% map_df(function(x) {
-    final_result <- try({
-        set1 <- filter(CCFs_features, Study == x[1])
-        set2 <- filter(CCFs_features, Study == x[2])
-
-        result <- set1 %>%
-            rbind(set2) %>%
-            group_by(feature) %>%
-            do(matchtest = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Match"])$p.value,
-               matchd = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Match"])$statistic,
-               nonmatchtest = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Non-Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Non-Match"])$p.value,
-               nonmatchd = ks.test(x = filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Non-Match"], y = filter(., Study == x[2])$value[filter(., Study == x[2])$match == "Known Non-Match"])$statistic,
-               set1_matchcount = length(filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Match"]),
-               set1_nonmatchcount = length(filter(., Study == x[1])$value[filter(., Study == x[1])$match == "Known Non-Match"]),
-               set2_matchcount = length(filter(., Study == x[2])$value[filter(., Study == x[1])$match == "Known Match"]),
-               set2_nonmatchcount = length(filter(., Study == x[2])$value[filter(., Study == x[1])$match == "Known Non-Match"])
-
-            )
-    })
-
-    if (inherits(final_result, "try-error")) return(NULL)
-
-    return(final_result %>% mutate(set1 = x[1], set2 = x[2]) %>% select(set1, set2, everything()) %>% as.data.frame)
-})
-
-final_result <- myresult %>%
-    mutate_each(funs(unlist)) %>%
-    mutate(set1 = gsub("Hamby", "H", set1),
-           set2 = gsub("Hamby", "H", set2),
-           matchtest = round(matchtest, digits = 4),
-           nonmatchtest = round(nonmatchtest, digits = 4)) %>%
-    as.data.frame
-final_result$matchtest[final_result$matchtest < .0001] <- "< 0.0001"
-final_result$nonmatchtest[final_result$nonmatchtest < .0001] <- "< 0.0001"
-
-print(xtable(final_result[,1:7], digits = c(0, 0, 0, 0, 0, 4, 0, 4), label = "tab:kstests", caption = "Results for the Kolmogrov-Smirnov distributional test."), comment = FALSE, include.rownames = FALSE)
-```
-
-```{r eval=FALSE, deghist, echo=FALSE, fig.cap='Histograms of matching probability, facetted by the degradation level and known match versus known non-match.', fig.height=9.5, fig.width=8, fig.pos='H'}
-p1 <- ggplot(data = filter(CCFs_withlands_plot, degrade_left == 0), aes(x = forest)) +
-    geom_histogram() +
-    facet_grid(match~degrade_right, scales = "free_y") +
-    ggtitle("Left Fixed") +
-    theme_bw()
-p2 <- ggplot(data = filter(CCFs_withlands_plot, degrade_right == 0), aes(x = forest)) +
-    geom_histogram() +
-    facet_grid(match~degrade_left, scales = "free_y") +
-    ggtitle("Right Fixed") +
-    theme_bw()
-p3 <- ggplot(data = filter(CCFs_withlands_plot, (degrade_left == 0 & degrade_right == 0) | (degrade_right != 0 & degrade_left != 0)), aes(x = forest)) +
-    geom_histogram() +
-    facet_grid(match~degrade_left+degrade_right, scales = "free_y") +
-    ggtitle("Middle Fixed") +
-    theme_bw()
-
-grid.arrange(p1, p2, p3, nrow = 3)
-```
-
-```{r eval=FALSE, firstscore, echo=FALSE, fig.cap='Score distributions for the naive approach to bullet matching, for known matches and known non-matches.', fig.pos='H'}
-result <- CCFs_withlands %>%
-    group_by(study.x, barrel.x, bullet.x, study.y, barrel.y, bullet.y) %>%
-    summarise(forest = mean(head(sort(forest, decreasing = TRUE))), match = any(match))
-
-ggplot(data = result, aes(x = match, y = forest)) +
-    geom_boxplot() +
-    theme_bw()
-```
-
 # References

### 2017-10-03 degraded-paper.tex
Paper updates

--- 2017-09-06 degraded-paper.tex
@@
   \postdate{\par}
-  \date{9/5/2017}
+  \date{10/05/2017}
 
@@
 (National Research Council 2009). There have been some attempts towards
-developing mathematical and statistical approaches to bullet
-matching.One such attempt was the definition of CMS, the Consecutively
-Matching Striae (Biasotti 1959), with a cutoff of six to separate
-matches from non-matches. Still, rigorous assessments of the
-applicability of such cutoffs have not to this point been described
-(Advisors on Science and Technology 2016).
+developing mathematical and statistical approaches to bullet matching.
+One such attempt was the definition of CMS, the Consecutively Matching
+Striae (Biasotti 1959), with a cutoff of six to separate matches from
+non-matches. Still, rigorous assessments of the applicability of such
+cutoffs have not to this point been described (Advisors on Science and
+Technology 2016).
 
@@
 instance, performance was assessed only on this single set of 35 bullets
-fired from a consecutively manufactured set of only ten gun barrels.
-Each of these bullets was part of controlled study, and the full lands
-were available for matching. While there were some data quality issues,
-this was still a near ideal test case for the algorithm.
+fired from a consecutively manufactured set of only ten known and 15
+unknown gun barrels. Each of these bullets was part of controlled study,
+and the full lands were available for matching. While there were some
+data quality issues, this was still a near ideal test case for the
+algorithm.
 
 Real world applications of bullet matching often involve the recovery of
-fragments of bullets from the crime scene. Traditional features used in
-forensic examination work well for a full land, but there has been less
-investigation into their performance in the case of a fragmented land.
-For example, the CMS is naturally limited by the portion of the land
-that can be recovered.
+fragments of bullets from the crime scene (Council and others 2004).
+Traditional features used in forensic examination work well for a full
+land, but there has been less investigation into their performance in
+the case of a fragmented land. For example, the CMS is naturally limited
+by the portion of the land that can be recovered and varies across
+manufacturers (W. Chu et al. 2011).
 
@@
 impressions} rather than on individual lands. This enables us, for
-instance, to compute the number of matching striae. Table \ref{tab:ccf1}
-and Table \ref{tab:ccf2} provide an example of six land-to-land
-comparisons and the derived features for these comparisons. The two
-\texttt{profile\_id} columns identify a particular land from the two
-Hamby datasets, and the remaining columns are the derived features for
-these comparisons.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrr}
-  \hline
-profile1\_id & profile2\_id & ccf & rough\_cor & lag & D & sd\_D \\ 
-  \hline
-49 & 540 & 0.2689 & 0.1091 & -0.1734 & 0.0026 & 0.0043 \\ 
-  49 & 1032 & 0.3878 & 0.3582 & 0.1938 & 0.0017 & 0.0030 \\ 
-  49 & 1540 & 0.2410 & 0.1320 & 0.2781 & 0.0023 & 0.0039 \\ 
-  49 & 2044 & 0.2892 & 0.1136 & 0.0828 & 0.0023 & 0.0038 \\ 
-  49 & 2583 & 0.1769 & 0.0214 & -0.0563 & 0.0026 & 0.0041 \\ 
-  49 & 3074 & 0.2715 & 0.1254 & 0.0938 & 0.0020 & 0.0033 \\ 
-   \hline
-\end{tabular}
-\caption{A sample of six land-to-land comparisons, and derived features for these comparisons.} 
-\label{tab:ccf1}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrr}
-  \hline
-signature\_length & overlap & matches & mismatches & cms & non\_cms & sum\_peaks \\ 
-  \hline
-1.9328 & 0.9103 & 1.1368 & 10.2251 & 0.5684 & 3.7182 & 2.0038 \\ 
-  1.9359 & 0.9193 & 5.0571 & 6.5593 & 3.3714 & 2.3426 & 6.5264 \\ 
-  1.9359 & 0.8757 & 1.7696 & 10.3662 & 1.1797 & 5.8592 & 2.0192 \\ 
-  1.9172 & 0.9829 & 1.5920 & 7.9441 & 0.5307 & 3.4756 & 1.8139 \\ 
-  1.8234 & 0.9692 & 1.1317 & 8.8479 & 0.5659 & 4.9155 & 0.8555 \\ 
-  1.9172 & 0.9690 & 2.6913 & 10.7645 & 1.0765 & 3.4251 & 3.5272 \\ 
-   \hline
-\end{tabular}
-\caption{The remaining derived features for the previous six land-to-land comparisons.} 
-\label{tab:ccf2}
-\end{table}
-
-We generalize the definitions of these features to account for the
+instance, to compute the number of matching striae between two lands. We
+generalize the definitions of these features to account for the
 possibility that we may be handling degraded bullet lands, where only
 fragments can be recovered. The definition of each feature is given
-below, where \(f(t)\) represents the height values of the first profile,
-and \(g(t)\) the height values of the second:
+below, where \(f(t)\) represents the height values of the first profile
+at position \(t\) along the signature, and \(g(t)\) the height values of
+the second. An indication of whether the feature is new since Hare,
+Hofmann, and Carriquiry (2016) is also given:
 
@@
 \item
-  \textbf{rough\_cor} (\%) is a new feature that quantifies the
-  correlation between the two signatures after performing a second LOESS
-  smoothing stage and then subtracting the result from the original
-  signatures. This attempts to model the roughness of the surface after
-  removing structure such as waviness.
+  \textbf{rough\_cor} (new) (\%) feature that quantifies the correlation
+  between the two signatures after performing a second LOESS smoothing
+  stage and then subtracting the result from the original signatures.
+  This attempts to model the roughness of the surface after removing
+  structure such as waviness.
 \item
@@
 \item
-  \textbf{overlap} (\%) provides the percentage of the two signatures
-  that overlap after the alignment stage.
+  \textbf{overlap} (new) (\%) provides the percentage of the two
+  signatures that overlap after the alignment stage.
 \item
@@
 Correlation side by side, differentiating between known matches and
-known non-matches. Figure \ref{fig:roughcordist} shows this. It can be
-seen that the separation of the two groups along both variables is quite
-strong, but many known-matches have low values for each. The
-distributions of known non-matches, on the other hand, is relatively
-symmetric and normal.
+known non-matches. Figure \ref{fig:roughcordist} displays this as an
+empirical CDF plot. It can be seen that the separation of known matches
+and known non-matches along both the CCF and the Roughness Correlation
+is quite strong and follows similar distributions (the known non-matches
+are relatively symmetric, while the known matches are very skewed left).
+However, some known non-matches with CCF values that would typically be
+indicative of a match have relatively lower values for the Roughness
+Correlation, which indicates that this feature could provide some added
+value when it comes to discriminating between matches and non-matches.
 
@@
 \includegraphics{degraded-paper_files/figure-latex/roughcordist-1.pdf}
-\caption{\label{fig:roughcordist}Distributions of the Roughness Correlation
-compared to the CCF for known matches and known non-matches}
+\caption{\label{fig:roughcordist}Empirical CDFs of the Roughness Correlation
+compared to the CCF for known matches and known non-matches. It can be
+seen that the distribution of each feature for the known non-matches is
+quite symmetric, while the distribution for each feature for the known
+matches is skewed left.}
 \end{figure}
@@
 of the features. There are currently three studies for bullets included
-in the NIST ballistics database (REF). Those are Hamby (Set 252), Hamby
-(Set 44), and Cary. For purposes of analysis we describe in this paper,
-we exclude the Cary bullets from consideration, because the study was
-designed to assess the persistence of striation markings over a series
-of fires from the same barrel. Thus, every Cary bullet is a known match
-to every other Cary bullet, although early firings do not produce the
-same markings that later firings do. Hence, we will consider Hamby (Set
-252) and Hamby (Set 44) only.
+in the NIST Ballistics Toolmark Research Database. Those are Hamby (Set
+252), Hamby (Set 44), and Cary. For purposes of the analysis we describe
+in this paper, we exclude the Cary bullets from consideration, because
+the study was designed to assess the persistence of striation markings
+over a series of fires from the same barrel. Thus, every Cary bullet is
+a known match to every other Cary bullet. Hence, we will consider Hamby
+(Set 252) and Hamby (Set 44) only. This leaves us a total of 83,028
+land-to-land comparisons, of which 1208 are among known matching land
+impressions and 81,820 are among known non-matching land impressions.
 
-We can now train the forest using the features we defined earlier. We
-split the data into an 80\% training, 20\% testing sets to assess the
-algorithm's out of sample performance, using the \texttt{caret} package
-(Jed Wing et al. 2016). Table \ref{tab:avgforest} displays the results
-in the form of a confusion matrix on the test set, averaged over ten
-independent random forests trained on ten random data subsets. There are
-a total of 83,028 land-to-land comparisons, of which 1208 are among
-known matching land impressions and 81,820 are among known non-matching
-land impressions. It can be seen that false positives are exceedingly
-rare, but false negatives occur more frequently (approximately 65 false
-negative land to land comparisons on the test set, compared with an
-average of less than four false positives).
+We can now train the forest using the features we defined earlier. Using
+the \texttt{caret} package (Jed Wing et al. 2016), we perform the
+following partitioning scheme. Out of the 50 barrels total, ten knowns
+and fifteen unknowns from each of the two Hamby sets, we hold out ten
+barrels randomly as a testing set, and use the remaining 40 to train the
+model. We repeat this procedure ten times and average the confusion
+matrix in order to assess the model accuracy with different holdout
+samples. Table \ref{tab:avgforest} displays the results in the form of a
+confusion matrix on the test set, averaged over these ten independent
+random forests trained on ten random barrel subsets. It can be seen that
+false positives are exceedingly rare, but false negatives occur more
+frequently (approximately 21` false negative land to land comparisons on
+the test set, compared with an average of less than two false
+positives).
 
@@
   \hline
-False Negative & 65.3 \\ 
-  False Positive & 3.8 \\ 
-  True Negative & 16363.1 \\ 
-  True Positive & 171.8 \\ 
+False Negative & 20.7 \\ 
+  False Positive & 1.9 \\ 
+  True Negative & 3716.8 \\ 
+  True Positive & 56.5 \\ 
    \hline
@@
 random forests, broken down by study. It can be seen that Hamby252 to
-Hamby252 comparisons exhibit the fewest errors, while Hamby252 to
-Hamby44 comparisons exhibit the most errors on average. This intuitively
-makes some sense given the potential presence of scanner operator
-effects, which we address further in this section.
+Hamby252 comparisons exhibit the fewest errors, while Hamby44 to Hamby44
+comparisons exhibit the most errors on average. This intuitively makes
+some sense given the potential presence of scanner operator effects,
+which we address further in this section.
 
@@
   \hline
-Hamby252\_Hamby252 & 0.29\% & 0.01\% & 98.77\% & 0.93\% \\ 
-  Hamby252\_Hamby44 & 0.47\% & 0.02\% & 98.35\% & 1.16\% \\ 
-  Hamby44\_Hamby44 & 0.34\% & 0.04\% & 98.72\% & 0.9\% \\ 
+Hamby252\_Hamby252 & 0.39\% & 0.02\% & 97.56\% & 2.03\% \\ 
+  Hamby252\_Hamby44 & 0.45\% & 0.05\% & 98.31\% & 1.19\% \\ 
+  Hamby44\_Hamby44 & 0.89\% & 0.09\% & 97.68\% & 1.34\% \\ 
    \hline
@@
 \caption{The average confusion matrix for the 10 random forests, broken down by study.
-It can be seen that Hamby252 to Hamby252 comparisons exhibit the fewest errors, while Hamby252 to Hamby44 comparisons exhibit the most on average.}
+It can be seen that Hamby252 to Hamby252 comparisons exhibit the fewest errors, while Hamby44 to Hamby44 comparisons exhibit the most on average.}
 \label{tab:avgforeststudy}
@@
 
-As a first stage to assessing feature robustness, we produce parallel
-coordinate plots of the various features based on true positives, true
-negatives, false positives, and false negatives. Figures \ref{fig:pcp}
-and \ref{fig:pcp2} display these plots. The means of the true positive
-and the false negative groups are shown, respectively, in the two
-figures. False positives tend to occur with anomalously high values of
-\texttt{sum\_peaks}, \texttt{matches}, and \texttt{ccf}, while false
-negatives tend to resemble very closely the feature distribution of the
-true negatives, with a slightly higher average \texttt{ccf}.
+Our goal is to assess the robustness of the previously defined features
+as it pertains to our bullet matching routines. This goal is both a
+backward looking assessment of our previous results for full
+land-to-land comparisons, and a forward looking one to help support the
+case that these can be used in the degraded land case. As a first stage
+to assessing this robustness, we produce parallel coordinate plots of
+the various features based on true positive, true negative, false
+positive, and false negative land-to-land matches. Figure \ref{fig:pcp}
+displays these plots. The means of the true positive and the true
+negative groups are shown as thick blue lines, respectively, in the two
+panels. The dashed lines represent individual land to land comparisons,
+with errors highlighted larger in red. It can be seen that the few false
+positives tendt o have anomalously high \texttt{ccf} or
+\texttt{matches}, while the false negatives tend have a lot of
+variability, though tending to also have a high \texttt{ccf} value.
 
@@
 \caption{\label{fig:pcp}Parallel coordinate plot of the features based on
-the random forest confusion matrix for true and false positives.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{degraded-paper_files/figure-latex/pcp2-1.pdf}
-\caption{\label{fig:pcp2}Parallel coordinate plot of the features based on
-the random forest confusion matrix for true and false negatives.}
+the random forest confusion matrix for true and false positives (above),
+and true and false negatives (below). False positives tend to have some
+feature anomalously high, while false negatives exhibit quite a spread,
+sometimes having very large values of CCF or CMS, for example.}
 \end{figure}
@@
 probability by fitting a new random forest which is designed to predict
-the study based on the derived features. Ideally, if the assumption of
-independence between lands holds across different operators, this forest
-should have poor performance - The set of derived features should be
-relatively consistent among known matches and known non-matches
-regardless of the study since the Hamby data in both sets originated
-from the same gun barrels.
+the study from which the scans came from based on the derived features.
+It should be noted that the two sets of scans were performed by two
+trained professionals at NIST, and therefore this analysis is intended
+to shed light on how slightly different operating procedures for the
+scans may lead to varying results in algorithms derived from these
+scans. Ideally, if the assumption of independence between lands holds
+across different operators, this forest should have poor performance -
+The set of derived features should be relatively consistent among known
+matches and known non-matches regardless of the study since the Hamby
+data in both sets originated from the same gun barrels.
 
@@
 proportions, for the random forest with study as the response. It can be
-seen that while overall the random forest performs poorly, as hoped,
-comparisons between Hamby252 bullets are more distinguishable from other
-comparisons.
+seen that indeed the random forest performs poorly, as hoped, indicating
+that a simple model to predict the study using the features available is
+not enough to detect the operator effects.
 
@@
   \hline
-Hamby252\_Hamby252 & 28.5\% & 10.23\% & 3.24\% \\ 
-  Hamby252\_Hamby44 & 70.57\% & 83.27\% & 79.78\% \\ 
-  Hamby44\_Hamby44 & 0.93\% & 6.51\% & 16.98\% \\ 
+Hamby252\_Hamby252 & 9.6\% & 8.32\% & 11.28\% \\ 
+  Hamby252\_Hamby44 & 81.47\% & 81.06\% & 78.68\% \\ 
+  Hamby44\_Hamby44 & 8.93\% & 10.63\% & 10.04\% \\ 
    \hline
@@
 \caption{Confusion Matrix (Column Proportions) for the random forest with study as the response.
-It can be seen that while overall the random forest performs poorly, as hoped, comparisons between Hamby252 bullets is more distinguishable from other comparisons.}
+It can be seen that the random forest performs poorly, as hoped, indicating that a simple model to predict the study using the features available is not enough to detect the operator effects.}
 \label{tab:studypred}
@@
 
-Figure \ref{fig:ccfstudy} shows the distributions of the features
-defined above, faceted by whether the lands are known to be fired from
-the same gun barrel, across different study to study comparisons. The
+Figure \ref{fig:ccfstudy} shows the distributions of the land-to-land
+features, faceted by whether the lands are known to be fired from the
+same gun barrel, across different study to study comparisons. The
 distributions among the known non-matches seem relatively consistent
@@
 each feature, between different set comparisons, and between known
-matches compared with known non-matches. Although the tests are
+matches compared with known non-matches. Although most of the tests are
 significant, looking at the raw values of the D statistic suggest that
@@
 Hamby252 lands, as the visual inspection of the boxplots also suggested.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{llllrlr}
+  \hline
+set1 & set2 & feature & matchtest & matchd & nonmatchtest & nonmatchd \\ 
+  \hline
+H252\_H252 & H252\_H44 & ccf & $<$ 0.0001 & 0.2723 & 1e-04 & 0.0189 \\ 
+  H252\_H252 & H252\_H44 & cms & $<$ 0.0001 & 0.1751 & $<$ 0.0001 & 0.0245 \\ 
+  H252\_H252 & H252\_H44 & D & $<$ 0.0001 & 0.2567 & $<$ 0.0001 & 0.1049 \\ 
+  H252\_H252 & H252\_H44 & matches & $<$ 0.0001 & 0.1933 & $<$ 0.0001 & 0.0327 \\ 
+  H252\_H252 & H252\_H44 & mismatches & $<$ 0.0001 & 0.2015 & 0.3537 & 0.0079 \\ 
+  H252\_H252 & H252\_H44 & overlap & 0.0492 & 0.0984 & $<$ 0.0001 & 0.0276 \\ 
+  H252\_H252 & H252\_H44 & rough\_cor & $<$ 0.0001 & 0.2647 & $<$ 0.0001 & 0.0970 \\ 
+  H252\_H252 & H252\_H44 & sum\_peaks & 8e-04 & 0.1426 & 0.0015 & 0.0162 \\ 
+  H252\_H252 & H44\_H44 & ccf & $<$ 0.0001 & 0.2160 & $<$ 0.0001 & 0.0257 \\ 
+  H252\_H252 & H44\_H44 & cms & $<$ 0.0001 & 0.2515 & $<$ 0.0001 & 0.0467 \\ 
+  H252\_H252 & H44\_H44 & D & $<$ 0.0001 & 0.2342 & $<$ 0.0001 & 0.1946 \\ 
+  H252\_H252 & H44\_H44 & matches & $<$ 0.0001 & 0.2770 & $<$ 0.0001 & 0.0713 \\ 
+  H252\_H252 & H44\_H44 & mismatches & $<$ 0.0001 & 0.2505 & 0.0414 & 0.0138 \\ 
+  H252\_H252 & H44\_H44 & overlap & 0.2432 & 0.0906 & $<$ 0.0001 & 0.0408 \\ 
+  H252\_H252 & H44\_H44 & rough\_cor & $<$ 0.0001 & 0.2242 & $<$ 0.0001 & 0.1718 \\ 
+  H252\_H252 & H44\_H44 & sum\_peaks & 1e-04 & 0.1926 & $<$ 0.0001 & 0.0289 \\ 
+  H252\_H44 & H44\_H44 & ccf & 0.1149 & 0.0883 & $<$ 0.0001 & 0.0259 \\ 
+  H252\_H44 & H44\_H44 & cms & 0.111 & 0.0888 & $<$ 0.0001 & 0.0262 \\ 
+  H252\_H44 & H44\_H44 & D & 0.2923 & 0.0724 & $<$ 0.0001 & 0.0906 \\ 
+  H252\_H44 & H44\_H44 & matches & 0.0603 & 0.0977 & $<$ 0.0001 & 0.0423 \\ 
+  H252\_H44 & H44\_H44 & mismatches & 0.3301 & 0.0700 & 0.1633 & 0.0096 \\ 
+  H252\_H44 & H44\_H44 & overlap & 0.8231 & 0.0465 & 1e-04 & 0.0190 \\ 
+  H252\_H44 & H44\_H44 & rough\_cor & 0.2671 & 0.0741 & $<$ 0.0001 & 0.0769 \\ 
+  H252\_H44 & H44\_H44 & sum\_peaks & 0.047 & 0.1011 & 0.006 & 0.0147 \\ 
+   \hline
+\end{tabular}
+\caption{Results for the Kolmogrov-Smirnov distributional test.} 
+\label{tab:kstests}
+\end{table}
 
@@
 operator effect in the bullet scanning procedure must be taken into
-account in order to assume pairwise independence.
+account in order to assume pairwise independence of bullet land scans
+between Hamby sets 252 and 44.
 
@@
 \centering
+\includegraphics{degraded-paper_files/figure-latex/deghist-1.pdf}
+\caption{\label{fig:deghist}Histograms of matching probability, facetted by
+the degradation level and known match versus known non-match.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
 \includegraphics{degraded-paper_files/figure-latex/senspe-1.pdf}
@@
 impression. Feature \ref{fig:featexp2} shows the feature expression for
-known non-matches by comparison.
+known non-matches by comparison. The non-matches don't exhibit the same
+pattern of better feature expression for the middle-fixed case, except
+perhaps for very low degradation levels. However, feature expression
+rises as a function of the land proportion, which indicates why the
+random forest begins predicting more positives, raising the sensitivity,
+but drastically lowering the specificity.
 
@@
 \caption{\label{fig:featexp2}Feature expression for known non-matches, as a
-function of land proportion. Patterns of feature expression as a
-function are far less visible in this case.}
+function of land proportion. As expected by the fact that the false
+positive rate increases as the land proportion decreases, so too does
+the feature expression. However, unlike for the known matches, fixing
+the middle portion of the land does not seem to lead to more expressed
+features, except perhaps for very low degradation levels.}
 \end{figure}
@@
 land impression was originally excluded from consideration (see (Hare,
-Hofmann, and Carriquiry 2016)). However, it appears that approximately
-half of the bullet land remains relatively unaffected. We extract a
-signature from the unaffected half and attempt to match this signature
-to its full known match.
+Hofmann, and Carriquiry 2016)) based on our subjective assessment of the
+quality of the land. However, it appears that approximately half of the
+bullet land remains relatively unaffected. We extract a signature from
+the unaffected half and attempt to match this signature to its full
+known match.
 
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics{degraded-paper_files/figure-latex/firstscore-1.pdf}
+\caption{\label{fig:firstscore}Score distributions for the naive approach to
+bullet matching, for known matches and known non-matches.}
+\end{figure}
+
 We can improve on this approach by exploiting the rotation of the bullet
@@
      &= 1 - ((1 - P(M1)) \times (1 - P(M2)) \times ... \times (1 - P(M6)))
-\end{align}
+\end{align*}
 
-Where \(M\) is the event that two bullets match, \(NM\) is the event
+where \(M\) is the event that two bullets match, \(NM\) is the event
 that two bullets do not match, \(M1\), \(M2\), \ldots{}, \(M6\) are the
@@
 is based on the Sequence Average Maximum (SAM) by Sensorfar (2017) in
-their bullet matching software application \texttt{SensoMatch}.
+their bullet matching software application \texttt{SensoMatch}. A
+similar approach using the cross correlation maximum was first proposed
+by W. Chu et al. (2010). Compared to that approach, ours uses random
+forest based probabilities compared to correlation values allowing for
+elements of probability theory to help determine the resulting bullet
+match probability.
 
@@
 
-We derive a score by computing the bullet matching probability on each
-set of six matrix diagonals using the previously defined formula, under
-the assumption of land to land indepndence. Finally, we take the maximum
-score obtained out of the six results as the final matching score for a
-bullet pair. After doing so, we can plot the scores for known matches
-and known non-matches separately. Figure \ref{fig:scores} provides the
-distribution of matching scores for known matches compared to known
-non-matches. It can be seen that the known matches all have scores of
-around 100\%, while no non-match achieves a score of above 30\%, and
-hence this procedure provides perfect discrimination between all pairs
-of bullets between and within the two Hamby datasets.
+We now describe four methods for deriving a score from this matrix. The
+results derived from these methods are shown in Figure
+\ref{fig:allscores}. In \textbf{Method 1}, we derive a score by
+computing the bullet matching probability on each set of six matrix
+diagonals using the previously defined formula, under the assumption of
+land to land indepndence. Finally, we take the maximum score obtained
+out of the six results as the final matching score for a bullet pair.
+After doing so, we can plot the scores for known matches and known
+non-matches separately. It can be seen that the known matches all have
+scores of around 100\%, while no non-match achieves a score of above
+30\%, and hence this procedure provides perfect discrimination between
+all pairs of bullets between and within the two Hamby datasets.
+
+On the other hand, \textbf{Method 2} is obtained by flipping this
+procedure around by assuming that a match occurs if and only if all six
+lands match. As it turns out, this does not discriminate quite as well.
+Every known bullet non-match achieves a score of about zero, but so do
+about 15 known bullet matches. This method performs poorly because our
+matching algorithm exhibits a larger false negative rate than the rate
+of false positives. Multiplying the probabilities together compounds the
+issue of false negatives and leads to some misidentification of matching
+bullets.
+
+\textbf{Method 3} is a hybrid of these two approaches, where we average
+the probabilities along the diagonal rather than multiplying those
+probabilities. Now, we once again differentiate the two groups well with
+no known non-match achieving a score above 10\%, and no known match with
+a score below 40\%.
+
+One more approach to generate bullet matching scores, which we call
+\textbf{Method 4}, would exploit the SAM procedure on individual
+features. For each diagonal in the \(6 \otimes 6\) matrix, we can
+compute an average value for each feature in our model. This yields six
+sets of feature values for all six diagonals. We can then feed all six
+sets of features into the random forest in order to obtain a matching
+probability for each, taking the highest resulting probability to locate
+the diagonal and thus identify land to land alignment. It can be seen
+that while this procedure does discriminate well, it yields some false
+negatives (matching bullets that our forest identifies as a non-match).
 
@@
 \centering
-\includegraphics{degraded-paper_files/figure-latex/scores-1.pdf}
-\caption{\label{fig:scores}Distribution of matching scores for known matches
-compared to known non-matches. It can be seen that the known matches all
-have scores of around 100\%, while no non-match achieves a score of
-above 30\%.}
-\end{figure}
-
-On the other hand, flipping this procedure around by assuming that a
-match occurs if and only if all six lands match does not discriminate
-quite as well, as can be seen in Figure \ref{fig:scores1}. Every known
-bullet non-match achieves a score of about zero, but so do about 15
-known bullet matches. This method performs poorly because our matching
-algorithm exhibits a larger false negative rate than the rate of false
-positives. Multiplying the probabilities together compounds the issue of
-false negatives and leads to some misidentification of matching bullets.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{degraded-paper_files/figure-latex/scores1-1.pdf}
-\caption{\label{fig:scores1}Distribution of matching scores for known
-matches compared to known non-matches when assuming a match occurs if
-and only if all six lands match. Now, it can be seen that the known
-non-matches all have scores of around 0, while most though not all known
-matches achieve much higher scores.}
-\end{figure}
-
-If we construct a hybrid of these two approaches, we can average the
-probabilities along the diagonal rather than multiplying those
-probabilities. Doing so yields the distributions shown in Figure
-\ref{fig:scores10}. Now, we once again differentiate the two groups well
-with no known non-match achieving a score above 10\%, and no known match
-with a score below 40\%.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{degraded-paper_files/figure-latex/scores10-1.pdf}
-\caption{\label{fig:scores10}Distribution of matching scores for known
-matches compared to known non-matches obtained by averaging the
-probabilities along the maximal diagonal. Once again, it can be seen
-that the known non-matches all have scores of around 0, but the known
-matches this time are well separated, with nearly all achieving scores
-above .5.}
-\end{figure}
-
-One more approach to match bullets would exploit the SAM procedure on
-individual features. For each diagonal in the \(6 \otimes 6\) matrix, we
-can compute an average value for each feature in our model. This yields
-six sets of feature values for all six diagonals. We can then feed all
-six sets of features into the random forest in order to obtain a
-matching probability for each, taking the highest resulting probability
-to locate the diagonal and thus identify land to land alignment. Figure
-\ref{fig:scores2} shows boxplots of the matching scores using this
-hybrid procedure. It can be seen that while this procedure does
-discriminate well, it yields some false negatives (matching bullets that
-our forest identifies as a non-match).
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{degraded-paper_files/figure-latex/scores2-1.pdf}
-\caption{\label{fig:scores2}Distribution of matching scores using a SAM
-procedure on the feature values for known matches compared to known
-non-matches.}
+\includegraphics{degraded-paper_files/figure-latex/allscores-1.pdf}
+\caption{\label{fig:allscores}Distribution of matching scores using four
+methods. Method 1 assumes a match if at least one pair of lands match.
+Method 2 assumes a match if all pairs of lands match. Method 3 averages
+the probabilities instead of multiplying them. Finally, Method 4 uses a
+SAM procedure on the feature values for known matches compared to known
+non-matches. While these methods have various levels of discriminatory
+power, and rely on slightly different assumptions, they do show clear
+and significant separation between matching bullets and non-matching
+bullets in general.}
 \end{figure}
@@
 so, we noted strong evidence of operator effects that resulted in
-differences in the quality of the microscope scans.
+differences in the quality of the microscope scans. These effects were
+noted despite the experience of the individuals conducting the scans,
+which implies that such effects could quite likely be more pronounced
+when scans are done by those more inexperienced, or with fewer standard
+operating procedures in place.
 
@@
 
-We begain exploring the robustness of the matching algorithm proposed in
+We began exploring the robustness of the matching algorithm proposed in
 (Hare, Hofmann, and Carriquiry 2016) to land degradation. As suspected,
@@
 
-As we have stated before, the lack of 3D images of bullets limits the
-extent to which these algorithms can be tested and validated. The
-degraded land simulation itself may be too simplistic and not faithfully
-represent realistic scenarios. However, as more data are collected, we
-can continue to update, train and test the matching algorithm in order
-to improve its performance in real datasets.
+Finally, one pleasing conclusion from these results is the fact that
+generalizing them to full bullet comparisons rather than the land to
+land level appears to work quite well. Depending on the assumptions
+made, the out of sample accuracy of bullet to bullet comparisons can
+range from nearly perfect to perfect. This result is encouraging in that
+real world use of these algorithms would be done on the bullet level,
+assuming enough of the bullet was recovered to make these procedures
+possible.
+
+As we have stated before, the lack of 3D images of bullets available in
+the public domain limits the extent to which these algorithms can be
+tested and validated. The degraded land simulation itself may be too
+simplistic and not faithfully represent realistic scenarios. However, as
+more data are collected, we can continue to update, train, and test the
+matching algorithm in order to improve its performance in real datasets.
 
@@
 
+\hypertarget{ref-chu:2010}{}
+Chu, W., J. Song, T. Vorburger, J. Yen, S. Ballou, and B. Bachrach.
+2010. ``Pilot study of automated bullet signature identification based
+on topography measurements and correlations.'' \emph{J. Forensic Sci.}
+55 (2): 341--47.
+
 \hypertarget{ref-thompson:2013}{}
@@
 of the American Mathematical Society} 35 (4). JSTOR: 824--54.
+
+\hypertarget{ref-national2004forensic}{}
+Council, National Research, and others. 2004. \emph{Forensic Analysis:
+Weighing Bullet Lead Evidence}. National Academies Press.
 

### 2017-10-06 degraded-paper.Rmd
Small changes

@@
 We propose an approach to  standardize the featues, to account for the fact that only a portion of the land impression may be recovered from the crime scene.
-With the standardized features, we tackle two issues that were not addressed in  [@2016arXiv160105788H].
+With the standardized features, we tackle two issues that were not addressed in  @2016arXiv160105788H.
 The first is the effect of the microscope operator  on the resulting images and consequent algorithm performance.
-The second issue has to do with the robustness of the land matching algorithm in [@2016arXiv160105788H] relative to the degree of degradation of the questioned land impression.
+The second issue has to do with the robustness of the land matching algorithm in @2016arXiv160105788H relative to the degree of degradation of the questioned land impression.
 Finally, we describe some of the initial steps toward generalizing a matching algorithm based on land-to-land comparisons, to one based on bullet-to-bullet comparisons, as would be of interest in a real world application of these ideas.
@@
 study_result_wide_col <- study_result_wide
-study_result_wide_col[,-1] <- apply(study_result_wide_col[,-1], 2, function(x) paste0(round(100 * x / sum(x), digits = 2), "%"))
+study_result_wide_col[,-1] <- apply(study_result_wide_col[,-1], 2, function(x) {
+    mystr <- paste0(round(100 * x / sum(x), digits = 2), "%")
+    mystr[stringr::str_length(mystr) %in% c(4, 5)] <- paste0("0", mystr[stringr::str_length(mystr) %in% c(4, 5)])
+
+    return(mystr)
+})
 names(study_result_wide_col)[1] <- "Prediction \\ Actual"
@@
 
-Though visual inspection clearly shows differences, we can more formally assess the differences between distributions with a Kolmogrov-Smirnov test.
+Though visual inspection clearly shows differences, we can more formally assess the differences between distributions with a Kolmogorov-Smirnov test.
 Table \ref{tab:kstests} gives the results of pairwise tests, for each feature, between different set comparisons, and between known matches compared with known non-matches.
@@
 
-print(xtable(final_result[,1:7], digits = c(0, 0, 0, 0, 0, 4, 0, 4), label = "tab:kstests", caption = "Results for the Kolmogrov-Smirnov distributional test."), comment = FALSE, include.rownames = FALSE, table.placement = 'H')
+print(xtable(final_result[,1:7], digits = c(0, 0, 0, 0, 0, 4, 0, 4), label = "tab:kstests", caption = "Results for the Kolmogorov-Smirnov distributional test."), comment = FALSE, include.rownames = FALSE, table.placement = 'H')
 ```
@@
 
-print(xtable(tidy(mytest), digits = c(0, 4, 4, 0, 0), label = "tab:mykstest", caption = "Results for the Kolmogrov-Smirnov distributional test between values of the ideal cross section for Hamby252 compared with Hamby44."), comment = FALSE, include.rownames = FALSE)
+print(xtable(tidy(mytest), digits = c(0, 4, 4, 0, 0), label = "tab:mykstest", caption = "Results for the Kolmogorov-Smirnov distributional test between values of the ideal cross section for Hamby252 compared with Hamby44."), comment = FALSE, include.rownames = FALSE)
 ```
@@
 This occurs because the algorithm begins producing more positive predictions in general, likely as the result of the ccf being arbitrarily higher for known non-matches due to the small signature.
-On the other hand, the specificity drops dramatically for left, middle, and right fixed degraded lands when less than 50%.
-of the land impression is available for examination.
+On the other hand, the specificity drops dramatically for left, middle, and right fixed degraded lands when less than 50% of the land impression is available for examination.
 For a more in-depth exploration of the matching probabilities, Figure \@ref(fig:deghist) provides histograms of the matching probability  by  degradation level and by known match versus known non-match categories.
@@
 \centering
-\includegraphics[width=\linewidth]{images/br9-2-4-greyflip.png}
-\caption{Land 4 of Bullet 2, from Barrel 9 of Hamby Set 252. It can be seen that this particular land exhibits some major tank rash on the right half.}
+\includegraphics[width=\linewidth]{images/br9-2-4-grey.png}
+\caption{Land 4 of Bullet 2, from Barrel 9 of Hamby Set 252. It can be seen that this particular land exhibits some major tank rash on the left half.}
 \label{fig:br924}
@@
 
-Table \ref{tab:br924pred} shows the values of the features, after extracting only the first 50% of the Hamby Barrel 9 Bullet 2, 4th land (and hence, simulating a left-fixed 50% degraded scenario), compared with a feature comparison between both full lands (and hence, including the tank rash striae).
+Table \ref{tab:br924pred} shows the values of the features, after extracting only the last 50% of the Hamby Barrel 9 Bullet 2, 4th land (and hence, simulating a right-fixed 50% degraded scenario), compared with a feature comparison between both full lands (and hence, including the tank rash striae).
 The features are derived in a comparison with its known match, the complete Bullet 1 third land fired from Barrel 9.
@@
 The matching diagonal is clear based on the high probabilities (cell $(1, 3)$, cell $(2, 4)$, cell $(3, 5)$, etc.) although it can be seen that one of the six comparisons has a relatively lower matching probability.
-This procedure is based on the Sequence Average Maximum (SAM) by @sensorfar in their bullet matching software application `SensoMatch`.
+This procedure is based on the Sequence Average Maximum (SAM) by @sensofar in their bullet matching software application `SensoMatch`.
 A similar approach using the cross correlation maximum was first proposed by @chu:2010.
@@
 
-We began exploring the robustness of the matching algorithm proposed in [@2016arXiv160105788H] to land degradation.
+We began exploring the robustness of the matching algorithm proposed in @2016arXiv160105788H to land degradation.
 As suspected, the algorithm performance declines as a function of the rate of degradation.

### 2017-10-06 degraded-paper.tex
Small changes

@@
 non-matches. Still, rigorous assessments of the applicability of such
-cutoffs have not to this point been described (Advisors on Science and
-Technology 2016).
+cutoffs have not to this point been described (President's Council of
+Advisors on Science and Technology 2016).
 
@@
 Real world applications of bullet matching often involve the recovery of
-fragments of bullets from the crime scene (Council and others 2004).
-Traditional features used in forensic examination work well for a full
-land, but there has been less investigation into their performance in
-the case of a fragmented land. For example, the CMS is naturally limited
-by the portion of the land that can be recovered and varies across
-manufacturers (W. Chu et al. 2011).
+fragments of bullets from the crime scene (National Research Council
+2004). Traditional features used in forensic examination work well for a
+full land, but there has been less investigation into their performance
+in the case of a fragmented land. For example, the CMS is naturally
+limited by the portion of the land that can be recovered and varies
+across manufacturers (W. Chu et al. 2011).
 
@@
 from the crime scene. With the standardized features, we tackle two
-issues that were not addressed in (Hare, Hofmann, and Carriquiry 2016).
+issues that were not addressed in Hare, Hofmann, and Carriquiry (2016).
 The first is the effect of the microscope operator on the resulting
 images and consequent algorithm performance. The second issue has to do
-with the robustness of the land matching algorithm in (Hare, Hofmann,
-and Carriquiry 2016) relative to the degree of degradation of the
+with the robustness of the land matching algorithm in Hare, Hofmann, and
+Carriquiry (2016) relative to the degree of degradation of the
 questioned land impression. Finally, we describe some of the initial
@@
   \hline
-Hamby252\_Hamby252 & 9.6\% & 8.32\% & 11.28\% \\ 
+Hamby252\_Hamby252 & 09.6\% & 08.32\% & 11.28\% \\ 
   Hamby252\_Hamby44 & 81.47\% & 81.06\% & 78.68\% \\ 
-  Hamby44\_Hamby44 & 8.93\% & 10.63\% & 10.04\% \\ 
+  Hamby44\_Hamby44 & 08.93\% & 10.63\% & 10.04\% \\ 
    \hline
@@
 Though visual inspection clearly shows differences, we can more formally
-assess the differences between distributions with a Kolmogrov-Smirnov
+assess the differences between distributions with a Kolmogorov-Smirnov
 test. Table \ref{tab:kstests} gives the results of pairwise tests, for
@@
 \end{tabular}
-\caption{Results for the Kolmogrov-Smirnov distributional test.} 
+\caption{Results for the Kolmogorov-Smirnov distributional test.} 
 \label{tab:kstests}
@@
 specificity drops dramatically for left, middle, and right fixed
-degraded lands when less than 50\%. of the land impression is available
+degraded lands when less than 50\% of the land impression is available
 for examination. For a more in-depth exploration of the matching
@@
 Table \ref{tab:br924pred} shows the values of the features, after
-extracting only the first 50\% of the Hamby Barrel 9 Bullet 2, 4th land
-(and hence, simulating a left-fixed 50\% degraded scenario), compared
+extracting only the last 50\% of the Hamby Barrel 9 Bullet 2, 4th land
+(and hence, simulating a right-fixed 50\% degraded scenario), compared
 with a feature comparison between both full lands (and hence, including
@@
 We began exploring the robustness of the matching algorithm proposed in
-(Hare, Hofmann, and Carriquiry 2016) to land degradation. As suspected,
+Hare, Hofmann, and Carriquiry (2016) to land degradation. As suspected,
 the algorithm performance declines as a function of the rate of
@@
 \hypertarget{refs}{}
-\hypertarget{ref-pcast2016}{}
-Advisors on Science, President's Council of, and Technology. 2016.
-``Report on Forensic Science in Criminal Courts: Ensuring Scientific
-Validity of Feature-Comparison Methods.''
-\url{https://www.whitehouse.gov/sites/default/files/microsites/ostp/PCAST/pcast_forensic_science_report_final.pdf}.
-
 \hypertarget{ref-biasotti:1959}{}
@@
 
-\hypertarget{ref-national2004forensic}{}
-Council, National Research, and others. 2004. \emph{Forensic Analysis:
-Weighing Bullet Lead Evidence}. National Academies Press.
-
 \hypertarget{ref-giannelli:2011}{}
@@
 
+\hypertarget{ref-national2004forensic}{}
+National Research Council. 2004. \emph{Forensic Analysis: Weighing
+Bullet Lead Evidence}. National Academies Press.
+
 \hypertarget{ref-NAS:2009}{}
-National Research Council. 2009. \emph{Strengthening Forensic Science in
-the United States: A Path Forward}. Washington, DC: The National
-Academies Press.
+---------. 2009. \emph{Strengthening Forensic Science in the United
+States: A Path Forward}. Washington, DC: The National Academies Press.
 doi:\href{https://doi.org/10.17226/12589}{10.17226/12589}.
@@
 
+\hypertarget{ref-pcast2016}{}
+President's Council of Advisors on Science and Technology. 2016.
+``Report on Forensic Science in Criminal Courts: Ensuring Scientific
+Validity of Feature-Comparison Methods.''
+\url{https://www.whitehouse.gov/sites/default/files/microsites/ostp/PCAST/pcast_forensic_science_report_final.pdf}.
+
 \hypertarget{ref-riva:2014}{}
@@
 
-\hypertarget{ref-sensorfar}{}
+\hypertarget{ref-sensofar}{}
 Sensorfar. 2017. \emph{SensoMATCH Bullet Comparison Software}.

### 2017-10-06 response.Rmd
Acknowledgment

@@
 ---
-title: "Response to reviewers"
+title: "Reviewer Response and Acknowledgment"
 author: "Eric Hare, Heike Hofmann, Alicia Carriquiry"
@@
 ---
+
+# Acknowledgment
+
+We wanted to begin by acknowledging the reviewers.
+The reviews were very thorough and immensely helpful despite being done in a very short time frame.
+We greatly appreciate the extra effort put forth in order to help us get this manuscript included in the special issue.
+
+This great feedback has allowed us to effectively revise the paper.
+We hope that you find the revised version of the manuscript substantially improved.
+Below, we have addressed the individual points raised by the two reviewers.
 

### 2017-10-06 degraded-paper.tex
Small references change
Acknowledgment

@@
 comparisons has a relatively lower matching probability. This procedure
-is based on the Sequence Average Maximum (SAM) by Sensorfar (2017) in
+is based on the Sequence Average Maximum (SAM) by Sensofar (2017) in
 their bullet matching software application \texttt{SensoMatch}. A
@@
 
+\section{Acknowledgment}\label{acknowledgment}
+
+We wish to acknowledge the anonymous reviewers, whose quick yet thorough
+and immensely helpful reviews helped us improve this manuscript. We
+greatly appreciate the extra effort put forth for us.
+
 \clearpage
@@
 \hypertarget{ref-sensofar}{}
-Sensorfar. 2017. \emph{SensoMATCH Bullet Comparison Software}.
+Sensofar. 2017. \emph{SensoMATCH Bullet Comparison Software}.
 

### 2017-10-20 degraded-paper.Rmd
Acknowledgment
all but fig 13

@@
 
+custom <- list(
+  theme_bw(),
+  scale_linetype_manual(values=3:2),
+  scale_colour_manual(values=c("black", "grey40")),
+  ylab(expression(paste("Relative Height (in ",mu,m,")", sep=""))), 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
+  )
+
 p1 <- ggplot(data = final_profs, aes(x = y, y = l30, colour = profile_id)) +
-    geom_line() +
-    theme_bw()
+    geom_line(aes(linetype=profile_id)) +
+    custom
 
 p2 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l30, colour = profile_id)) +
-    geom_line() +
-    geom_smooth(inherit.aes = FALSE, aes(x = y, y = avgl30), colour = "purple") +
-    theme_bw()
+    geom_line(aes(linetype=profile_id)) +
+    geom_smooth(inherit.aes = FALSE, aes(x = y, y = avgl30), size = .5, se=FALSE, colour="black") +
+    custom
 
 p3 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l50, colour = profile_id)) +
-    geom_line() +
-    theme_bw()
+    geom_line(aes(linetype=profile_id)) +
+    geom_hline(yintercept=0, size = .5, colour="black") +
+    custom +
+  ylab("Adjusted relative height")
 
@@
 
-```{r roughcordist, echo=FALSE, fig.cap='Empirical CDFs of the Roughness Correlation compared to the CCF for known matches and known non-matches.
+```{r roughcordist, echo=FALSE, out.width='.5\\textwidth',  fig.cap='Empirical CDFs of the Roughness Correlation compared to the CCF for known matches and known non-matches.
 It can be seen that the distribution of each feature for the known non-matches is quite symmetric, while the distribution for each feature for the known matches is skewed left.'}
@@
 
-ggplot(data = cor_compare, aes(x = value, color = match)) +
+custom <- list(
+  theme_bw(),
+  scale_linetype_manual("Bullet Land Pairs", values=3:2),
+  scale_colour_manual("Bullet Land Pairs", values=c("black", "grey40"))
+  )
+
+ggplot(data = cor_compare, aes(x = value, colour=match, linetype = match)) +
   stat_ecdf() +
   facet_wrap(~feature, nrow = 2) +
-  theme_bw() + ylab("") + xlab("") +
-  scale_color_brewer("Bullet Land Pairs", palette="Paired") +
+  ylab("") + xlab("") +
   theme(legend.background = element_rect(colour="grey75"),
         axis.title.y=element_blank(),
-        plot.margin = unit(c(0,0,0,0), unit="cm"))
+        plot.margin = unit(c(0,0,0,0), unit="cm")) +
+  custom +
+    theme(legend.position=c(0.2, 0.85)) 
 ```
@@
   select(id, everything()) %>%
-  gather(key = Feature, value = Value, 3:ncol(.))
+  gather(key = Feature, value = Value, 3:ncol(.)) %>%
+  mutate(x = as.numeric(as.factor(Feature)))
+
+truepos <- filter(CCFs_tall, Type %in% c("True Positive"))
+falsepos <- filter(CCFs_tall, Type %in% c("False Positive"))
+
+truepos_poly <- truepos %>% 
+  group_by(Feature) %>% 
+  summarize(min = min(Value), max=max(Value), mean=mean(Value), x=x[1])
+tpps <- truepos_poly %>% gather(Type, Result, min:max)
+tpps$order <-c(1:8, 9+8:1)
+tpps$x <-c(1:8, 1:8)
+
+tpps <- tpps %>% arrange(order)
 
 p1 <- ggplot() +
-  geom_line(data = filter(CCFs_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
-  geom_line(data = filter(CCFs_tall, Type %in% c("False Positive")), aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.8, linetype = "dashed") +
-  geom_line(data = filter(pcp_data_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
+  geom_polygon(aes(x = x, y=Result), data = tpps, fill="grey50", alpha=.35) +
+  geom_line(aes(x = x, y = mean), data = truepos_poly, colour="grey50", size=1) +
+  geom_line(data = truepos, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 0.1, size = 0.2) + #, linetype = "dashed") +
+  scale_x_continuous(breaks=1:8, labels=unique(tpps$Feature)) +
+  geom_line(data = falsepos, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 1, size = 0.8) +
   ylab("Normalized Value") +
+  xlab("") +
   theme_bw() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_color_manual(values = c("#e60000", "#000099")) +
-  scale_size_manual(values = c(1, 1))
+  scale_color_manual(values = c("grey10", "grey50")) +
+#  scale_color_manual(values = c("#e60000", "#000099")) +
+  scale_size_manual(values = c(1, 1)) +
+  scale_linetype_manual(values=2:1)
+
+# p1 <- ggplot() +
+# #  geom_polygon(aes(x = x, y=Result), data = tpps, fill="#000099", alpha=.5) +
+#   geom_line(data = truepos, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
+#   geom_line(data = falsepos, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.8, linetype = "dashed") +
+#   geom_line(data = filter(pcp_data_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
+#   ylab("Normalized Value") +
+#   theme_bw() +
+#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+#   scale_color_manual(values = c("#e60000", "#000099")) +
+#   scale_size_manual(values = c(1, 1))
 
@@
 mysamp2 <- sample(CCFs_tall$id[CCFs_tall$Type == "False Negative"], 20)
+
+CCFs_tall <- CCFs_tall %>% mutate(x = as.numeric(as.factor(Feature)))
+
 trueneg_sample <- CCFs_tall %>%
@@
 
+trueneg_poly <- trueneg_sample %>% 
+  group_by(Feature) %>% 
+  summarize(min = min(Value), max=max(Value), mean=mean(Value), x=x[1])
+tnps <- trueneg_poly %>% gather(Type, Result, min:max)
+tnps$order <-c(1:8, 9+8:1)
+tnps$x <-c(1:8, 1:8)
+
+tnps <- tnps %>% arrange(order)
+
 p2 <- ggplot() +
-  #geom_line(data = filter(CCFs_tall, Type %in% c("True Negative", "False Negative")), aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type, size = Type)) +
-  geom_line(data = trueneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
-  geom_line(data = falseneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.4, linetype = "dashed") +
+  geom_polygon(aes(x = x, y=Result), data = tnps, fill="grey50", alpha=.35) +
+  geom_line(aes(x = x, y = mean), data = trueneg_poly, colour="grey50", size=1) +
+  geom_line(data = trueneg_sample, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 0.1, size = 0.2) + #, linetype = "dashed") +
+scale_x_continuous(breaks=1:8, labels=unique(tpps$Feature)) +
+  geom_line(data = falseneg_sample, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 1, size = 0.8) +
+  ylab("Normalized Value") +
+  xlab("") +
   theme_bw() +
-  geom_line(data = filter(pcp_data_tall, Type %in% c("True Negative")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
-  ylab("Normalized Value") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_color_manual(values = c("#e60000", "#000099"))
+  scale_color_manual(values = c("grey10", "grey50")) +
+#  scale_color_manual(values = c("#e60000", "#000099")) +
+  scale_size_manual(values = c(1, 1)) +
+  scale_linetype_manual(values=2:1)
+
+# p2 <- ggplot() +
+#   #geom_line(data = filter(CCFs_tall, Type %in% c("True Negative", "False Negative")), aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type, size = Type)) +
+#   geom_line(data = trueneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
+#   geom_line(data = falseneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.4, linetype = "dashed") +
+#   theme_bw() +
+#   geom_line(data = filter(pcp_data_tall, Type %in% c("True Negative")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
+#   ylab("Normalized Value") +
+#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+#   scale_color_manual(values = c("#e60000", "#000099"))
 
@@
 ## Produce plot
-ggplot(result, aes(x = level, y = value, colour = fixed)) +
+ggplot(result, aes(x = level, y = value, colour = fixed, linetype=fixed)) +
     facet_wrap(~metric, nrow = 2) +
@@
     scale_x_continuous(breaks = c(0, .125, .25, .375, .5, .625, .75), labels = c("100%", "87.5%", "75%", "62.5%", "50%", "37.5%", "25%")) +
-    xlab("Land Proportion")
+    xlab("Land Proportion") +
+  scale_colour_manual(values=c("black", "grey20", "grey30", "grey40")) +
+  theme(legend.key.width=unit(3,"line"))
 ```
@@
 
-ggplot(data = feature_exp, aes(x = level, y = value, colour = fixed)) +
+ggplot(data = feature_exp, aes(x = level, y = value, colour = fixed, linetype=fixed)) +
     geom_point() +
@@
     xlab("Land Proportion") +
-    facet_wrap(~feature, scales = "free_y")
+    facet_wrap(~feature, scales = "free_y") +
+    scale_colour_manual(values=c("black", "grey20", "grey30", "grey40")) +
+  theme(legend.key.width=unit(3,"line"))
+
 ```
@@
 
-ggplot(data = feature_exp2, aes(x = level, y = value, colour = fixed)) +
+ggplot(data = feature_exp2, aes(x = level, y = value, colour = fixed, linetype=fixed)) +
     geom_point() +
@@
     theme_bw() +
-    theme(legend.background = element_rect(colour="grey75"),
+    theme(legend.position="none",
           axis.text.y = element_blank(),
-          axis.title.y = element_blank())
+          axis.title.y = element_blank()) +
+  annotate("text", x=2.25, y= 0.8, label="Known matches (KM)", hjust=1, colour="grey40") +
+  annotate("text", x=0.75, y= 0.05, label="Known non-matches (KNM)", hjust=0, colour="grey40")
+
 ```
@@
 
+# Acknowledgment
+
+We wish to acknowledge the anonymous reviewers, whose quick yet thorough and immensely helpful reviews helped us improve this manuscript. We greatly appreciate the extra effort put forth for us.
+
 \clearpage

### 2017-10-21 degraded-paper.Rmd
fig 13 ...

@@
           axis.text.y = element_blank(),
-          axis.title.y = element_blank())
+          axis.title.y = element_blank()) +
+  theme(legend.position="none") +
+  annotate("text", x=2.25, y= 0.8, label="Known matches (KM)", hjust=1, colour="grey40") +
+  annotate("text", x=0.75, y= 0.05, label="Known non-matches (KNM)", hjust=0, colour="grey40")
 ```

### 2017-10-22 degraded-paper.Rmd
fig 13 now not blind

--- 2017-10-20 degraded-paper.Rmd
@@
 
+dframe <- data.frame(expand.grid(match=c("Known non-matches (KNM)", "Known matches (KM)"), method=1:4))
+dframe$x <- c(0.75, 2.25)[as.numeric(as.factor(dframe$match))]
+dframe$y <- c(0.15, 0.9)[as.numeric(as.factor(dframe$match))]
+idx <- with(dframe, which(method==1 & match=="Known non-matches (KNM)"))
+dframe$y[idx] <- 0.2
+idx <- with(dframe, which(method==1 & match=="Known matches (KM)"))
+dframe$y[idx] <- 0.875
+idx <- with(dframe, which(method==3 & match=="Known matches (KM)"))
+dframe$y[idx] <- 0.7
+idx <- with(dframe, which(method==4 & match=="Known matches (KM)"))
+dframe$x[idx] <- 1.45
+
 mypaste <- function(strings) { paste("Method", strings) }
-
 ggplot(data = final_result, aes(x = match, y = score, fill = match)) +
@@
           axis.text.y = element_blank(),
-          axis.title.y = element_blank())
+          axis.title.y = element_blank()) +
+  theme(legend.position="none") +
+  geom_text(aes(x=x,y=y, label=match), colour="grey40", data=dframe)
+
 ```

### 2017-10-22 degraded-paper.Rmd
I don't remember what changes these were

--- 2017-10-20 degraded-paper.Rmd
@@
 
-custom <- list(
-  theme_bw(),
-  scale_linetype_manual(values=3:2),
-  scale_colour_manual(values=c("black", "grey40")),
-  ylab(expression(paste("Relative Height (in ",mu,m,")", sep=""))), 
-  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
-  )
-
 p1 <- ggplot(data = final_profs, aes(x = y, y = l30, colour = profile_id)) +
-    geom_line(aes(linetype=profile_id)) +
-    custom
+    geom_line() +
+    theme_bw()
 
 p2 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l30, colour = profile_id)) +
-    geom_line(aes(linetype=profile_id)) +
-    geom_smooth(inherit.aes = FALSE, aes(x = y, y = avgl30), size = .5, se=FALSE, colour="black") +
-    custom
+    geom_line() +
+    geom_smooth(inherit.aes = FALSE, aes(x = y, y = avgl30), colour = "purple") +
+    theme_bw()
 
 p3 <- ggplot(data = final_doublesmoothed, aes(x = y, y = l50, colour = profile_id)) +
-    geom_line(aes(linetype=profile_id)) +
-    geom_hline(yintercept=0, size = .5, colour="black") +
-    custom +
-  ylab("Adjusted relative height")
+    geom_line() +
+    theme_bw()
 
@@
 
-```{r roughcordist, echo=FALSE, out.width='.5\\textwidth',  fig.cap='Empirical CDFs of the Roughness Correlation compared to the CCF for known matches and known non-matches.
+```{r roughcordist, echo=FALSE, fig.cap='Empirical CDFs of the Roughness Correlation compared to the CCF for known matches and known non-matches.
 It can be seen that the distribution of each feature for the known non-matches is quite symmetric, while the distribution for each feature for the known matches is skewed left.'}
@@
 
-custom <- list(
-  theme_bw(),
-  scale_linetype_manual("Bullet Land Pairs", values=3:2),
-  scale_colour_manual("Bullet Land Pairs", values=c("black", "grey40"))
-  )
-
-ggplot(data = cor_compare, aes(x = value, colour=match, linetype = match)) +
+ggplot(data = cor_compare, aes(x = value, color = match)) +
   stat_ecdf() +
   facet_wrap(~feature, nrow = 2) +
-  ylab("") + xlab("") +
+  theme_bw() + ylab("") + xlab("") +
+  scale_color_brewer("Bullet Land Pairs", palette="Paired") +
   theme(legend.background = element_rect(colour="grey75"),
         axis.title.y=element_blank(),
-        plot.margin = unit(c(0,0,0,0), unit="cm")) +
-  custom +
-    theme(legend.position=c(0.2, 0.85)) 
+        plot.margin = unit(c(0,0,0,0), unit="cm"))
 ```
@@
   select(id, everything()) %>%
-  gather(key = Feature, value = Value, 3:ncol(.)) %>%
-  mutate(x = as.numeric(as.factor(Feature)))
-
-truepos <- filter(CCFs_tall, Type %in% c("True Positive"))
-falsepos <- filter(CCFs_tall, Type %in% c("False Positive"))
-
-truepos_poly <- truepos %>% 
-  group_by(Feature) %>% 
-  summarize(min = min(Value), max=max(Value), mean=mean(Value), x=x[1])
-tpps <- truepos_poly %>% gather(Type, Result, min:max)
-tpps$order <-c(1:8, 9+8:1)
-tpps$x <-c(1:8, 1:8)
-
-tpps <- tpps %>% arrange(order)
+  gather(key = Feature, value = Value, 3:ncol(.))
 
 p1 <- ggplot() +
-  geom_polygon(aes(x = x, y=Result), data = tpps, fill="grey50", alpha=.35) +
-  geom_line(aes(x = x, y = mean), data = truepos_poly, colour="grey50", size=1) +
-  geom_line(data = truepos, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 0.1, size = 0.2) + #, linetype = "dashed") +
-  scale_x_continuous(breaks=1:8, labels=unique(tpps$Feature)) +
-  geom_line(data = falsepos, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 1, size = 0.8) +
+  geom_line(data = filter(CCFs_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
+  geom_line(data = filter(CCFs_tall, Type %in% c("False Positive")), aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.8, linetype = "dashed") +
+  geom_line(data = filter(pcp_data_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
   ylab("Normalized Value") +
-  xlab("") +
   theme_bw() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_color_manual(values = c("grey10", "grey50")) +
-#  scale_color_manual(values = c("#e60000", "#000099")) +
-  scale_size_manual(values = c(1, 1)) +
-  scale_linetype_manual(values=2:1)
-
-# p1 <- ggplot() +
-# #  geom_polygon(aes(x = x, y=Result), data = tpps, fill="#000099", alpha=.5) +
-#   geom_line(data = truepos, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
-#   geom_line(data = falsepos, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.8, linetype = "dashed") +
-#   geom_line(data = filter(pcp_data_tall, Type %in% c("True Positive")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
-#   ylab("Normalized Value") +
-#   theme_bw() +
-#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-#   scale_color_manual(values = c("#e60000", "#000099")) +
-#   scale_size_manual(values = c(1, 1))
+  scale_color_manual(values = c("#e60000", "#000099")) +
+  scale_size_manual(values = c(1, 1))
 
@@
 mysamp2 <- sample(CCFs_tall$id[CCFs_tall$Type == "False Negative"], 20)
-
-CCFs_tall <- CCFs_tall %>% mutate(x = as.numeric(as.factor(Feature)))
-
 trueneg_sample <- CCFs_tall %>%
@@
 
-trueneg_poly <- trueneg_sample %>% 
-  group_by(Feature) %>% 
-  summarize(min = min(Value), max=max(Value), mean=mean(Value), x=x[1])
-tnps <- trueneg_poly %>% gather(Type, Result, min:max)
-tnps$order <-c(1:8, 9+8:1)
-tnps$x <-c(1:8, 1:8)
-
-tnps <- tnps %>% arrange(order)
-
 p2 <- ggplot() +
-  geom_polygon(aes(x = x, y=Result), data = tnps, fill="grey50", alpha=.35) +
-  geom_line(aes(x = x, y = mean), data = trueneg_poly, colour="grey50", size=1) +
-  geom_line(data = trueneg_sample, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 0.1, size = 0.2) + #, linetype = "dashed") +
-scale_x_continuous(breaks=1:8, labels=unique(tpps$Feature)) +
-  geom_line(data = falseneg_sample, aes(x = x, y = Value, colour = Type, group = id, linetype=Type), alpha = 1, size = 0.8) +
+  #geom_line(data = filter(CCFs_tall, Type %in% c("True Negative", "False Negative")), aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type, size = Type)) +
+  geom_line(data = trueneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
+  geom_line(data = falseneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.4, linetype = "dashed") +
+  theme_bw() +
+  geom_line(data = filter(pcp_data_tall, Type %in% c("True Negative")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
   ylab("Normalized Value") +
-  xlab("") +
-  theme_bw() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_color_manual(values = c("grey10", "grey50")) +
-#  scale_color_manual(values = c("#e60000", "#000099")) +
-  scale_size_manual(values = c(1, 1)) +
-  scale_linetype_manual(values=2:1)
-
-# p2 <- ggplot() +
-#   #geom_line(data = filter(CCFs_tall, Type %in% c("True Negative", "False Negative")), aes(x = Feature, y = Value, colour = Type, group = id, alpha = Type, size = Type)) +
-#   geom_line(data = trueneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 0.1, size = 0.2, linetype = "dashed") +
-#   geom_line(data = falseneg_sample, aes(x = Feature, y = Value, colour = Type, group = id), alpha = 1, size = 0.4, linetype = "dashed") +
-#   theme_bw() +
-#   geom_line(data = filter(pcp_data_tall, Type %in% c("True Negative")), aes(x = Feature, y = Value, colour = Type, group = Type), size = 2) +
-#   ylab("Normalized Value") +
-#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-#   scale_color_manual(values = c("#e60000", "#000099"))
+  scale_color_manual(values = c("#e60000", "#000099"))
 
@@
 ## Produce plot
-ggplot(result, aes(x = level, y = value, colour = fixed, linetype=fixed)) +
+ggplot(result, aes(x = level, y = value, colour = fixed)) +
     facet_wrap(~metric, nrow = 2) +
@@
     scale_x_continuous(breaks = c(0, .125, .25, .375, .5, .625, .75), labels = c("100%", "87.5%", "75%", "62.5%", "50%", "37.5%", "25%")) +
-    xlab("Land Proportion") +
-  scale_colour_manual(values=c("black", "grey20", "grey30", "grey40")) +
-  theme(legend.key.width=unit(3,"line"))
+    xlab("Land Proportion")
 ```
@@
 
-ggplot(data = feature_exp, aes(x = level, y = value, colour = fixed, linetype=fixed)) +
+ggplot(data = feature_exp, aes(x = level, y = value, colour = fixed)) +
     geom_point() +
@@
     xlab("Land Proportion") +
-    facet_wrap(~feature, scales = "free_y") +
-    scale_colour_manual(values=c("black", "grey20", "grey30", "grey40")) +
-  theme(legend.key.width=unit(3,"line"))
-
+    facet_wrap(~feature, scales = "free_y")
 ```
@@
 
-ggplot(data = feature_exp2, aes(x = level, y = value, colour = fixed, linetype=fixed)) +
+ggplot(data = feature_exp2, aes(x = level, y = value, colour = fixed)) +
     geom_point() +
@@
     theme_bw() +
-    theme(legend.position="none",
+    theme(legend.background = element_rect(colour="grey75"),
           axis.text.y = element_blank(),
-          axis.title.y = element_blank()) +
-  annotate("text", x=2.25, y= 0.8, label="Known matches (KM)", hjust=1, colour="grey40") +
-  annotate("text", x=0.75, y= 0.05, label="Known non-matches (KNM)", hjust=0, colour="grey40")
-
+          axis.title.y = element_blank())
 ```
@@
 
+This research was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
+
 \clearpage

### 2017-10-22 degraded-paper.tex
I don't remember what changes these were

@@
 
+This research was partially funded by the Center for Statistics and
+Applications in Forensic Evidence (CSAFE) through Cooperative Agreement
+\#70NANB15H176 between NIST and Iowa State University, which includes
+activities carried out at Carnegie Mellon University, University of
+California Irvine, and University of Virginia.
+
 \clearpage

### 2017-10-22 degraded-paper.tex
Rebuild the pdf I guess

@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{degraded-paper_files/figure-latex/roughcordist-1.pdf}
-\caption{\label{fig:roughcordist}Empirical CDFs of the Roughness Correlation
-compared to the CCF for known matches and known non-matches. It can be
-seen that the distribution of each feature for the known non-matches is
-quite symmetric, while the distribution for each feature for the known
-matches is skewed left.}
+\begin{figure}
+\includegraphics[width=.5\textwidth]{degraded-paper_files/figure-latex/roughcordist-1} \caption{Empirical CDFs of the Roughness Correlation compared to the CCF for known matches and known non-matches.
+It can be seen that the distribution of each feature for the known non-matches is quite symmetric, while the distribution for each feature for the known matches is skewed left.}\label{fig:roughcordist}
 \end{figure}
@@
   \hline
-False Negative & 20.7 \\ 
-  False Positive & 1.9 \\ 
-  True Negative & 3716.8 \\ 
-  True Positive & 56.5 \\ 
+False Negative & 20.6 \\ 
+  False Positive & 2.0 \\ 
+  True Negative & 3716.7 \\ 
+  True Positive & 56.6 \\ 
    \hline
@@
   \hline
-Hamby252\_Hamby252 & 0.39\% & 0.02\% & 97.56\% & 2.03\% \\ 
-  Hamby252\_Hamby44 & 0.45\% & 0.05\% & 98.31\% & 1.19\% \\ 
-  Hamby44\_Hamby44 & 0.89\% & 0.09\% & 97.68\% & 1.34\% \\ 
+Hamby252\_Hamby252 & 0.38\% & 0.02\% & 97.56\% & 2.04\% \\ 
+  Hamby252\_Hamby44 & 0.45\% & 0.05\% & 98.3\% & 1.19\% \\ 
+  Hamby44\_Hamby44 & 0.88\% & 0.09\% & 97.68\% & 1.35\% \\ 
    \hline
@@
   \hline
-Hamby252\_Hamby252 & 09.6\% & 08.32\% & 11.28\% \\ 
-  Hamby252\_Hamby44 & 81.47\% & 81.06\% & 78.68\% \\ 
-  Hamby44\_Hamby44 & 08.93\% & 10.63\% & 10.04\% \\ 
+Hamby252\_Hamby252 & 09.93\% & 08.47\% & 011.3\% \\ 
+  Hamby252\_Hamby44 & 81.24\% & 80.82\% & 78.27\% \\ 
+  Hamby44\_Hamby44 & 08.83\% & 010.7\% & 10.43\% \\ 
    \hline
